### 2020-02-03 README.Rmd
README.md / rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eurobarometer

<!-- badges: start -->
<!-- badges: end -->

The goal of eurobarometer is to ...

## Installation

You can install the released version of eurobarometer from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("eurobarometer")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("antaldaniel/eurobarometer")
```
## Reading in an SPSS file from GESIS

First of all, we create a representation of the GESIS archive SPSS file that can be programatically used in R.  One problem is that SPSS variable names may contain reserved characters. For further programmatic use in R, the best is to rename the variables.

```{r example, eval=FALSE}
library(eurobarometer)
library(eurobarometer)
gesis_spss_read ( path = file.path('data-raw', 'example.sav'), 
                  rename = TRUE, 
                  unique_id = TRUE)
```

This setting the function calls `canonical_file_rename()` which removes spaces and special characters, and converts all variable names to lowercase. Similarly, by default the serial
identifiers of observations (rows) in the SPSS file are brought to the harmonized name `unique_id`. This ID is unique in the single survey data file, but not if you create panel trend files.

If you are planning to work in R with certain files, it makes sense to save the result of this function as an R data object with `save()` or `saveRDS()` somewhere. Conversion from SPSS is a rather resource-intensive operation.

### 2020-02-03 README.Rmd
hopefully

@@
 
-You can install the released version of eurobarometer from [CRAN](https://CRAN.R-project.org) with:
-
-``` r
-install.packages("eurobarometer")
-```
-
-And the development version from [GitHub](https://github.com/) with:
+At this moment yuo cannot install yest the packgae from [CRAN](https://CRAN.R-project.org), but there is a development version on [GitHub](https://github.com/):
 

### 2020-05-13 R/gesis_metadata_create.R
convert_class

@@ -25,14 +25,16 @@ gesis_metadata_create <- function ( dat ) {
 
   r_class   <- vapply ( dat, class, character(1) )
   spss_name <- vapply ( dat, sjlabelled::get_label, character(1) )
-  canonical_name <- canonical_name_create(spss_name)
+  normalized_names <- normalize_names(spss_name)
+
+
 
   ##Creating the basic metadata ----
   metadata <- data.frame (
     r_name = names ( dat ),
     r_class = r_class,
     spss_name = spss_name,
-    canonical_name = canonical_name
+    normalized_name = normalized_names
   )
 
 
@@ -52,31 +54,5 @@ gesis_metadata_create <- function ( dat ) {
     sapply ( metadata$factor_levels, unlist ),
     length, numeric(1) )  #number of categories in categorical variables
 
-  ##Adding constants -----
-
-  if ( "doi" %in% names(dat)) {
-    metadata$doi = dat$doi[1]
-  } else { metadata$doi <- NA_character_ }
-
-  if ( "version" %in% names(dat)) {
-    metadata$version = dat$version[1]
-  } else { metadata$version <- NA_character_ }
-
-  study_number <- which ( metadata$canonical_name == "study_number_distributor")
-  gesis_archive_version <-  which ( metadata$canonical_name == "gesis_archive_version")
-  id_serial_number <- which ( metadata$canonical_name == "id_serial_number")
-
-  if ( length(study_number) > 0 ) {
-    metadata$study_number <- as.character(dat [1, study_number])
-  } else {
-    metadata$study_number<- NA
-  }
-
-  if ( length(gesis_archive_version) > 0 ) {
-    metadata$gesis_archive_number <- as.character(dat [1, gesis_archive_version])
-  } else {
-    metadata$gesis_archive_number <- NA
-  }
-
- metadata
+  suggest_conversion( metadata)
 }

### 2020-05-13 R/normalize_names.R

@@ -1,12 +1,12 @@
-#' Create a canonical variable name for the R objects
+#' Create Normalized Variable Names for GESIS Columns
 #'
 #' @param x A vector of the GESIS variable names
 #' @importFrom stringr str_sub
 #' @examples
-#' canonical_name_create ( c("UPPER CASE VAR", "VAR NAME WITH % SYMBOL") )
+#' normalize_names ( c("UPPER CASE VAR", "VAR NAME WITH % SYMBOL") )
 #' @export
 
-canonical_name_create <- function(x) {
+normalize_names <- function(x) {
 
   x <- trimws(tolower(as.character(x)))
 

### 2020-05-13 R/utils.R

@@ -1,0 +1,27 @@
+#' Suggest Conversion
+#'
+#' @param metadata A metadata file created by gesis_metadata_create
+#' @importFrom magrittr %>%
+#' @importFrom dplyr select mutate case_when
+
+suggest_conversion <- function(metadata) {
+
+  factor_levels <- sapply ( metadata$factor_levels,
+                        function(x) sort(tolower(unlist(x))) )
+
+  first_level  <- unlist(lapply(factor_levels, function(x) x[1]))
+  second_level <- unlist(lapply(factor_levels, function(x) x[2]))
+
+  suggestions <- metadata %>%
+    dplyr::select ( r_class,factor_levels, level_length ) %>%
+    mutate ( first_level  = first_level,
+             second_level = second_level ) %>%
+    mutate ( suggestion = case_when (
+      r_class %in% c("numeric","character") ~ r_class,
+      r_class == 'haven_labelled' & level_length == 1 ~ 'character',
+      first_level == "mentioned" & second_level == "not mentioned" ~  'dummy',
+      TRUE ~ "factor" ))
+
+  metadata$conversion_suggestion <- suggestions$suggestion
+  metadata
+}

### 2020-05-13 data-raw/create_vocabulary.R

@@ -1,7 +1,0 @@
-
-eurobarometer_vocabulary <- read.csv ( file.path ( 'data-raw',
-                                                   'vocabulary.csv'),
-           stringsAsFactors = FALSE)
-
-
-usethis::use_data(eurobarometer_vocabulary, internal=TRUE, overwrite=TRUE)

### 2020-05-13 README.Rmd
New README

@@
 <!-- badges: start -->
+[![Codecov test coverage](https://codecov.io/gh/antaldaniel/eurobarometer/branch/master/graph/badge.svg)](https://codecov.io/gh/antaldaniel/eurobarometer?branch=master)
+[![Project Status: Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
+"[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)"
+"[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/regions)](https://cran.r-project.org/package=eurobarometer)"
 <!-- badges: end -->
 
-The goal of eurobarometer is to ...
+The goal of eurobarometer is converting Eurobarometer microdata files, as stored by GESIS, into tidy R data frames and help common pre-processing problems.
+
+Please report all issues to [github.com/antaldaniel/eurobarometer/issues](https://github.com/antaldaniel/eurobarometer/issues).
+
+Pull requests are welcome on [github.com/antaldaniel/eurobarometer]((https://github.com/antaldaniel/eurobarometer/issues))  from all potential contributors who abide by  the terms of the [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
 
@@
 
-At this moment yuo cannot install yest the packgae from [CRAN](https://CRAN.R-project.org), but there is a development version on [GitHub](https://github.com/):
+At this moment you cannot install yest the packgae from [CRAN](https://CRAN.R-project.org), but there is a development version on [GitHub](https://github.com/):
 
@@
 Conversion from SPSS is a rather resource-intensive operation.
+
+## Code of Conduct
+
+Please note that the eurobarometer project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
+By contributing to this project, you agree to its terms.

### 2020-05-13 README.Rmd
README

@@
 "[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/regions)](https://cran.r-project.org/package=eurobarometer)"
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3825700.svg)](https://doi.org/10.5281/zenodo.3825700)
 <!-- badges: end -->
@@
 Pull requests are welcome on [github.com/antaldaniel/eurobarometer]((https://github.com/antaldaniel/eurobarometer/issues))  from all potential contributors who abide by  the terms of the [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
+
+If you use `eurobarometer` in your work, please [cite the package](https://doi.org/10.5281/zenodo.3825700).
 

### 2020-05-13 R/canonical_names.R
remove unnecessary old functions

@@ -1,9 +1,14 @@
-#' Create Normalized Variable Names for GESIS Columns
+#' Create Canonical Names for GESIS Columns
 #'
-#' @param x A vector of the GESIS variable names
+#' Create canonical names that do not vary across several SPSS files
+#' from different years.
+#' @param metadata A metadata data frame created by
+#' \code{\link{gesis_metadata_create}}
 #' @importFrom stringr str_sub
+#' @family naming functions
 #' @examples
-#' normalize_names ( c("UPPER CASE VAR", "VAR NAME WITH % SYMBOL") )
+#' canonical_names ( c("digital_object_identifier",
+#'                     "tns_unique_case_id") )
 #' @export
 
 canonical_names <- function(metadata) {
@@ -25,8 +30,6 @@ canonical_names <- function(metadata) {
   x <- ifelse ( grepl("wex_weight_extra", x),
                 "wex",
                 x)
-
-
 
   x <- gsub( '\\s', '_', x)
   x <- gsub( '___', '_', x)

### 2020-05-13 R/eurobarometer.R

@@ -1,0 +1,16 @@
+#' eurobarometer: A package to create tidy Eurobarometer microdata files.
+#'
+#' The regions package provides four categories of important functions:
+#' validate, recode, impute and aggregate.
+#'
+#' @section naming functions:
+#' The naming functions make the GESIS SPSS files usable in a programmatic 
+#' context.
+#' \code{\link{normalize_names}} removes regex symbols, whitespace, and 
+#' basic inconsistencies in abbreviations.
+#' \code{\link{canonical_names}} harmonizes the names across various GESIS
+#' files, so that they can be joined into panels across time.
+#' 
+#' @docType package
+#' @name eurobarometer
+NULL

### 2020-05-13 R/normalize_names.R

@@ -2,6 +2,7 @@
 #'
 #' @param x A vector of the GESIS variable names
 #' @importFrom stringr str_sub
+#' @family naming functions
 #' @examples
 #' normalize_names ( c("UPPER CASE VAR", "VAR NAME WITH % SYMBOL") )
 #' @export

### 2020-05-13 R/recode_age_education.R

@@ -1,57 +1,0 @@
-#' @title Recode Age Education
-#'
-#' If age exact is labelled to a number less than 10, coded to 15.
-#' Refused answers are coded to missing.
-#' Still studying answers are recoded to the exact age, when given,
-#' otherwise to missing.
-#' @param var Age Education variable
-#' @param age_exact The exact age variable, defaults to \code{NULL}.
-#' @importFrom dplyr case_when
-#' @importFrom haven as_factor
-#' @family recode functions
-#' @examples
-#' #Exact age is given
-#' recode_age_education  (
-#'  var =  c( "2 years", "75 years", NA_real_,
-#'            "Still studying", "Refusal",
-#'            "No full-time education"),
-#'  age_exact = c(87, 75, 34,23, 19, 45)
-#'  )
-#'
-#' #Exact age is not given, school living is recoded to missing
-#' recode_age_education  (
-#'  var =  c( "2 years", "75 years", NA_real_,
-#'            "Still studying", "Refusal",
-#'            "No full-time education")
-#'            )
-#' @export
-
-recode_age_education <- function ( var, age_exact = NULL ) {
-
-  if ( class(var) == "haven_labelled" )  {
-    var <- haven::as_factor (var)
-  }
-
-  if ( is.null(age_exact)) age_exact = as.character(var)
-  age_exact <- as.character(age_exact)
-
-  age_education_special <- tolower(as.character(var))
-  var [ is.na(age_education_special)]
-
-  tested <- case_when (
-    age_education_special == "no full-time education" ~ "15",
-    grepl ( " years", age_education_special) ~ gsub(" years", "", as.character(age_education_special)) ,
-    age_education_special == "still studying" ~ as.character(age_exact),
-    age_education_special %in% c("dK", "declined", "refusal", "na")  ~ NA_character_,
-    is.na(age_education_special) ~ NA_character_,
-    TRUE ~ as.character(age_education_special)
-  )
-
-  case_when (
-    as.numeric ( tested ) < 15 ~ 15,
-    TRUE ~ as.numeric(tested)
-    )
-
-}
-
-

### 2020-05-13 R/recode_gender.R

@@ -1,21 +1,0 @@
-#' @title Recode gender variable
-#'
-#' The harmonize demography function create a harmonized labelling or value
-#' setting for standard demography variables.
-#'
-#' @param var The gender variable, such as \code{gender}, \code{D10 GENDER},
-#' or the Turkish Cypriot Community equivalents.
-#' @family recode functions
-#' @examples
-#' #not case sensitive
-#' recode_gender  ( var = c("Man", "Female", "man", "Woman" ))
-#' @export
-
-
-
-recode_gender <- function ( var ) {
-
-  var <- tolower(as.character(var))
-  ifelse ( var %in% c("man", "male"), "m", "f")
-
-}

### 2020-05-13 R/recode_mentioned.R

@@ -1,24 +1,0 @@
-#' @title Recode A Binary Variable in Multiple Choice
-#'
-#' @param var A binary variale with multiple choice options
-#' @family recode functions
-#' @examples
-#' #not case sensitive
-#' recode_binary  ( var = c("Mentioned", "Not mentioned", "mentioned", "Refusal", "DK") )
-#' @export
-
-recode_mentioned <- function ( var ) {
-
-  if ( class(var) == "haven_labelled" )  {
-    var <- haven::as_factor (var)
-  }
-
-  dplyr::case_when (
-    tolower(as.character(var)) == "mentioned" ~ 1,
-    tolower(as.character(var)) == "not mentioned" ~ 0,
-    TRUE ~ NA_real_)
-
-}
-
-
-

### 2020-05-13 R/rename_var.R

@@ -1,46 +1,0 @@
-#' @title Harmonize gender variable
-#'
-#' @param df A data frame like object with the raw statistical data.
-#' @param harmonized_name The harmonized name of the variable.
-#' @return The \code{df} with the gender binary variable in the data frame
-#' names as \code{gender}
-#' @examples
-#' \dontrun{
-#' gesis_spss_read ( path = file.path('data-raw', 'example.sav'),
-#'                   rename = TRUE,
-#'                   unique_id = TRUE ) %>%
-#'    recode_var (., harmonized_name = "gender")
-#' }
-#' @export
-
-rename_var <- function ( df,
-                         harmonized_name = "gender" ) {
-
-  . <- canonical_name <- harmonized_var <- NULL
-
-
-  ## Check if the harmonized name can be found in the vocabulary --------
-
-  if ( ! harmonized_name %in% eurobarometer_vocabulary$harmonized_name ) {
-    stop ( "Variable '", harmonized_name, "' cannot be found in the internal package vocabulary." )
-  }
-
-  names_df <- names ( df )
-
-  harmonize_vars <- eurobarometer_vocabulary %>%
-    dplyr::filter ( harmonized_name == harmonized_name ) %>%
-    dplyr::select ( harmonized_name, canonical_name )
-
-  if ( any (harmonize_vars$canonical_name %in% names (df)) ) {
-
-    possible_harmonized_var <-  names_df [ names_df %in% harmonize_vars$canonical_name]
-
-    if ( length( possible_harmonized_var ) != 1 ) {
-      stop ( "Could not identify '", harmonized_name , "' variables.\nCandidates are ", possible_harmonized_var )
-    } else {
-      names(df)[which (  names(df)  == possible_harmonized_var)] <- harmonized_var
-    }
-
-  }
-  df
-}

### 2020-05-17 README.Rmd
Add NEWS.md

@@
 ```
-## Reading in an SPSS file from GESIS
 
-First of all, we create a representation of the GESIS archive SPSS file that can be programatically used in R.
-One problem is that SPSS variable names may contain reserved characters.
-For further programmatic use in R, the best is to rename the variables.
+## Data Preparation For Programmatic Use
 
-```{r example, eval=FALSE}
-library(eurobarometer)
-library(eurobarometer)
-gesis_spss_read ( path = file.path('data-raw', 'example.sav'), 
-                  rename = TRUE, 
-                  unique_id = TRUE)
-```
+The SPSS files in GESIS are not ready for programmatic use in R, because the variable names often contain reserved special characters.
+You should use `gesis_metadta_create()` to understand what are the possible problems in the SPSS file and how to resolve them.
+(See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html))
 
-This setting the function calls `canonical_file_rename()` which removes spaces and special characters, and converts all variable names to lowercase. Similarly, by default the serial
-identifiers of observations (rows) in the SPSS file are brought to the harmonized name `unique_id`. This ID is unique in the single survey data file, but not if you create panel trend files.
+* The naming functions `normalize_names()` removes regex symbols, whitespace, and basic inconsistencies in abbreviations.
+The function `canonical_names()`harmonizes the names across various GESIS files, so that they can be joined into panels across time.
 
-If you are planning to work in R with certain files, it makes sense to save the result of this function as an R data object with `save()` or `saveRDS()` somewhere.
-Conversion from SPSS is a rather resource-intensive operation.
+* You can read in the SPSS file to R in haven, which will result in many `labelled` variables.
+The metadata file contains a suggestion of the most practical class conversion for data analysis.
+You can review and change this suggestion and then make the conversions at once with `convert_class()`
+
+## Joining Serveral Files
+
+The vocabulary functions help to create data panels across Eurobarometer surveys taken in different times.
+
+There are two main issues that need to be resolved in such cases
+
+*  Make sure that the same questions are used, and they are consistently named (this problem is treated in the metadata level)
+
+* In case of structured questions or question blocks the answer options are coded exactly the same (for example “Fully agree”and “Totally agree” are the same.) This will be handled by the vocabulary functions.
+(Not yet implemented.)
+
+## Joining With Eurostat & Other Data Tables
+
+*  The regional boundaries are consistently used, coded and named - this will be harmonized with the package `regions` , currently under review in CRAN, which will help resolving common issues, and to properly join various Eurobarometer files with Eurostat data tables and Google data tables.
+
+* This will also allow you to create regional statistics from Eurobarometer microdata files.
+
+* Currently `code_nuts1()` and `code_nuts2()` is implemented but this will be replaced with a more comprehensive NUTS0-NUTS3 level helper function to make the regional coding consistent.
 
@@
 
-Please note that the eurobarometer project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
+Please note that the `eurobarometer` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
 By contributing to this project, you agree to its terms.

### 2020-05-17 README.Rmd
README

@@
 
-*  The regional boundaries are consistently used, coded and named - this will be harmonized with the package `regions` , currently under review in CRAN, which will help resolving common issues, and to properly join various Eurobarometer files with Eurostat data tables and Google data tables.
+*  The regional boundaries are consistently used, coded and named - this will be harmonized with the package [regions](http://regions.danielantal.eu/), currently under review in CRAN, which will help resolving common issues, and to properly join various Eurobarometer files with Eurostat data tables and Google data tables.
 

### 2020-06-04 README.Rmd
README

--- 2020-05-17 README.Rmd
@@
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3825700.svg)](https://doi.org/10.5281/zenodo.3825700)
+[![Follow author](https://img.shields.io/twitter/follow/antaldaniel.svg?style=social)](https://twitter.com/intent/follow?screen_name=antaldaniel)
+
 <!-- badges: end -->
@@
 The SPSS files in GESIS are not ready for programmatic use in R, because the variable names often contain reserved special characters.
-You should use `gesis_metadta_create()` to understand what are the possible problems in the SPSS file and how to resolve them.
+You should use `gesis_metadata_create()` to understand what are the possible problems in the SPSS file and how to resolve them.
 (See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html))
@@
 
-* You can read in the SPSS file to R in haven, which will result in many `labelled` variables.
+* You can read in the SPSS file to R in [haven](https://haven.tidyverse.org/), which will result in many `labelled` variables.
 The metadata file contains a suggestion of the most practical class conversion for data analysis.
 You can review and change this suggestion and then make the conversions at once with `convert_class()`
+
+See `vignette("metadata")`
 
@@
 
-*  The regional boundaries are consistently used, coded and named - this will be harmonized with the package `regions` , currently under review in CRAN, which will help resolving common issues, and to properly join various Eurobarometer files with Eurostat data tables and Google data tables.
+*  The regional boundaries are consistently used, coded and named - this will be harmonized with the package [regions](http://regions.danielantal.eu/), which helps resolving common issues, and to properly join various Eurobarometer files with Eurostat data tables and Google data tables.
 

### 2020-06-11 R/data_eurobarometer_corpus.R
documentation

@@ -10,6 +10,8 @@
 #'   \item{numeric_value}{The original SPSS numeric values, if applicable}
 #'   \item{character_value}{The original SPSS character value, if applicable}
 #'   \item{label}{The label of the questionnaire item}
+#'   \item{item_no}{The rank number of the alphabetically sorted answer option.}
+#'   \item{item_of}{The number of answer options in the questionnaire item.}
 #'   \item{label_normalized}{Normalized version of the questionnaire item}
 #'   \item{filename}{The name of the original GESIS file.}
 #' }

### 2020-06-11 R/gesis_metadata_create.R

@@ -1,7 +1,7 @@
-#' Get a metadata file
+#' Get A Metadata File
 #'
-#' Retrieve a metadata file with from your data directory or the temporary directory
-#' of your current R session.
+#' Retrieve a metadata file with from your data directory or the
+#' temporary directory of your current R session.
 #' @param zacat_id For example, \code{"ZA5688_v6-0-0"}, as analyzed and saved by
 #' \code{\link{analyze_gesis_file}}.
 #' Defaults to \code{"last_data_frame"}, which is
@@ -55,5 +55,5 @@ gesis_metadata_create <- function ( dat ) {
     sapply ( metadata$factor_levels, unlist ),
     length, numeric(1) )  #number of categories in categorical variables
 
-  suggest_conversion( metadata)
+  suggest_conversion(metadata)
 }

### 2020-06-11 data-raw/create_corpus.R

@@ -4,9 +4,12 @@ gesis_dir <- file.path("C:/Users/Daniel Antal/OneDrive - Visegrad Investments/",
 gesis_files <- file.path(gesis_dir,
           dir(gesis_dir)[grepl(".sav", dir(gesis_dir))])
 
-dat = gesis_files[2]
 
 this_corpus <- gesis_corpus_create( file_path=gesis_dir)
 eurobarometer_corpus <- this_corpus
 
 usethis::use_data(eurobarometer_corpus, overwrite =TRUE, internal = FALSE)
+
+
+
+

### 2020-06-13 R/gesis_metadata_create.R
renaming internal variables

@@ -1,18 +1,30 @@
 #' Get A Metadata File
 #'
+#' @description
 #' Retrieve a metadata file with from your data directory or the
 #' temporary directory of your current R session.
+#'
+#' @details The structure of the returned tibble (data frame)
+#' \emph{var_name_orig}: The original variable name in SPSS;
+#' \emph{class_orig}: The original variable class after importing
+#' \emph{class_orig}: The original variable class after importing
+#' with \code{\link{haven::read_spss}};
+#' \emph{var_label_orig}: The original variable label in SPSS;
+#' \emph{var_label_norm}: A normalized version of the variable labels;
+#' \emph{factor_levels}: A list of factor levels, i.e. value labels in SPSS;
+#' \emph{class_suggested}: A suggested class conversion.
+#'
 #' @param zacat_id For example, \code{"ZA5688_v6-0-0"}, as analyzed and saved by
 #' \code{\link{analyze_gesis_file}}.
 #' Defaults to \code{"last_data_frame"}, which is
 #' the last file in the temporary directory of the current session.
 #' @param data_dir Defaults to \code{"NULL"}. In this case the \code{"tempdir()"}
 #' of your current R session will be used.
-#' @importFrom sjlabelled get_label get_labels
+#' @importFrom labelled val_labels var_label
 #' @importFrom dplyr full_join
-#' @seealso canonical_name_create
-#' @return A data frame with the original file names, the suggested canonical
-#' names, factor lists, suggested class conversions.
+#' @importFrom tibble tibble
+#' @return A data frame with the original variable attributes and
+#' suggested conversions and changes.
 #' @examples
 #' \dontrun{
 #' ##use your own file:
@@ -23,37 +35,40 @@
 
 gesis_metadata_create <- function ( dat ) {
 
-  r_class   <- vapply ( dat,
-                        function(x) class(x)[1],
-                        character(1) )
-  spss_name <- vapply ( dat, sjlabelled::get_label, character(1) )
-  normalized_names <- normalize_names(x = spss_name )
-  normalized_names
-
-  ##Creating the basic metadata ----
-  metadata <- data.frame (
-    r_name = names ( dat ),
-    r_class = r_class,
-    spss_name = spss_name,
-    normalized_names = normalized_names
-  )
-
-
-  ##Creating a catalog of possible categories / factor levels ----
-  value_labels <- sapply ( dat, sjlabelled::get_labels )
[…]

### 2020-06-13 R/gesis_vocabulary_create.R

@@ -2,23 +2,22 @@
 #'
 #' Get every single unique questionnaire item from a survey file.
 #'
-#' @param dat For example,
-#' @importFrom sjlabelled get_label get_labels
+#' @details The variables in the returned data frame
+#' \describe{
+#'   \item{var_label_orig}{The name of the variable in the data file.}
+#'   \item{val_numeric_orig}{The original SPSS numeric values, if applicable}
+#'   \item{val_label_orig}{The original SPSS character value, if applicable}
+#'   \item{val_order_alpha}{The rank number of the alphabetically sorted answer option.}
+#'   \item{val_order_length}{The number of answer options in the questionnaire item.}
+#'   \item{val_label_normalized}{Normalized version of the questionnaire item}
+#' }
+#' @param dat A data frame read by \code{\link{haven::read_spss}}.
 #' @importFrom dplyr bind_rows mutate arrange
-#' @importFrom labelled val_labels
+#' @importFrom labelled var_label
 #' @importFrom tibble tibble
 #' @seealso canonical_name_create
 #' @return The vocabulary of the survey in a data frame. The return
 #' data frame has 7 columns:
-#' \describe{
-#'   \item{r_name}{The name of the variable in the data file.}
-#'   \item{numeric_value}{The original SPSS numeric values, if applicable}
-#'   \item{character_value}{The original SPSS character value, if applicable}
-#'   \item{label}{The label of the questionnaire item}
-#'   \item{item_no}{The rank number of the alphabetically sorted answer option.}
-#'   \item{item_of}{The number of answer options in the questionnaire item.}
-#'   \item{label_normalized}{Normalized version of the questionnaire item}
-#' }
 #' @examples
 #' \dontrun{
 #' ##use your own file:
@@ -29,66 +28,83 @@
 
 gesis_vocabulary_create <- function ( dat ) {
 
-  r_class   <- vapply ( dat,
-                        function(x) class(x)[1],
-                        character(1) )
-  spss_name <- vapply ( dat, sjlabelled::get_label, character(1) )
-  normalized_names <- normalize_names(x = spss_name )
-  normalized_names
+  class_orig   <- vapply (
+    dat, function(x) class(x)[1], character(1)
+  )
 
-  get_items <- function ( r_name  = 'nuts') {
-    itemize <- dat[, r_name ]
+  var_label_orig <- vapply ( dat, labelled::var_label, character(1) )
[…]

### 2020-06-17 data-raw/trust_tables.Rmd
new vignette

---
title: "Creating Variable and Values Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
options(width = 160)
```

```{r setup}
library(eurobarometer)
library(dplyr)
library(tibble)
library(knitr)
library(kableExtra)
# The examples of this vignette can be found run with  
# source(
#  file.path("not_included", "vignette_vocabulary_examples.R")
#  )
```

## Procedure

The procedure involves three steps:

1. Concept definition: Trust in institutions: general trust without speficying the concrete domains of activity where the trust applies (these specific questions would require special treatment).

2. Selection of survey questions that correspond to the concept by filtering on variable and value labels.

3. Standardizing value labels across questions.

4. Standardizing variable names to have the same name for all variables that correspond to the same question.

## Metadata

`metadata_database` is a table with variable metadata created from Eurobarometer SPSS files. It has the following columns:

`var_name_orig`: variable name in the original dataset  
`class_orig`: column class in the original dataset   
`var_label_orig`: variable label in the original dataset  
`var_label_norm`: normalized variable label  
`var_name_suggested`: suggested variable label  
`factor_levels`: vector of value labels as a list (they have no fixed length)  
`n_categories`: total number of response options `= length(factor_levels)`   
`class_suggested`: suggested conversion to an R class, it should correspond later to a conversion function, so that the researcher can just simply approve and get a correct R representation.  
`filename`: original name of file as obtained from GESIS; it contains a disambigous version information, too.
`val_numeric_orig`: numeric value code (if available, empty otherwise)   
`val_label_orig`: value label in the original dataset corresponding to a response option
`val_order_alpha`: alphabetical number (position) of response option in the set of value labels, aftering sorting with `sort()` (we use alphabetical order instead of `levels()` because levels may not be the same in different survey files, and we revert to a basic but disambigous sorting)   
`val_order_length`:  
`val_label_norm`: value label normalized with `label_normalize()`

```{r}
metadata_database <-  readRDS( 
  file.path('..', 'data-raw',
  'eb_metadata_database_large.rds')) 
```

## Selection of variables

This is an iterative process.

### Iteration 1

Let's filter out questions with `trust` in their variable labels.

```{r}
select_metadata_vars <- c("filename", "var_name_orig", "var_label_orig", "var_label_norm", 
                          "val_label_orig", "val_label_norm" , "val_numeric_orig", 
                          "val_order_alpha", "n_categories")

trust_metadata <- metadata_database  %>%
  filter (
    grepl( "trust", var_label_norm )
  )  %>%
  select ( all_of(select_metadata_vars)) %>%
  arrange ( var_label_norm, val_label_norm, filename )

trust_metadata %>%
  head(10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Scrolling through the variable names it looks like binary trust in institutions questions have in their label "trust in institutions" or "trust political parties" or "- trust" at the end. Let's take a closer look at those questions and see what response options they have.

Several variables from Eurobarometer 69.2 (ZA4744) ask more detailed questions about specific reasons for trusting or not trusting selected institutions. These do not fit together with the more general items.

### Iteration 2

```{r}
trust_metadata <- trust_metadata %>% 
  filter (
    grepl( "trust_in_institutions|trust_political_parties|_trust$", var_label_norm )
  ) %>%
  filter( ! (filename == "ZA4744_v5-0-0.sav" & var_name_orig %in% c("v336", "v347","v292", "v303",
                                                                    "v314", "v325", "v270", "v281")))
trust_metadata %>%
  arrange(filename, var_name_orig, val_order_alpha) %>%
  head(10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Check how many unique response values do the selected variables have.

```{r}
trust_metadata %>%
  arrange(val_label_norm) %>%
  select(val_label_norm, everything()) %>% 
  count(n_categories) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

These variables have 2, 3 or 4, or as many as 13 unique response options. Let's take a look at the one with 13 unique responses:

```{r}
trust_metadata %>%
  filter(n_categories == 13) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

The question asks:
*Who do you trust the most to fight effectively in [country] against the European Union and its budget being defrauded?*
(see documentation in the [GESIS documentation](https://search.gesis.org/research_data/ZA3938)).

This refers to trust with regard to a specific activity, so does not fit into our concept definition. Let's see what we are left with after excluding it.

### Iteration 3

```{r}
trust_metadata <- trust_metadata %>%
  filter( ! (filename == "ZA3938_v1-0-1.sav" & var_name_orig == "v511"))

trust_metadata %>%
  count(val_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

* `dk` stands for "don't know",
* `inap*` stands for "inapplicable" for various reasons, which are maybe worth exploring,
* `mentioned` and `not mentioned` suggest questions where the respondent is asked to list all objects to which some condition applies, e.g. "here is a list of institutions. which of those would you say you trust?". These questions require special attention, because "no" is blended with "no answer / don't know". Plus, it's a very different type of question for the respondent,
* `na` and `lt_na_gt` are missing values,  
* `tend not to trust` is a negative answer to the trust question,
* `tend to trust` is a positive answer to the trust question,
* `you generally do not trust stories published on online social networks` is a mystery.

Let's start with the mystery.

```{r}
trust_metadata %>% 
  filter( grepl("you_generally_do_not_trust", val_label_norm)) %>%
  select(-var_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

That's likely not what we are interested in, but still worth checking with the [GESIS documentation](https://search.gesis.org/research_data/ZA6861).

The question reads "QD5 When you see or read a story published on online social networks, what makes you consider the story trustworthy?" and has several "mentioned/not mentioned" items. All those items are listed as a separate variables, and one of them ends with "trust", which is why we filtered it out. We don't want it and will remove it.

We refine the variable filter.

### Iteration 4

```{r}
trust_metadata <- trust_metadata %>% 
  filter( ! (filename == "ZA6861_v1-2-0.sav" & var_name_orig == "qd5.6"))

trust_metadata %>%
  count(val_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Now let's at the questions that have the `mentioned` / `not mentioned` response options. It's not what we want.

```{r}
exclusions <- trust_metadata %>%
  filter(val_label_norm %in% c("mentioned", "not_mentioned")) %>%
  print() %>%
  count(filename, var_name_orig)
```

### Iteration 5

```{r}
trust_metadata <- trust_metadata %>%
  anti_join(exclusions)

trust_metadata %>%
  count(val_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

## Standardizing value labels across questions

Now we need to create a table with all value labels, as well as topics / variable types.

```{r}
val_labels_trust <- trust_metadata %>%
  count(val_label_norm)

trust_vocabulary <- tibble::tibble (
  # maybe we can use a generic controlled vocabulary, for example
  # Library of Congress
  topic_1 = 'trust institutions',
  # And if we find them, we can add GESIS or TNS/Kantar keywords here
  topic_2 = 'trust, binary',
  val_label_norm = val_labels_trust %>% pull(val_label_norm), 
  level = 3 # missingness should be harmonized in character form 
)

trust_vocabulary %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

We add the harmonized values and labels, as well as a flag indicating whether the value is substantive of one of missing value codes, to create the `trust_values_table`.

```{r}
trust_values_table <- trust_vocabulary %>%
  mutate(character_value = case_when(
      # and create a surely harmonized character representation 
      grepl("dk|inap|na", val_label_norm) ~ NA_character_, 
      substr(val_label_norm, 1,7) == "tend_to" ~ "tend_to_trust",
      substr(val_label_norm, 1,11) == "tend_not_to" ~ "tend_not_to_trust",
    TRUE ~ "ERROR"), 
    numeric_value = case_when ( 
      character_value == "tend_to_trust" ~ 1,
      character_value == "tend_not_to_trust" ~ 0,
      TRUE ~ NA_real_),
    missing = case_when (
      # it is useful for faster filtering of missingness
      # and true value labels  
      is.na(numeric_value) ~ TRUE, 
      TRUE ~ FALSE)
  ) %>%
  arrange(numeric_value)

trust_values_table %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

## Standardizing variable names

The `trust_variable_table` will include original and harmonized variable names, as well as keywords and topics to match with the `trust_values_table`.

```{r}
trust_variable_table <- trust_metadata %>%
  filter ( val_label_norm %in% trust_values_table$val_label_norm ) %>%
  filter ( ! grepl("_recoded", var_label_norm ) ) %>%
  mutate ( institution = gsub("trust_in_institutions_|in_|_trust|trust_|[0-9]_|^a_|^b_|^q16c_|the_", "",
                              var_label_norm ),
           institution = gsub("charitable_org", "charities", institution ),
           institution = gsub("europ_court_of_auditors", "eu_court_of_auditors", institution ),
           institution = gsub("eur_court_of_auditors", "eu_court_of_auditors", institution ),
           institution = gsub("europ_court_of_justice", "eu_court_of_justice", institution ),
           institution = gsub("european_court_of_justice", "eu_court_of_justice", institution ),
           institution = gsub("justice_legal_system", "justice", institution ),
           institution = gsub("justice_nat_legal_system", "justice", institution ),
           institution = gsub("nat_parliament", "national_parliament", institution ),
           institution = gsub("nat_government", "national_government", institution ),
           institution = gsub("non_govmnt_org|non_govnmt_org", "ngo", institution ),
           institution = gsub("polit_parties", "political_parties", institution ),
           institution = gsub("reg_loc_public_authorities", "reg_loc_authorities", institution ),
           institution = gsub("reg_local_authorities", "reg_loc_authorities", institution ),
           institution = gsub("reg_local_public_authorities", "reg_loc_authorities", institution ),
           institution = gsub("rg_lc_authorities", "reg_loc_authorities", institution ),
           institution = gsub("written_press", "press", institution ),
           institution = gsub("econ_and_social_committee", "econ_and_soc_committee", institution ),
           institution = gsub("economic_and_soc_committee", "econ_and_soc_committee", institution )) %>%
  mutate ( geo_qualifier = case_when(
    grepl("_tcc", institution) ~ "tcc",
    TRUE ~ NA_character_),
    institution = gsub("_tcc", "", institution)
  ) %>%
  mutate ( var_name_suggested = paste0("trust_",
                                    institution, "_",
                                    geo_qualifier) ) %>%
  mutate ( var_name_suggested = gsub("_NA", "", var_name_suggested),
           topic_1 = "trust institutions",
           topic_2 = "trust, binary") %>%
  select ( -contains("label"),
           -val_numeric_orig, -val_order_alpha, -n_categories) %>%
  rename ( keyword_1 = institution ) %>%
  distinct_all() %>%
  mutate(var_name_suggested = ifelse(var_name_suggested == "trust_political_partiess_tcc",
                                    "trust_political_parties_tcc", var_name_suggested)) %>%
  arrange(var_name_suggested)

trust_variable_table %>% 
  head(10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )

```

How many different harmonized variables do we get? 54, each in between 1 and 42 Eurobarometer editions.

```{r}
trust_variable_table %>%
  count(var_name_suggested) %>%
  # arrange(desc(n)) %>%
  # head(10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

### 2020-06-18 data-raw/create_trust_tables.Rmd
renaming labelling functions

--- 2020-06-17 data-raw/trust_tables.Rmd
@@
----
-title: "Creating Variable and Values Tables"
-output: rmarkdown::html_vignette
-vignette: >
-  %\VignetteIndexEntry{create_tables}
-  %\VignetteEngine{knitr::rmarkdown}
-  %\VignetteEncoding{UTF-8}
----
-
-```{r, include = FALSE}
-knitr::opts_chunk$set(
-  message = FALSE,
-  warning = FALSE, 
-  collapse = TRUE,
-  comment = "#>"
-)
-options(width = 160)
-```
-
-```{r setup}
 library(eurobarometer)
@@
 #  )
-```
 
-## Procedure
-
-The procedure involves three steps:
-
-1. Concept definition: Trust in institutions: general trust without speficying the concrete domains of activity where the trust applies (these specific questions would require special treatment).
-
-2. Selection of survey questions that correspond to the concept by filtering on variable and value labels.
-
-3. Standardizing value labels across questions.
-
-4. Standardizing variable names to have the same name for all variables that correspond to the same question.
-
-## Metadata
-
-`metadata_database` is a table with variable metadata created from Eurobarometer SPSS files. It has the following columns:
-
-`var_name_orig`: variable name in the original dataset  
-`class_orig`: column class in the original dataset   
-`var_label_orig`: variable label in the original dataset  
-`var_label_norm`: normalized variable label  
-`var_name_suggested`: suggested variable label  
-`factor_levels`: vector of value labels as a list (they have no fixed length)  
-`n_categories`: total number of response options `= length(factor_levels)`   
-`class_suggested`: suggested conversion to an R class, it should correspond later to a conversion function, so that the researcher can just simply approve and get a correct R representation.  
-`filename`: original name of file as obtained from GESIS; it contains a disambigous version information, too.
-`val_numeric_orig`: numeric value code (if available, empty otherwise)   
-`val_label_orig`: value label in the original dataset corresponding to a response option
-`val_order_alpha`: alphabetical number (position) of response option in the set of value labels, aftering sorting with `sort()` (we use alphabetical order instead of `levels()` because levels may not be the same in different survey files, and we revert to a basic but disambigous sorting)
-`val_order_length`:  
-`val_label_norm`: value label normalized with `label_normalize()`
-
-```{r}
 metadata_database <-  readRDS( 
@@
   'eb_metadata_database_large.rds'))
-```
 
-## Selection of variables
-
-This is an iterative process.
-
-### Iteration 1
-
-Let's filter out questions with `trust` in their variable labels.
-
-```{r}
 select_metadata_vars <- c("filename", "var_name_orig", "var_label_orig", "var_label_norm", 
@@
                   font_size = 10 )
-```
 
-Scrolling through the variable names it looks like binary trust in institutions questions have in their label "trust in institutions" or "trust political parties" or "- trust" at the end.
-Let's take a closer look at those questions and see what response options they have.
-
-Several variables from Eurobarometer 69.2 (ZA4744) ask more detailed questions about specific reasons for trusting or not trusting selected institutions.
-These do not fit together with the more general items.
-
-### Iteration 2
-
-```{r}
 trust_metadata <- trust_metadata %>% 
@@
                   font_size = 10 )
-```
 
-Check how many unique response values do the selected variables have.
-
-```{r}
 trust_metadata %>%
@@
                   font_size = 10 )
-```
 
-These variables have 2, 3 or 4, or as many as 13 unique response options. Let's take a look at the one with 13 unique responses:
-
-```{r}
 trust_metadata %>%
@@
                   font_size = 10 )
-```
 
-The question asks:
-*Who do you trust the most to fight effectively in [country] against the European Union and its budget being defrauded?*
-(see documentation in the [GESIS documentation](https://search.gesis.org/research_data/ZA3938)).
-
-This refers to trust with regard to a specific activity, so does not fit into our concept definition. Let's see what we are left with after excluding it.
-
-### Iteration 3
-
-```{r}
 trust_metadata <- trust_metadata %>%
@@
                   font_size = 10 )
-```
 
-* `dk` stands for "don't know",
-* `inap*` stands for "inapplicable" for various reasons, which are maybe worth exploring,
-* `mentioned` and `not mentioned` suggest questions where the respondent is asked to list all objects to which some condition applies, e.g.
-"here is a list of institutions.
-which of those would you say you trust?".
-These questions require special attention, because "no" is blended with "no answer / don't know".
-Plus, it's a very different type of question for the respondent,
-* `na` and `lt_na_gt` are missing values,  
-* `tend not to trust` is a negative answer to the trust question,
-* `tend to trust` is a positive answer to the trust question,
-* `you generally do not trust stories published on online social networks` is a mystery.
-
-Let's start with the mystery.
-
-```{r}
 trust_metadata %>% 
@@
                   font_size = 10 )
-```
 
-That's likely not what we are interested in, but still worth checking with the [GESIS documentation](https://search.gesis.org/research_data/ZA6861).
-
-The question reads "QD5 When you see or read a story published on online social networks, what makes you consider the story trustworthy?" and has several "mentioned/not mentioned" items.
-All those items are listed as a separate variables, and one of them ends with "trust", which is why we filtered it out.
-We don't want it and will remove it.
-
-We refine the variable filter.
-
-### Iteration 4
-
-```{r}
 trust_metadata <- trust_metadata %>% 
@@
                   fixed_thead = T,
-                  font_size = 10 )
-```
 
-Now let's at the questions that have the `mentioned` / `not mentioned` response options. It's not what we want.
-
-```{r}
 exclusions <- trust_metadata %>%
@@
   count(filename, var_name_orig)
-```
 
-### Iteration 5
-
-```{r}
 trust_metadata <- trust_metadata %>%
@@
                   font_size = 10 )
-```
 
-## Standardizing value labels across questions
-
-Now we need to create a table with all value labels, as well as topics / variable types.
-
-```{r}
 val_labels_trust <- trust_metadata %>%
@@
 
-trust_vocabulary %>%
-  kable %>%
-  kable_styling(bootstrap_options =
-                  c("striped", "hover", "condensed"),
-                  fixed_thead = T,
-                  font_size = 10 )
-```
-
-We add the harmonized values and labels, as well as a flag indicating whether the value is substantive of one of missing value codes, to create the `trust_values_table`.
-
-```{r}
 trust_values_table <- trust_vocabulary %>%
@@
                   font_size = 10 )
-```
 
-## Standardizing variable names
-
-The `trust_variable_table` will include original and harmonized variable names, as well as keywords and topics to match with the `trust_values_table`.
-
-```{r}
 trust_variable_table <- trust_metadata %>%
@@
 trust_variable_table %>%
-  head(10) %>%
-  kable %>%
-  kable_styling(bootstrap_options =
-                  c("striped", "hover", "condensed"),
-                  fixed_thead = T,
-                  font_size = 10 )
-
-```
-
-How many different harmonized variables do we get? 54, each in between 1 and 42 Eurobarometer editions.
-
-```{r}
-trust_variable_table %>%
   count(var_name_suggested) %>%
@@
                   font_size = 10 )
-```

### 2020-06-18 R/data-eb_sample.R

@@ -1,0 +1,22 @@
+#' Vocabulary To Rename Categorical Variables With Two Valid Categories
+#'
+#' Later the filename may be omitted.
+#' The \code{canonical_name} is suggested for programmatic use in R files.
+#' The \code{topic_1}, \code{topic_2} keywords should be later replaced with
+#' consistent keywords from various thesauri, if necessary. Or they may be
+#' omitted.
+#'
+#' The treatment of declined answers in \code{level} should probably be
+#' excluded later, but in a harmonized way.
+#'
+#' @format A data frame with 6 rows in 9 variables:
+#' \describe{
+#'   \item{topic_1}{A temporary first level keyword.}
+#'   \item{topic_2}{A temporary second level keyword.}
+#'   \item{label_normalized}{The variable label after normalization, as imported from the GESIS SPSS file.}
+#'   \item{level}{The number of levels excluding decline.}
+#'   \item{character_value}{The preferred variable label as.character}
+#'   \item{numeric_value}{A suggested numerical coding, when applicable.}
+#'   \item{missing}{A logical variable to indicate if the label in the row represents some level of missingess.}
+#'   }
+"data-eb_sample"

### 2020-06-18 R/eurobarometer.R

@@ -1,11 +1,11 @@
 #' eurobarometer: A package to create tidy Eurobarometer microdata files.
 #'
-#' @section naming functions:
+#' @section labelling functions:
 #' The naming functions make the GESIS SPSS files usable in a programmatic
 #' context.
 #' \code{\link{label_normalize}} removes regex symbols, whitespace, and
 #' basic inconsistencies in abbreviations.
-#' \code{\link{canonical_names}} harmonizes the names across various GESIS
+#' \code{\link{label_suggest}} harmonizes the names across various GESIS
 #' files, so that they can be joined into panels across time.
 #'
 #' @section metadata functions:

### 2020-06-18 R/gesis_metadata_create.R

@@ -35,8 +35,8 @@ gesis_metadata_create <- function ( dat ) {
 
   var_label_orig <- vapply ( dat, labelled::var_label, character(1) )
   var_label_norm <- label_normalize(x = var_label_orig )
-  var_label_suggested <- canonical_names(var_label_norm,
-                                         names(dat))
+  var_label_suggested <- label_suggest( var_label_norm,
+                                        names(dat) )
 
   ## Creating the basic metadata ----
   metadata <- tibble::tibble (

### 2020-06-18 R/gesis_vocabulary_create.R

@@ -15,7 +15,7 @@
 #' @importFrom dplyr bind_rows mutate arrange
 #' @importFrom labelled var_label
 #' @importFrom tibble tibble
-#' @seealso canonical_name_create
+#' @seealso label_suggest
 #' @return The vocabulary of the survey in a data frame. The return
 #' data frame has 6 columns.
 #' @examples

### 2020-06-18 R/label_normalize.R

@@ -1,38 +1,33 @@
-#' Create Normalized Variable Names for GESIS Columns
+#' Create Normalized Variable & Value Labels
 #'
-#' @param x A vector of the GESIS variable names
-#' @importFrom stringr str_sub
+#' Remove relative reference to question blocks, special characters
+#' used in regular expressions, whitespace and bring the variable and
+#' value labels to snake_case_format.
+#'
+#' @param x A vector of the (GESIS) variable or value labels
+#' @importFrom stringr str_sub str_trim
 #' @family naming functions
+#' @return A character vector with the normalized labels
 #' @examples
-#' label_normalize ( c("UPPER CASE VAR", "VAR NAME WITH % SYMBOL") )
+#' label_normalize ( c(
+#' "QA17_1 EUROPEAN PARLIAMENT - TRUST",
+#' "QA13_1 TRUST IN INSTITUTIONS: NAT GOVERNMENT" )
+#'  )
 #' @export
 
 label_normalize <- function(x) {
 
-  y <- trimws(tolower(as.character(x)))
+  ## unit tests for this function are in
+  ## tests/testthat/test-label_normalize.R
+  ## please add test after correcting for unexpected results
 
+  y <- trimws(tolower(as.character(x)))
 
   ##do the abbreviations first that may have a . sign
   y <- gsub( '\\ss\\.a\\.', '_sa', y)
 
-  ##remove special regex characters
-  y <- gsub( '\\&', '_and_', y)
-  y <- gsub( '\\+', '_plus_', y)
-  y <- gsub( '\\%', '_pct_', y)
-  y <- gsub( '<', '_lt_', y)
-  y <- gsub( '>', '_gt_', y)
-  y <- gsub('\\.|-|\\:|\\;|\\/|\\(|\\)|\\!', '_', y)
-
-  ##remove space(s) and some other characters
-  y <- gsub('\\s\\s', '\\s', y)
-  y <- gsub("\\'", "",  y)
-
-  y <- ifelse ( test = stringr::str_sub ( y, -1, -1 ) == '_',
-                yes  = stringr::str_sub ( y,  1, -2 ),
-                no   = y  )
-
-
-
+  ## remove prefix of question block numbers
+  y <- gsub( '^q[abcde]\\d{1,2}_\\d{1,2}', '', y )  # remove QA117_1
[…]

### 2020-06-18 R/label_suggest.R

@@ -1,27 +1,30 @@
-#' Create Canonical Names for GESIS Columns
+#' Create Suggested Names for GESIS Columns
 #'
-#' Create canonical names that do not vary across several SPSS files
-#' from different years.
-#'
-#' This function may get a new name and some new functionality.
+#' Create canonical variable names (labels) that do not vary across
+#' several SPSS files from different years.
 #'
 #' @param var_names A character string of variable labels.
 #' @param var_name_orig A character string of variable names.
 #' @importFrom stringr str_sub
-#' @family naming functions
+#' @family labelling functions
 #' @examples
-#' canonical_names ( c("digital_object_identifier",
-#'                     "tns_unique_case_id") )
+#' label_suggest( c("digital_object_identifier",
+#'                  "tns_unique_case_id") )
 #' @export
 
-canonical_names <- function(var_names,
-                            var_name_orig) {
+label_suggest <- function(var_label_orig,
+                          var_name_orig) {
 
-  x <- trimws(tolower(as.character(var_names)))
+  x <- trimws(tolower(as.character(var_label_orig)))
   x <- label_normalize (x)
   x <- ifelse (var_name_orig == "caseid", "caseid", x)
   x <- ifelse (var_name_orig == "uniqid", "uniqid", x)
   x <- ifelse (var_name_orig == "doi", "doi", x)
+
+
+  y  <- gsub("wex_weight_extra_population_gt_15", "wex", y)
+  y  <- ifelse ( substr(y,1,3) =="w1_",
+                 'w1', y)
 
   ## Hard replacmenet ------------------------
   x <- ifelse ( grepl("unique_case_id", x),

### 2020-06-18 data-raw/create_eurobarometer_sample.R

@@ -5,7 +5,7 @@ ZA5479_sample <- dplyr::sample_n( ZA5479_sample, 2000 )
 ZA5479_sample <- ZA5479_sample[, c(1:8, 53, 104:114, 520:524, 551:622)]
 eb_sample <- ZA5479_sample
 
-usethis::use_data(eb_sample, internal = TRUE, overwrite = TRUE)
+usethis::use_data(eb_sample, internal = FALSE, overwrite = TRUE)
 
 
 sample <- eurobarometer:::eb_sample

### 2020-06-18 data-raw/create_trust_tables.R

@@ -1,0 +1,135 @@
+library(eurobarometer)
+library(dplyr)
+library(tibble)
+library(knitr)
+library(kableExtra)
+# The examples of this vignette can be found run with
+# source(
+#  file.path("not_included", "vignette_vocabulary_examples.R")
+#  )
+
+
+metadata_database <-  readRDS(
+  file.path( 'data-raw',
+  'eb_metadata_database_large.rds'))
+
+select_metadata_vars <- c("filename", "var_name_orig", "var_label_orig", "var_label_norm",
+                          "val_label_orig", "val_label_norm" , "val_numeric_orig",
+                          "val_order_alpha", "n_categories")
+
+trust_metadata <- metadata_database  %>%
+  filter (
+    grepl( "trust", var_label_norm )
+  )  %>%
+  select ( all_of(select_metadata_vars)) %>%
+  arrange ( var_label_norm, val_label_norm, filename )
+
+trust_metadata <- trust_metadata %>%
+  filter (
+    grepl( "trust_in_institutions|trust_political_parties|_trust$", var_label_norm )
+  ) %>%
+  filter( ! (filename == "ZA4744_v5-0-0.sav" & var_name_orig %in% c("v336", "v347","v292", "v303",
+                                                                    "v314", "v325", "v270", "v281")))
+trust_metadata %>%
+  arrange(filename, var_name_orig, val_order_alpha) %>%
+  head(10) %>%
+  kable %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 )
+
+
+trust_metadata <- trust_metadata %>%
+  filter( ! (filename == "ZA3938_v1-0-1.sav" &
+               var_name_orig == "v511"))
+
+
+trust_metadata <- trust_metadata %>%
+  filter( ! (filename == "ZA6861_v1-2-0.sav" &
+               var_name_orig == "qd5.6"))
+
+exclusions <- trust_metadata %>%
+  filter(val_label_norm %in% c("mentioned", "not_mentioned")) %>%
+  print() %>%
+  count(filename, var_name_orig)
+
+trust_metadata <- trust_metadata %>%
+  anti_join(exclusions,
+            by = c("filename", "var_name_orig"))
+
+
+val_labels_trust <- trust_metadata %>%
+  count(val_label_norm)
+
+trust_vocabulary <- tibble::tibble (
[…]

### 2020-06-18 tests/testthat/test-label_normalize.R

@@ -1,0 +1,45 @@
+
+
+### Test based on trust variable labels --------------------------------
+#paste0("'",
+#       paste( var_label_orig[c(1,122,455,899,1234,1989, 2400)], collapse = "', '"),
+#       "'")
+
+example_trust_labels <- c('QA17_1 EUROPEAN PARLIAMENT - TRUST',
+                          'Q26 COUNCIL OF MINISTERS - TRUST',
+                          'Q17 EUROPEAN COMMISSION - TRUST',
+                          'Q10 TRUST IN INSTITUTIONS: CIVIL SERVICE',
+                          'TRUST IN INSTITUTIONS: JUSTICE / LEGAL SYSTEM',
+                          'TRUST IN INSTITUTIONS: REG/LOC PUBLIC AUTHORITIES',
+                          'TRUST IN INSTITUTIONS: UNITED NATIONS')
+
+tested_trust_labels <- label_normalize ( example_trust_labels)
+
+test_that("correct vector is returned", {
+  expect_equal(length(tested_trust_labels), 7)
+  expect_equal(class(tested_trust_labels), "character")
+})
+tested_trust_labels
+
+test_that("qb preffix is removed", {
+  expect_equal(all(grepl( "qa1|q1", tested_trust_labels)), FALSE)
+})
+
+test_that("/ is removed", {
+  expect_equal(all(grepl( "\\/", tested_trust_labels)), FALSE)
+})
+
+test_that("/ is removed", {
+  expect_equal(all(grepl( "\\/", tested_trust_labels)), FALSE)
+})
+
+tested_trust_endings <- substr (x = tested_trust_labels,
+                                start = length(tested_trust_labels),
+                                stop = length(tested_trust_labels))
+
+test_that("no whitespace or clutter on end", {
+  expect_equal(all(grepl ( "\\s|_", tested_trust_endings )), FALSE)
+})
+
+
+

### 2020-06-18 R/create_panel.R
create_panel

@@ -1,4 +1,4 @@
-#' @title Create A Panel Skeleton With Unique ID
+#' @title Create A Skeleton Survey Panel With Unique ID
 #'
 #' @param survey_list A list of data frames containing surveys.
 #' @param id_vars A vector of ID variables to form a panel ID var,
@@ -16,14 +16,14 @@
 #' survey_list <- list ( 'ZA7576_sample' = ZA7576_sample,
 #'                       'ZA7562_sample' = ZA7562_sample,
 #'                       'ZA7489_sample' = ZA7489_sample)
-#' create_panel_id (survey_list, id_vars =c("uniqid", "doi") )
+#' panel_create (survey_list, id_vars =c("uniqid", "doi") )
 #' @export
 
 
-create_panel_id <- function (survey_list, id_vars = c("uniqid", "doi") ) {
+panel_create <- function (survey_list, id_vars = c("uniqid", "doi") ) {
 
   # in case of unexpected results, after fixing add a unique test
-  # to tests/testthat/test-create_panel_id.R
+  # to tests/testthat/test-panel_create.R
 
   create_id <- function (dat, id_vars = c("uniqid", "doi") ) {
 

### 2020-06-18 tests/testthat/test-panel_create.R

@@ -9,8 +9,8 @@ survey_list <- list ( 'ZA7576_sample' = ZA7576_sample,
                       'ZA7489_sample' = ZA7489_sample)
 
 
-panel_skeleton <- create_panel_id (survey_list,
-                                   id_vars =c("uniqid", "doi") )
+panel_skeleton <- panel_create (survey_list,
+                                id_vars =c("uniqid", "doi") )
 
 test_that("Correct data frame is created", {
   expect_equal(nrow ( panel_skeleton),

### 2020-06-18 R/panel_create.R
Progress in the workflow!

@@ -20,12 +20,13 @@
 #' @export
 
 
-panel_create <- function (survey_list, id_vars = c("uniqid", "doi") ) {
+panel_create <- function (survey_list,
+                          id_vars = c("uniqid", "doi") ) {
 
   # in case of unexpected results, after fixing add a unique test
   # to tests/testthat/test-panel_create.R
 
-  create_id <- function (dat, id_vars = c("uniqid", "doi") ) {
+  create_id <- function (dat ) {
 
     if ( ! all(id_vars %in% names(dat)) ) {
       stop ( "Not (all) of ",

### 2020-06-18 tests/testthat/test-panel_create.R

@@ -4,17 +4,29 @@ data (ZA7576_sample)
 data (ZA7562_sample)
 data (ZA7489_sample)
 
-survey_list <- list ( 'ZA7576_sample' = ZA7576_sample,
+survey_list_1 <- list ( 'ZA7576_sample' = ZA7576_sample,
                       'ZA7562_sample' = ZA7562_sample,
                       'ZA7489_sample' = ZA7489_sample)
 
+import_file_names <- c(
+  'ZA7576_sample','ZA7562_sample','ZA7489_sample'
+)
 
-panel_skeleton <- panel_create (survey_list,
+survey_list_2 <- read_surveys (
+  import_file_names, .f = 'read_example_file' )
+
+panel_skeleton_1 <- panel_create (survey_list_1,
                                 id_vars =c("uniqid", "doi") )
 
+panel_skeleton_2 <- panel_create (survey_list_2,
+                                id_vars =c("uniqid", "filename") )
+
+names ( panel_skeleton_2)
+
 test_that("Correct data frame is created", {
-  expect_equal(nrow ( panel_skeleton),
+  expect_equal(nrow ( panel_skeleton_1),
                nrow (ZA7576_sample) + nrow(ZA7562_sample) +
                  nrow(ZA7489_sample))
-  expect_equal(names(panel_skeleton),c("panel_id", "uniqid", "doi"))
+  expect_equal(names(panel_skeleton_1),c("panel_id", "uniqid", "doi"))
+  expect_equal(names(panel_skeleton_2),c("panel_id", "uniqid", "filename"))
 })

### 2020-06-19 README.Rmd
covr::codecov

@@
 "[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)"
-"[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/regions)](https://cran.r-project.org/package=eurobarometer)"
+"[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/eurobarometer)](https://cran.r-project.org/package=eurobarometer)"
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3825700.svg)](https://doi.org/10.5281/zenodo.3825700)

### 2020-06-19 README.Rmd
readme

@@
 
-The goal of eurobarometer is converting Eurobarometer microdata files, as stored by GESIS, into tidy R data frames and help common pre-processing problems.
+The goal of `eurobarometer` is converting Eurobarometer microdata files, as stored by GESIS, into tidy R data frames and help common pre-processing problems.
 

### 2020-06-19 README.Rmd
codecov token Readme cleaning

@@
 ```
+## Workflow
+
+Not all elements of the [Workflow](http://eurobarometer.danielantal.eu/articles/workflow.html) are implemented yet.
 
@@
 
-## Joining Serveral Files
-
-The vocabulary functions help to create data panels across Eurobarometer surveys taken in different times.
-
-There are two main issues that need to be resolved in such cases
-
-*  Make sure that the same questions are used, and they are consistently named (this problem is treated in the metadata level)
-
-* In case of structured questions or question blocks the answer options are coded exactly the same (for example “Fully agree”and “Totally agree” are the same.) This will be handled by the vocabulary functions.
-(Not yet implemented.)
-
 ## Joining With Eurostat & Other Data Tables

### 2020-06-19 R/utils.R
qb identification

@@ -34,37 +34,46 @@ class_suggest <- function(metadata) {
 
 #' Identify A Question Block
 #'
+#' Add a question block identifier in column \code{qb}.
 #' @param metadata A metadata file created by metadata_create within
 #' \code{\link{gesis_metadata_create}}
-#' @return A question block identifier.
+#' @return The metadata data frame with a new column \code{qb}.
 #' @importFrom magrittr %>%
-#' @importFrom dplyr case_when
+#' @importFrom dplyr case_when mutate
+#' @importFrom tidyselect all_of
 #'
 
 question_block_identify <- function (metadata) {
 
-  suggested <- as.character(metadata$var_name_suggested)
+  var_name_suggested <- as.character(metadata$var_name_suggested)
   var_label_orig <- as.character(metadata$var_label_orig)
   n_categories <- metadata$n_categories
 
-  qb <- vector ( mode = 'character', length = length(suggested))
-  orig_1 <- tolower(substr(var_label_orig, 1, 1))
-  orig_2 <- tolower(substr(var_label_orig, 1, 2))
+  qb <- vector ( mode = 'character', length = length(var_name_suggested))
 
- qb <- dplyr::case_when (
 
-    suggested %in% c("filename", "doi", "uniqid" ) ~  "id",
-    grepl( "gesis_archive", suggested)             ~  "id",
+  metadata_tmp <- metadata %>%
+    mutate (
+      orig_1 = tolower(substr(var_name_orig, 1, 1)),
+      orig_2 = tolower(substr(var_name_orig, 1, 2))
+      ) %>%
+    mutate (qb  = case_when (
+
+    var_name_suggested %in%
+      c("filename", "doi", "uniqid" )              ~  "id",
+    grepl( "gesis_archive", var_name_suggested)    ~  "id",
     orig_1 == "p"                                  ~  "metadata",
-    var_label_orig %in%
-      c("d7", "d8", "d25", "d60")                  ~  "demography",
+    var_name_orig == "isocntry"                    ~  "metadata",
+    var_name_orig %in%
+      c("d7", "d8", "d25", "d60")                  ~  "socio-demography",
     orig_1 == "w"                                  ~  "weights",
-    grepl ( "trust_in|_trust", suggested) &
+    grepl ( "trust_in|_trust", var_name_suggested) &
       n_categories  == 4                           ~  "trust",
 
 
[…]

### 2020-06-19 README.Rmd
README update

@@
 
+[![Travis build status](https://travis-ci.com/antaldaniel/eurobarometer.svg?branch=master)](https://travis-ci.com/antaldaniel/eurobarometer)
 <!-- badges: end -->
@@
 
-* Currently `code_nuts1()` and `code_nuts2()` is implemented but this will be replaced with a more comprehensive NUTS0-NUTS3 level helper function to make the regional coding consistent.
+* Currently `code_nuts1()` and `code_nuts2()` is implemented but this will be replaced with a more comprehensive `NUTS0`-`NUTS3` level helper function to make the regional coding consistent.
 

### 2020-06-19 README.Rmd
news update

@@
 <!-- badges: start -->
+[![Travis build status](https://travis-ci.com/antaldaniel/eurobarometer.svg?branch=master)](https://travis-ci.com/antaldaniel/eurobarometer)
 [![Codecov test coverage](https://codecov.io/gh/antaldaniel/eurobarometer/branch/master/graph/badge.svg)](https://codecov.io/gh/antaldaniel/eurobarometer?branch=master)
@@
 [![Follow author](https://img.shields.io/twitter/follow/antaldaniel.svg?style=social)](https://twitter.com/intent/follow?screen_name=antaldaniel)
+[![Follow contributor](https://img.shields.io/twitter/follow/martakolcz.svg?style=social)](https://twitter.com/martakolcz?lang=en)
 
-[![Travis build status](https://travis-ci.com/antaldaniel/eurobarometer.svg?branch=master)](https://travis-ci.com/antaldaniel/eurobarometer)
 <!-- badges: end -->

### 2020-06-19 README.Rmd
README update

--- 2020-06-19 README.Rmd
@@
 [![Follow author](https://img.shields.io/twitter/follow/antaldaniel.svg?style=social)](https://twitter.com/intent/follow?screen_name=antaldaniel)
-
-[![Travis build status](https://travis-ci.com/antaldaniel/eurobarometer.svg?branch=master)](https://travis-ci.com/antaldaniel/eurobarometer)
+[![Follow contributor](https://img.shields.io/twitter/follow/martakolcz.svg?style=social)](https://twitter.com/martakolcz?lang=en)
 <!-- badges: end -->

### 2020-06-19 README.Rmd
Changes on travis-ci, new documentation

--- 2020-06-19 README.Rmd
@@
 "[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/eurobarometer)](https://cran.r-project.org/package=eurobarometer)"
-[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3825700.svg)](https://doi.org/10.5281/zenodo.3825700)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3901724.svg)](https://doi.org/10.5281/zenodo.3901724)
 [![Follow author](https://img.shields.io/twitter/follow/antaldaniel.svg?style=social)](https://twitter.com/intent/follow?screen_name=antaldaniel)
-
-[![Travis build status](https://travis-ci.com/antaldaniel/eurobarometer.svg?branch=master)](https://travis-ci.com/antaldaniel/eurobarometer)
+[![Follow contributor](https://img.shields.io/twitter/follow/martakolcz.svg?style=social)](https://twitter.com/martakolcz?lang=en)
 <!-- badges: end -->

### 2020-06-20 R/id_create.R
id_create must make unique IDs

@@ -29,19 +29,32 @@ id_create <- function (dat,
            " present in <dat>.")
   }
 
+  first_id_var <- id_vars[1]
+  last_id_var <- id_vars[length(id_vars)]
+
   tmp <- dat %>%
-    select ( all_of ( id_vars ))
+    select ( all_of ( id_vars )) %>%
+    mutate_all ( as.character )
 
   if ( nrow(tmp) == 0 ) {
     stop ( "The id_vars are not present in the rows." )
   }
 
-  tmp %>%
+  tmp <- tmp %>%
     tidyr::unite ( col = 'panel_id',
-                   all_of(id_vars), remove=FALSE ) %>%
-    mutate ( panel_id = label_normalize(panel_id))
+                   !!first_id_var:!!last_id_var, remove=FALSE ) %>%
+    mutate ( panel_id = gsub("[^[:alnum:] ]", "_", panel_id)) %>%
+    select ( panel_id, everything())
 
+  unicity_test <- tmp %>% summarize_all ( n_unique )
+
+  if ( unicity_test$panel_id < nrow(tmp) ) {
+    #warning("The id_vars=c('", paste(id_vars, collapse = "', '"),
+    #        "') do not form unique IDs, row number is added as a prefix.")
+
+    tmp$panel_id <- paste0(1:nrow(tmp), "_", tmp$panel_id)
+  }
+
+  tmp
 }
-
-dat <- survey_list[[1]]
 

### 2020-06-20 tests/testthat/test-id_create.R

@@ -1,0 +1,22 @@
+
+test_id_create <- data.frame (
+  uniqid = c(1,2,3,4,5,5),
+  doi = rep("doi:10.4232/1.13393", 6))
+
+corrected <- id_create ( dat = test_id_create,
+                         id_vars = c("uniqid", "doi")
+                         )
+
+#test_that("warning is given", {
+#  expect_warning( id_create (test_id_create,
+#                             id_vars = c("uniqid", "doi")))
+#})
+
+test_that("correct ID is returned", {
+  expect_equal( id_create (test_id_create,
+                             id_vars = c("uniqid", "doi"))$panel_id,
+                paste0 ( 1:6, "_", c(1,2,3,4,5,5), "_", "doi_10_4232_1_13393") )
+})
+
+
+

### 2020-06-20 R/id_create.R
spacing

@@ -3,21 +3,16 @@
 #' @param dat A data frames containing one survey.
 #' @param id_vars A vector of ID variables to form a panel ID var,
 #' defaults to \code{c("uniqid", "doi")}.
-#' @importFrom dplyr select mutate summarize_all
+#' @importFrom dplyr select mutate summarize_all mutate_all
 #' @importFrom tidyr unite
 #' @importFrom magrittr %>%
 #' @importFrom tidyselect all_of everything
 #' @return A tibble with the number of original ID vars and their
 #' \code{panel_id} as their concatenation.
+#' @seealso panel_create
 #' @examples
 #' data (ZA7576_sample)
-#' data (ZA7562_sample)
-#' data (ZA7489_sample)
-#'
-#' survey_list <- list ( 'ZA7576_sample' = ZA7576_sample,
-#'                       'ZA7562_sample' = ZA7562_sample,
-#'                       'ZA7489_sample' = ZA7489_sample)
-#' panel_create (survey_list, id_vars =c("uniqid", "doi") )
+#' id_create (dat=ZA7576_sample, id_vars =c("uniqid", "doi") )
 #' @export
 
 id_create <- function (dat,

### 2020-06-20 R/id_create.R
documentation improvements

@@ -8,7 +8,7 @@
 #' @importFrom magrittr %>%
 #' @importFrom tidyselect all_of everything
 #' @return A tibble with the number of original ID vars and their
-#' \code{panel_id} as their concatenation.
+#' \code{panel_id} as their concatenation for a single survey.
 #' @family harmonization functions
 #' @examples
 #' data (ZA7576_sample)

### 2020-06-20 R/panel_create.R

@@ -4,7 +4,7 @@
 #' @param id_vars A vector of ID variables to form a panel ID var,
 #' defaults to \code{c("uniqid", "doi")}.
 #' @return A tibble with the number of original ID vars and their
-#' \code{panel_id} as their concatenation.
+#' \code{panel_id} as their concatenation for several surveys.
 #' @family harmonization functions
 #' @examples
 #' data (ZA7576_sample)

### 2020-06-20 R/harmonize_qb_vars.R
distinct_at depedency + documentation

@@ -15,7 +15,7 @@
 #' a column in \code{metadata}.
 #' @param conversion Defaults to \code{class_suggested}. Must be
 #' a column in \code{metadata}.
-#' @importFrom dplyr select filter bind_cols
+#' @importFrom dplyr select filter bind_cols distinct_at
 #' @importFrom purrr set_names
 #' @importFrom magrittr %>%
 #' @importFrom tidyselect all_of

### 2020-06-20 R/harmonize_qb_vars.R
non standard evaluation

@@ -42,7 +42,8 @@ harmonize_qb_vars <- function ( survey_list,
                                 conversion = "class_suggested" ) {
 
   id_vars_input <- id_vars
-  id_vars <- NULL
+  id_vars <- qb <- survey_subset <- var_name_orig <- NULL
+  panel_id <- NULL
 
   if ( ! "list" %in% class(survey_list) ) {
     if ( "data.frame" %in% class(survey_list) ) {

### 2020-06-21 R/data-metadata_filter_example.R
Trying to make the vignette build with external data

@@ -1,0 +1,19 @@
+#' Example file for the Creating Variable and Values Tables vignette
+#'
+#' @format A data frame with 4743 observations and 14 variables:
+#' \describe{
+#'   \item{var_label_orig}{The original variable label in SPSS}
+#'   \item{class_orig}{The original variable class after importing with\code{\link[haven]{read_spss}}}
+#'   \item{var_name_orig}{The original variable name in SPSS}
+#'   \item{var_label_norm}{A normalized version of the variable labels}
+#'   \item{var_name_suggested}{A partly canonized variable name.}
+#'   \item{factor_levels}{A list of factor levels, i.e. value labels in SPSS}
+#'   \item{class_suggested}{A suggested class conversion.}
+#'   \item{filename}{The original file name}
+#'   \item{var_numeric_orig}{The original numeric value}
+#'   \item{val_label_orig}{The original value label}
+#'   \item{var_order_alpha}{Sorted order of the value label}
+#'   \item{var_order_length}{Total number of value labels of the variable}
+#'   \item{val_label_norm}{A normalized version of the value labels}
+#'   }
+"metadata_filter_example"

### 2020-06-21 data-raw/create_metadata_filter_example.R

@@ -1,0 +1,12 @@
+metadata_rel_path_from_root <- find_root_file(
+  "data-raw", "eb_metadata_database_large.rds",
+  criterion = has_file("DESCRIPTION"))
+
+metadata_database <- readRDS( metadata_rel_path_from_root )
+
+metadata_filter_example <- metadata_database  %>%
+  filter (
+    grepl( "trust|concert|democracy", var_label_norm )
+  )
+
+usethis::use_data ( metadata_filter_example, overwrite=TRUE)

### 2020-06-21 README.Rmd
spelling

@@
 
-At this moment you cannot install yest the packgae from [CRAN](https://CRAN.R-project.org), but there is a development version on [GitHub](https://github.com/):
+At this moment you cannot install yet the package from [CRAN](https://CRAN.R-project.org), but there is a development version on [GitHub](https://github.com/):
 

### 2020-06-22 R/data-za5913_sample.R
change of data files

@@ -1,0 +1,54 @@
+#' ZA5913 Eurobarometer Sample
+#'
+#' ZA5913: Eurobarometer 81.2 (March 2014): Europeans in 2014,
+#' Financial and Economic Crisis, European Citizenship, and
+#' European Values
+#'
+#' A reduced part of a raw GESIS file to demonstrate changes in
+#' variable labels and variable names.
+#'
+#' @format A data frame with 4123 rows in 36 variables:
+#' \describe{
+#'   \item{doi}{DIGITAL OBJECT IDENTIFIER}
+#'   \item{version}{GESIS ARCHIVE VERSION & DATE}
+#'   \item{uniqid}{UNIQUE RESPONDENT ID (CASEID BY TNS COUNTRY CODE)}
+#'   \item{isocntry}{COUNTRY CODE - ISO 3166}
+#'   \item{p1}{DATE OF INTERVIEW}
+#'   \item{p3}{DURATION OF INTERVIEW (MINUTES)}
+#'   \item{p4}{N OF PERSONS PRESENT DURING INTERVIEW}
+#'   \item{p5}{RESPONDENT COOPERATION}
+#'   \item{nuts}{REGION - NUTS CODES}
+#'   \item{d7}{MARITAL STATUS}
+#'   \item{d8}{AGE EDUCATION}
+#'   \item{d25}{TYPE OF COMMUNITY}
+#'   \item{d60}{DIFFICULTIES PAYING BILLS - LAST YEAR}
+#'   \item{qa10_3}{EUROPEAN CENTRAL BANK - TRUST}
+#'   \item{qa10_2}{EUROPEAN COMMISSION - TRUST}
+#'   \item{qa10_1}{EUROPEAN PARLIAMENT - TRUST}
+#'   \item{qa7_4}{TRUST IN INSTITUTIONS: EUROPEAN UNION}
+#'   \item{qa7_2}{TRUST IN INSTITUTIONS: NATIONAL GOVERNMENT}
+#'   \item{qa7_3}{TRUST IN INSTITUTIONS: NATIONAL PARLIAMENT}
+#'   \item{qa7_1}{TRUST IN INSTITUTIONS: POLITICAL PARTIES}
+#'   \item{qa7_5}{TRUST IN INSTITUTIONS: REG/LOC AUTHORITIES}
+#'   \item{qd3_1}{IMPORTANT VALUES PERS: RULE OF LAW}
[…]

### 2020-06-22 R/data-za6863_sample.R

@@ -1,0 +1,60 @@
+#' ZA6863 Eurobarometer Sample
+#'
+#' A reduced part of a raw GESIS file to demonstrate changes in
+#' variable labels and variable names.
+#'
+#' @format A data frame with 5074 rows in 50 variables:
+#' \describe{
+#'   \item{doi}{DIGITAL OBJECT IDENTIFIER}
+#'   \item{version}{GESIS ARCHIVE VERSION & DATE}
+#'   \item{uniqid}{UNIQUE RESPONDENT ID (CASEID BY TNS COUNTRY CODE)}
+#'   \item{isocntry}{COUNTRY CODE - ISO 3166}
+#'   \item{p1}{DATE OF INTERVIEW}
+#'   \item{p2}{TIME OF INTERVIEW}
+#'   \item{p3}{DURATION OF INTERVIEW}
+#'   \item{p4}{N OF PERSONS PRESENT DURING INTERVIEW}
+#'   \item{p5}{RESPONDENT COOPERATION}
+#'   \item{nuts}{REGION - NUTS CODES}
+#'   \item{d7}{MARITAL STATUS}
+#'   \item{d8}{AGE EDUCATION}
+#'   \item{d25}{TYPE OF COMMUNITY}
+#'   \item{d60}{DIFFICULTIES PAYING BILLS - LAST YEAR}
+#'   \item{qa14_3}{EUROPEAN CENTRAL BANK - TRUST}
+#'   \item{qa14_2}{EUROPEAN COMMISSION - TRUST}
+#'   \item{qa14_1}{EUROPEAN PARLIAMENT - TRUST}
+#'   \item{qa8a_3}{TRUST IN INSTITUTIONS: ARMY}
+#'   \item{qa8a_9}{TRUST IN INSTITUTIONS: EUROPEAN UNION}
+#'   \item{qa8b_2}{TRUST IN INSTITUTIONS: EUROPEAN UNION (TCC)}
+#'   \item{qa8a_1}{TRUST IN INSTITUTIONS: JUSTICE / LEGAL SYSTEM}
+#'   \item{qa8a_7}{TRUST IN INSTITUTIONS: NATIONAL GOVERNMENT}
+#'   \item{qa8a_8}{TRUST IN INSTITUTIONS: NATIONAL PARLIAMENT}
+#'   \item{qa8a_2}{TRUST IN INSTITUTIONS: POLICE}
[…]

### 2020-06-22 R/data-za7489_sample.R

@@ -1,145 +1,0 @@
-#' ZA7489 Eurobarometer Sample
-#'
-#' A reduced part of a raw GESIS file to demonstrate changes in
-#' variable labels and variable names.
-#'
-#' @format A data frame with 6613 rows in 133 variables:
-#' \describe{
-#'   \item{doi}{DIGITAL OBJECT IDENTIFIER}
-#'   \item{version}{GESIS ARCHIVE VERSION & DATE}
-#'   \item{uniqid}{UNIQUE RESPONDENT ID (CASEID BY KANTAR COUNTRY CODE)}
-#'   \item{isocntry}{COUNTRY CODE - ISO 3166}
-#'   \item{p1}{DATE OF INTERVIEW}
-#'   \item{p2}{TIME OF INTERVIEW}
-#'   \item{p3}{DURATION OF INTERVIEW}
-#'   \item{p4}{N OF PERSONS PRESENT DURING INTERVIEW}
-#'   \item{p5}{RESPONDENT COOPERATION}
-#'   \item{d7}{MARITAL STATUS}
-#'   \item{d8}{AGE EDUCATION}
-#'   \item{d25}{TYPE OF COMMUNITY}
-#'   \item{d60}{DIFFICULTIES PAYING BILLS - LAST YEAR}
-#'   \item{qa13_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa13_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa13_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa13_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa13_3}{EUROPEAN CENTRAL BANK - TRUST}
-#'   \item{qa13_3}{EUROPEAN CENTRAL BANK - TRUST}
-#'   \item{qa13_3}{EUROPEAN CENTRAL BANK - TRUST}
-#'   \item{qa13_3}{EUROPEAN CENTRAL BANK - TRUST}
-#'   \item{qa13_2}{EUROPEAN COMMISSION - TRUST}
-#'   \item{qa13_2}{EUROPEAN COMMISSION - TRUST}
-#'   \item{qa13_2}{EUROPEAN COMMISSION - TRUST}
[…]

### 2020-06-22 R/data-za7576_sample.R

@@ -1,10 +1,15 @@
 #' ZA7576 Eurobarometer Sample
 #'
+#' ZA7576: Eurobarometer 91.5 (June-July 2019):
+#' Standard Eurobarometer, European Parliament
+#' Post-election survey 2019, and EU citizens and
+#' development cooperation
+#'
 #' A reduced part of a raw GESIS file to demonstrate changes in
 #' variable labels and variable names.
 #'
 #'
-#' @format A data frame with 5547 rows in 105 variables:
+#' @format A data frame with 6579 rows in 53 variables:
 #' \describe{
 #'   \item{doi}{DIGITAL OBJECT IDENTIFIER}
 #'   \item{version}{GESIS ARCHIVE VERSION & DATE}
@@ -15,104 +20,52 @@
 #'   \item{p3}{DURATION OF INTERVIEW}
 #'   \item{p4}{N OF PERSONS PRESENT DURING INTERVIEW}
 #'   \item{p5}{RESPONDENT COOPERATION}
+#'   \item{nuts}{REGION - NUTS CODES}
 #'   \item{d7}{MARITAL STATUS}
 #'   \item{d8}{AGE EDUCATION}
 #'   \item{d25}{TYPE OF COMMUNITY}
 #'   \item{d60}{DIFFICULTIES PAYING BILLS - LAST YEAR}
 #'   \item{qa14_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa14_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa14_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa14_5}{COUNCIL OF THE EU - TRUST}
-#'   \item{qa14_3}{EUROPEAN CENTRAL BANK - TRUST}
-#'   \item{qa14_3}{EUROPEAN CENTRAL BANK - TRUST}
-#'   \item{qa14_3}{EUROPEAN CENTRAL BANK - TRUST}
 #'   \item{qa14_3}{EUROPEAN CENTRAL BANK - TRUST}
 #'   \item{qa14_2}{EUROPEAN COMMISSION - TRUST}
[…]

### 2020-06-22 R/eurobarometer.R

@@ -27,6 +27,13 @@
 #' metadata analyais.
 #' \code{\link{harmonize_qb_vars}} is harmonizing the names and classes of
 #' questions in a question block.
+#'
+#' @section sample data files:
+#' \code{\link{ZA7489_sample}}
+#'
+#' \code{\link{ZA7572_sample}}
+#'
+#' \code{\link{ZA7576_sample}}
 #' @docType package
 #' @name eurobarometer
 NULL

### 2020-06-22 data-raw/create_small_eb_panel.R

@@ -4,12 +4,18 @@ source (file.path("data-raw", "create_trust_tables.R"))
 source(file.path("not_included", "daniel_env.R")) ##loads Daniel's local gesis directory
 
 selected_files <- sort ( unique ( trust_metadata$filename),
-                         decreasing = T)[1:3]
+                         decreasing = T)
+
+selected_files <- c("ZA5913_v2-0-0.sav",
+                    "ZA7576_v1-0-0.sav",
+                    "ZA6863_v1-0-0.sav")
+
 selected_trust_vars <- trust_metadata %>%
   filter ( filename %in% selected_files ) %>%
   select (
     all_of(c("filename", "var_name_orig", "var_label_norm"))
-    )
+    ) %>%
+  distinct_all()
 
 id_vars <- metadata_database %>%
   select (
@@ -53,7 +59,8 @@ weight_vars <- metadata_database %>%
   distinct_all()
 
 
-metadata_var_names_orig <- c("p1", "p2", "p3", "p4", "p5", "isocntry", "doi")
+metadata_var_names_orig <- c("p1", "p2", "p3", "p4", "p5",
+                             "isocntry", "doi", "v713", "nuts")
 
 metadata_vars <- metadata_database %>%
   select (
@@ -74,7 +81,7 @@ other_vars <- metadata_database %>%
   filter (
     filename %in% selected_files
   ) %>%
-  filter ( grepl( "voters_|political_discussion|vaccine_info_source",
+  filter ( grepl( "voters_|important_values_pers|vaccine_info_source",
                   var_name_suggested )) %>%
   distinct_all()
 
@@ -86,32 +93,36 @@ subsetting_vars <- id_vars %>%
   full_join ( weight_vars ) %>%
   select ( all_of(c("filename", "var_name_orig")))
 
+selected_files <- c("ZA5913_v2-0-0.sav",
+                    "ZA7576_v1-0-0.sav",
+                    "ZA6863_v1-0-0.sav")
+
 for (f in selected_files) { cat (f)}
 
 
-ZA7489_sample <- haven::read_spss( file.path(gesis_dir, "ZA7489_v1-0-0.sav"))
-ZA7562_sample <- haven::read_spss( file.path(gesis_dir, "ZA7562_v1-0-0.sav"))
+ZA5913_sample <- haven::read_spss( file.path(gesis_dir, "ZA5913_v2-0-0.sav"))
[…]

### 2020-06-24 R/convert_class.R
0.1.5 value label harmonization

@@ -51,6 +51,15 @@ convert_class <- function(dat, metadata,
     select ( tidyselect::all_of(character_vars) ) %>%
     mutate_all ( as.character )
 
+  labelled_vars <- class_conversion %>%
+    filter ( conversion == 'harmonized_labelled') %>%
+    select ( all_of("var")) %>%
+    unlist() %>% as.character()
+
+  labelled_df <- dat %>%
+    select ( tidyselect::all_of(labelled_vars) ) %>%
+    mutate_all ( harmonize_value_labels )
+
   dummy_vars <- class_conversion %>%
     filter ( conversion == 'dummy') %>%
     select ( all_of("var")) %>%
@@ -85,7 +94,7 @@ convert_class <- function(dat, metadata,
     mutate_all ( relabel_factors  )
 
   remerged_dat <- dplyr::bind_cols (
-    factor_df, character_df, numeric_df, dummy_df
+    factor_df, character_df, numeric_df, dummy_df, labelled_df
   )
 
   missing_from_return <- names(dat) [! names(dat) %in% names (remerged_dat)]

### 2020-06-24 R/data-label_harmonization_table.R

@@ -1,0 +1,14 @@
+#' Vocabulary To Rename Categorical Variables With Two Valid Categories
+#'
+#' The harmonization table of categorical value labels. Currently contains
+#' only valid binary values.  Missing labels are harmonized with the
+#' internal function \code{\link{harmonize_missing_values}}.
+#'
+#' @format A data frame with 84 rows in 3 variables:
+#' \describe{
+#'   \item{label_norm}{Value labels created by \code{\link{label_normalize}}}
+#'   \item{value_numeric}{Harmonized numeric values}
+#'   \item{label_harmonized}{Harmonized labels}
+#'   }
+#'   @seealso harmonize_value_labels label_normalize
+"label_harmonization_table"

### 2020-06-24 R/eurobarometer.R

@@ -27,6 +27,8 @@
 #' metadata analyais.
 #' \code{\link{harmonize_qb_vars}} is harmonizing the names and classes of
 #' questions in a question block.
+#' \code{\link{harmonize_missing_values}} is harmonizing the missing values
+#' in a labelled vector.
 #'
 #' @section sample data files:
 #' \code{\link{ZA5913_sample}} A small subset of the ZA5913:

### 2020-06-24 R/harmonize_qb_vars.R

@@ -46,11 +46,7 @@ harmonize_qb_vars <- function ( survey_list,
   . <- panel_id <- NULL
 
   if ( ! "list" %in% class(survey_list) ) {
-    if ( "data.frame" %in% class(survey_list) ) {
-      survey_list <- list ( dat = survey_list )
-    } else {
-      stop ( "'survey_list' must be a data frame or a list of data frames." )
-    }
+    survey_list <- to_survey_list(survey_list)
   }
 
   if ( ! var_name %in% names(metadata) ) {
@@ -90,8 +86,9 @@ harmonize_qb_vars <- function ( survey_list,
 
     if ( ncol(question_block) > 0 ) {
       question_block <- question_block %>%
-        purrr::set_names ( new_names ) %>%
-        convert_class (., subset_metadata,
+        purrr::set_names ( new_names )  %>%
+        convert_class (.,
+                       metadata = subset_metadata,
                        conversion = conversion )
 
       dplyr::bind_cols(skeleton, question_block)

### 2020-06-24 R/harmonize_value_labels.R

@@ -1,0 +1,104 @@
+#' Harmonize Value Labels
+#'
+#' Harmonize the numeric values and the value labels of variables.
+#' Currently works only with binary values.
+#'
+#' @importFrom labelled to_character labelled
+#' @importFrom dplyr case_when mutate left_join if_else distinct_all
+#' @importFrom tibble tibble
+#' @return A labelled vector containing harmonized numeric values and
+#' labels, and in \code{attr(v, num_orig)} the original numeric values,
+#' \code{attr(v, labels_orig)} the original labelling.
+#' @examples
+#' v <- labelled(c(3,4,4,3,8, 9),
+#'              c(YES = 3, NO = 4, wrong_label = 8, refused = 9)
+#'              )
+#' harmonize_value_labels(v)
+#' @export
+
+harmonize_value_labels <- function (labelled_var) {
+
+  data ( "label_harmonization_table", envir = environment())
+
+  if ("haven_labelled_spss" %in% class ( labelled_var )) {
+
+    new_labelled_var <- labelled_var
+    labelled_var <- labelled::labelled(
+      unclass(new_labelled_var),
+      labels = attr(new_labelled_var, "labels")
+    )
+  }
+
+  harmonized_1 <- tibble (
+    numeric = as.numeric(labelled_var),
+    label_orig = labelled::to_character(labelled_var),
+    label_norm = label_normalize(label_orig),
+  ) %>%
+    left_join ( label_harmonization_table, by = 'label_norm' ) %>%
+    mutate ( label_harmonized = if_else(
+      is.na(label_harmonized),
+      label_norm,
+      label_harmonized
+    )) %>%
+    mutate ( value_numeric = if_else(
+      is.na(value_numeric),
+      as.numeric(numeric),
+      as.numeric(value_numeric)
+    ))
+
+  valid_harmonized_values <- unique (
+    label_harmonization_table$label_harmonized
+  )
+
+  harmonized <- harmonized_1 %>%
+    left_join( harmonize_missing_values(labelled_var) %>%
+                 distinct_all(),
+               by = 'label_norm') %>%
+    mutate ( value_numeric = if_else (
+      ! label_harmonized %in% valid_harmonized_values,
+      8999,
+      value_numeric
+    )) %>%
+    mutate ( value_numeric = if_else (
+      label_harmonized == na_harmonized,
+      value_numeric,
+      na_numeric_value
+    )) %>%
+    mutate ( label_harmonized = if_else (
[…]

### 2020-06-24 R/label_normalize.R

@@ -78,6 +78,7 @@ label_normalize <- function(x) {
   y <- gsub( '\\s', '_', y)
   y <- gsub( '___', '_', y)
   y <- gsub( '__', '_', y)
+  y <- gsub( '\\,_', '_', y)
 
   # Shall we remove Central & Eastern European characters, Greek characters?
   #x <- gsub( 'á', 'a', x)
@@ -90,6 +91,10 @@ label_normalize <- function(x) {
 
   y <- ifelse ( test = stringr::str_sub ( y, 1,  1 ) == '_',
                 yes  = stringr::str_sub ( y, 2, -1 ),
+                no   = y  )
+
+  y <- ifelse ( test = stringr::str_sub(y, -3, -1) == " - ",
+                yes  = stringr::str_sub ( y, 1, -4 ),
                 no   = y  )
 
   stringr::str_trim(y, side = "both")

### 2020-06-24 R/read_surveys.R

@@ -6,8 +6,8 @@
 #' @param import_file_names A vector of file names to import.
 #' @param .f A function to import the surveys with.
 #' Defaults to \code{'read_example_file'}. For SPSS files,
-#' \code{'read_spss_survey'} is recommended, which is a well-parameterised
-#' version of \code{\link{[haven]{read_spss}}}.
+#' \code{'read_spss_survey'} is recommended, which is a
+#' well-parameterised version of \code{\link[haven]{read_spss}}.
 #' @return A list of the surveys.  Each element of the list is a data
 #' frame. The respective file names are added to each data frame as a
 #' constant column \code{filename}.

### 2020-06-24 R/utils.R

@@ -28,7 +28,7 @@ class_suggest <- function(metadata) {
       class_orig %in% c("numeric","character") ~ class_orig,
       class_orig == 'haven_labelled' & n_categories  == 1 ~ 'character',
       first_level == "mentioned" & second_level == "not mentioned" ~  'dummy',
-      TRUE ~ "factor" ))
+      TRUE ~ "harmonized_labelled" ))
 
   metadata$class_suggested <- suggestions$suggestion
   metadata
@@ -95,4 +95,46 @@ to_survey_list <- function (x) {
     }
   }
   x
+}
+
+#' Harmonize Missing Values
+#'
+#' @param x A vector of labelled or the character vector of
+#' value labels.
+#' @importFrom labelled to_character labelled
+#' @importFrom dplyr case_when mutate
+#' @importFrom tibble tibble
+#' @family harmonize functions
+#' @return A helper table to harmonize missing values.
+#' @examples
+#' v <- labelled(c(3,4,4,3,8, 9),
+#'              c(YES = 3, NO = 4, wrong_label =8, refused = 9)
+#'              )
+#' harmonize_missing_values(v)
+
+harmonize_missing_values <- function (x) {
+
+  if ( all( c("haven_labelled", "double") %in% class(x)) ) {
+    x <- labelled::to_character(x)
+  } else {
+    x <- as.character(x)
+  }
+
+  tmp <- tibble::tibble (
+    label_norm = label_normalize(x)
+  )
+
+  tmp %>%
+    mutate ( na_harmonized = dplyr::case_when (
+      grepl("inap", label_norm) ~ "inap",
+      grepl("decline|dk|refuse", label_norm) ~ "decline",
+      grepl("dont_know", label_norm) ~ "do_not_know",
+      TRUE ~ label_norm )
+    ) %>%
+    mutate ( na_numeric_value =  case_when (
+      na_harmonized == "inap"  ~  9999,
+      na_harmonized == "decline" ~ 9998,
+      na_harmonized == "do_not_know" ~ 9997,
+      TRUE ~ 8999
+    ))
 }

### 2020-06-24 data-raw/create_binary_harmonization_table.R

@@ -1,0 +1,22 @@
+binary_harm_table <- read.csv ( file.path('data-raw',
+                                          'binary_value_harmonization_table.csv'),
+                                sep = ';' )
+
+binary_harmonization <- binary_harm_table %>%
+  select ( ends_with("_1")) %>%
+  purrr::set_names( ., gsub("_1", "", names(.))) %>%
+  dplyr::bind_rows(binary_harm_table %>%
+                     select ( ends_with("_2")) %>%
+                     purrr::set_names( ., gsub("_2", "", names(.)))) %>%
+  purrr::set_names ( c("label_norm", "value_numeric", "label_harmonized"))
+
+
+binary_harmonization_table <- binary_harmonization %>%
+  distinct_all()
+
+label_harmonization_table <- binary_harmonization_table
+
+## Overwrite only if you know what you do!
+usethis::use_data(label_harmonization_table, overwrite=FALSE, internal=FALSE)
+
+

### 2020-06-24 data-raw/create_binary_metadata_table.R

@@ -88,3 +88,28 @@ unique_na_3 <- unique( label_normalize(binary_metadata$na_lab_3))
 
 c( unique_na_1 , unique_na_2, unique_na_3 )
 c ( unique_val_1, unique_val_2 )
+
+binary_value_pairs <- binary_metadata %>%
+  select ( all_of (c("label_val_1", "label_val_2")) ) %>%
+  distinct_all() %>%
+  mutate_all ( label_normalize )
+
+write.csv(binary_value_pairs, file.path("data-raw", "binary_value_pairs.csv"))
+
+
+na_harmonization <- tibble (
+  normalized_labels = c( unique_na_1 , unique_na_2, unique_na_3 )
+  ) %>%
+  filter ( nchar (normalized_labels)>0 ) %>%
+  mutate ( na_harmonized = dplyr::case_when (
+   grepl("inap", normalized_labels) ~ "inap",
+   grepl("decline|dk|refuse", normalized_labels) ~ "decline",
+   grepl("dont_know", normalized_labels) ~ "do_not_know",
+   TRUE ~ normalized_labels )
+  ) %>%
+  mutate ( na_numeric_value =  case_when (
+    na_harmonized == "inap"  ~  9999,
+    na_harmonized == "decline" ~ 9998,
+    na_harmonized == "do_not_know" ~ 9997,
+    TRUE ~ 8999
+  ))

### 2020-06-24 tests/testthat/test-harmonize_qb_vars.R

@@ -2,7 +2,6 @@ import_file_names <- c(
   'ZA7576_sample','ZA5913_sample','ZA6863_sample'
 )
 
-
 my_survey_list <- read_surveys (
   import_file_names, .f = 'read_example_file' )
 
@@ -24,12 +23,35 @@ data_frame_test <- harmonize_qb_vars (
   var_name = "var_name_suggested",
   conversion = "class_suggested" )
 
+small_trust_test <- my_survey_list[[1]]
+small_trust_test <- small_trust_test[1:4, c(1, 3, 26)]
+small_trust_metadata <- gesis_metadata_create(small_trust_test)
+
+trust_test <-  harmonize_qb_vars (
+  survey_list = small_trust_test,
+  metadata = small_trust_metadata,
+  question_block = "trust",
+  var_name = "var_name_suggested",
+  conversion = "class_suggested" )
+
+a <- as.numeric(unlist(unclass(small_trust_test$qa6a_8)))
+a
+
+harmonized_trust_labels <- names(attr(trust_test$trust_in_institutions_national_government, "labels"))
+
+test_that("correct value is returned", {
+  expect_equal(as.numeric(trust_test$trust_in_institutions_national_government),
+  c(9998,1,0, 0))
+  expect_equal(harmonized_trust_labels,
+               c('decline', 'tend_to_trust', 'tend_not_to_trust'))
+})
+
 test_that("correct structure is returned", {
   expect_equal("data.frame" %in% class(harmonized_test), TRUE)
 })
 
 test_that("data.frame converted to list and correctly returned", {
-  expect_equal("data.frame" %in% class(data_frame_test ), TRUE)
+  expect_equal("data.frame" %in% class(data_frame_test), TRUE)
 })
 
 test_that("error handling works in harmonize_qb_vars", {

### 2020-06-24 tests/testthat/test-harmonize_to_numeric.R

@@ -1,0 +1,10 @@
+v <- labelled::labelled(c(3,4,4,3,8,9),
+             c(YES = 3, NO = 4, wrong_label = 8, refused = 9)
+             )
+
+input_labelled_var = harmonize_value_labels(v)
+
+test_that("correct values are retured", {
+  expect_equal(harmonize_to_numeric(harmonized_labelled_var = input_labelled_var),
+               c(1,0,0,1,NA_real_, NA_real_))
+})

### 2020-06-24 tests/testthat/test-harmonize_value_labels.R

@@ -1,0 +1,11 @@
+
+v <- labelled(c(3,4,4,3,8, 9),
+             c(YES = 3, NO = 4, wrong_label = 8, refused = 9)
+             )
+
+test_that("multiplication works", {
+  expect_equal(as.numeric ( harmonize_value_labels(v) ),
+               c(1,0,0, 1, 8999,9998))
+  expect_equal(labelled::to_character ( harmonize_value_labels(v) ),
+               c("yes", "no", "no", "yes", "wrong_label", "decline"))
+})

### 2020-06-25 R/gesis_metadata_create.R
many exceptions with various labelled classes

@@ -84,6 +84,7 @@ gesis_metadata_create <- function ( survey_list ) {
     value_labels_df$factor_levels <- all_val_labels
     value_labels_df$na_levels <- sapply ( dat, labelled::na_values)
 
+    ## Establish the valid range of categories without missings -----
     fn_valid_range <- function ( x ) {
       element_name <- names ( value_labels_df$factor_levels[x] )
       f <- unlist(value_labels_df$factor_levels[x])
@@ -92,8 +93,9 @@ gesis_metadata_create <- function ( survey_list ) {
 
       if ( ! is.null(valid_range) ) {
         names( valid_range) <- gsub(paste0(element_name, "."), "", names(valid_range))
-     }
-      valid_range
+      }
+      normalized_labels <- label_normalize(names(valid_range))
+      valid_range [which ( ! grepl( "inap|refus|dk|decline", normalized_labels))]
     }
 
     value_labels_df$valid_range <- sapply (

### 2020-06-25 R/harmonize_qb_vars.R

@@ -70,7 +70,8 @@ harmonize_qb_vars <- function ( survey_list,
     select_qb <- names(dat)[ names(dat) %in% qb_vars_orig ]
 
     subset_metadata <- qb_metadata %>%
-      filter ( var_name_orig %in% select_qb )
+      filter ( var_name_orig %in% select_qb ) %>%
+      filter ( filename == unique(dat$filename)[1] )
 
     new_names <- unique(as.character(subset_metadata[[var_name]]))
 
@@ -86,10 +87,12 @@ harmonize_qb_vars <- function ( survey_list,
 
     if ( ncol(question_block) > 0 ) {
       question_block <- question_block %>%
-        purrr::set_names ( new_names )  %>%
-        convert_class (.,
-                       metadata = subset_metadata,
-                       conversion = conversion )
+        purrr::set_names ( new_names )
+
+      question_block <- convert_class (
+         dat = question_block,
+         metadata = subset_metadata,
+         conversion = conversion )
 
       dplyr::bind_cols(skeleton, question_block)
     } else {
@@ -120,7 +123,6 @@ harmonize_qb_vars <- function ( survey_list,
 
       for ( i in 2:number_surveys  ) {
         this_survey <- subsetted_surveys[[i]]
-
 
         if (
           ## Only join surveys with existing variables in the QB

### 2020-06-25 R/utils.R

@@ -31,7 +31,8 @@ class_suggest <- function(metadata) {
 
   suggestions <- metadata %>%
     mutate ( dummy = vapply(valid_range, is_dummy, logical(1))) %>%
-    mutate ( conversion_suggestion = dplyr::case_when (
+    mutate ( conversion_suggested = dplyr::case_when (
+      qb         == "id" ~ 'character',
       class_orig %in% c("numeric","character") ~ class_orig,
       class_orig == 'haven_labelled' & length_cat_range == 2  ~ 'harmonized_labelled',
       class_orig == 'haven_labelled_spss' & length_cat_range == 2  ~ 'harmonized_labelled',

### 2020-06-25 tests/testthat/test-gesis_metadata_create.R

@@ -1,5 +1,5 @@
 library(testthat)
-
+library(eurobarometer)
 import_file_names <- c(
   'ZA7576_sample','ZA5913_sample','ZA6863_sample'
 )
@@ -9,7 +9,7 @@ data ("ZA5913_sample")
 survey_list_2 <- read_surveys (
   import_file_names, .f = 'read_example_file' )
 
-returned_metadata <- gesis_metadata_create( survey_list_2 )
+returned_metadata <- gesis_metadata_create( survey_list = survey_list_2 )
 
 single_data_frame <- gesis_metadata_create( survey_list = ZA5913_sample )
 
@@ -17,8 +17,9 @@ test_that("correct structure is returned", {
   expect_equal(names(returned_metadata),
                c('filename', 'qb', 'var_name_orig',  'var_label_orig',
                  'var_label_norm', 'var_name_suggested',
-                 "length_cat_range", "length_na_range",'n_categories',
+                 "length_cat_range", "length_na_range", "length_total_range",
+                 'n_categories',
                  'factor_levels',  'valid_range', 'na_levels',
-                 'class_orig', 'class_suggested'))
+                 'class_orig', 'conversion_suggested'))
   expect_equal(unique(single_data_frame$filename), "not_given")
 })

### 2020-06-25 tests/testthat/test-harmonize_qb_vars.R

@@ -6,7 +6,7 @@ my_survey_list <- read_surveys (
   import_file_names, .f = 'read_example_file' )
 
 my_metadata <- gesis_metadata_create( survey_list = my_survey_list )
-
+names ( my_metadata)
 wrong_character_input <- c("a", "b")
 
 harmonized_test <- harmonize_qb_vars (
@@ -14,14 +14,7 @@ harmonized_test <- harmonize_qb_vars (
   metadata = my_metadata,
   question_block = "socio-demography",
   var_name = "var_name_suggested",
-  conversion = "class_suggested" )
-
-data_frame_test <- harmonize_qb_vars (
-  survey_list = my_survey_list[[1]],
-  metadata = my_metadata,
-  question_block = "socio-demography",
-  var_name = "var_name_suggested",
-  conversion = "class_suggested" )
+  conversion = "conversion_suggested" )
 
 small_trust_test <- my_survey_list[[1]]
 small_trust_test <- small_trust_test[1:4, c(1, 3, 26)]
@@ -32,18 +25,20 @@ trust_test <-  harmonize_qb_vars (
   metadata = small_trust_metadata,
   question_block = "trust",
   var_name = "var_name_suggested",
-  conversion = "class_suggested" )
+  conversion = "conversion_suggested" )
 
 a <- as.numeric(unlist(unclass(small_trust_test$qa6a_8)))
 a
 
-harmonized_trust_labels <- names(attr(trust_test$trust_in_institutions_national_government, "labels"))
+harmonized_trust_labels <- names(
+  attr(trust_test$trust_in_institutions_national_government, "labels"))
+harmonized_trust_labels
 
 test_that("correct value is returned", {
   expect_equal(as.numeric(trust_test$trust_in_institutions_national_government),
   c(9998,1,0, 0))
   expect_equal(harmonized_trust_labels,
-               c('decline', 'tend_to_trust', 'tend_not_to_trust'))
+               c('declined', 'tend_to_trust', 'tend_not_to_trust'))
 })
 
 test_that("correct structure is returned", {
@@ -51,7 +46,7 @@ test_that("correct structure is returned", {
 })
 
 test_that("data.frame converted to list and correctly returned", {
-  expect_equal("data.frame" %in% class(data_frame_test), TRUE)
+  expect_equal("data.frame" %in% class(trust_test), TRUE)
 })
 
 test_that("error handling works in harmonize_qb_vars", {
[…]

### 2020-06-25 R/harmonize_qb_vars.R
remove testing

@@ -110,9 +110,9 @@ harmonize_qb_vars <- function ( survey_list,
     }
   }
 
-  dat_1 <- survey_subset ( dat = survey_list [[1]] )
-  dat_2 <- survey_subset ( dat = survey_list [[2]] )
-  dat_3 <- survey_subset ( dat = survey_list [[3]] )
+  #dat_1 <- survey_subset ( dat = survey_list [[1]] )
+  #dat_2 <- survey_subset ( dat = survey_list [[2]] )
+  #dat_3 <- survey_subset ( dat = survey_list [[3]] )
 
   subsetted_surveys <- lapply (survey_list, survey_subset)
 

### 2020-06-26 R/eurobarometer.R
correct spelling

@@ -1,7 +1,7 @@
 #' eurobarometer: A package to create tidy Eurobarometer microdata files.
 #'
 #' The package eurobarometer has a number of related functions to
-#' analyze metadata; labelling variables with the help of metadata;
+#' analyse metadata; labelling variables with the help of metadata;
 #' harmonizing variable names, classes and labels.
 #'
 #' @section labelling functions:
@@ -23,11 +23,13 @@
 #' \code{\link{panel_create}} creates a skeleton panel with unique
 #' \code{panel_id} and other identification variables.
 #' \code{\link{convert_class}} to make class conversions based on a
-#' metadata analyais.
+#' metadata analysis.
 #' \code{\link{harmonize_qb_vars}} is harmonizing the names and classes of
 #' questions in a question block.
 #' \code{\link{harmonize_missing_values}} is harmonizing the missing values
 #' in a labelled vector.
+#' \code{\link{harmonize_value_lables}} is harmonizing the the value
+#' labels in 2 or 3 category labelled variables.
 #'
 #' @section sample data files:
 #' \code{\link{ZA5913_sample}} A small subset of the ZA5913:

### 2020-06-26 R/harmonize_value_labels.R

@@ -12,6 +12,7 @@
 #' @return A labelled vector containing harmonized numeric values and
 #' labels, and in \code{attr(v, num_orig)} the original numeric values,
 #' \code{attr(v, labels_orig)} the original labelling.
+#' @family harmonization functions
 #' @examples
 #' v <- labelled::labelled(
 #'        c(3,4,4,3,8, 9),

### 2020-06-29 R/read_surveys.R
new vignette

@@ -7,9 +7,10 @@
 #' @param .f A function to import the surveys with.
 #' Defaults to \code{'read_example_file'}. For SPSS files,
 #' \code{'read_spss_survey'} is recommended, which is a
-#' well-parameterised version of \code{\link[haven]{read_spss}}.
+#' well-parametrised version of \code{\link[haven]{read_spss}}.
 #' @param file_path A file path where the \code{import_file_names}
-#' can be found. Defaults to the working directory: \code{file_path = NULL}.
+#' can be found.
+#' Defaults to the working directory: \code{file_path = NULL}.
 #' @return A list of the surveys.  Each element of the list is a data
 #' frame. The respective file names are added to each data frame as a
 #' constant column \code{filename}.
@@ -28,7 +29,7 @@ read_surveys <- function ( import_file_names,
 
 
   if ( !is.null(file_path) ) {
-    if ( dir.exists(file_path)) {
+    if ( dir.exists(file_path) ) {
       read_file_names <- file.path(file_path, import_file_names)
     } else {
       stop(file_path, " cannot be found.")

### 2020-06-29 data-raw/create_3_values_harmonization_table.R

@@ -1,6 +1,7 @@
 three_harm_table <- read.csv ( file.path('data-raw',
                                           'three_value_harmonization_table.csv'),
-                                sep = ';' )
+                                sep = ';' ) %>%
+  filter (!is.na(label_harmonized_1 ))
 
 three_harmonization <- three_harm_table %>%
   select ( ends_with("_1")) %>%

### 2020-07-01 R/eurobarometer_labelled.R
new class constructor with test

@@ -1,0 +1,94 @@
+#' Construct A eurobarometer_labelled Variable
+#'
+#' @param x A vector of numeric values of class integer, numeric,
+#' haven_labelled or haven_labelled_spss.
+#' @param value_labels The \code{labels} attribute, defaults to \code{NULL}.
+#' @param na_values The \code{na_values} attribute, defaults to \code{NULL}.
+#' @param na_range The \code{na_range} attribute, defaults to \code{NULL}.
+#' @param labels_orig The \code{labels_orig} attribute which contains the original
+#' value_labels before harmonization or normalization, defaults to \code{NULL}.
+#' @param var_name_orig The \code{var_name_orig} attribute, which
+#' contains the original variable name,  defaults to \code{NULL}.
+#' @param qb The \code{qb} attribute, which contains the identifier of the
+#' question block, defaults to \code{NULL}.
+#' @param conversion The \code{conversion} attribute which contains the
+#' name of the function that created the variable, defaults to \code{NULL}.
+#' @return A vector with metadata attributes of class \code{eurobarometer_labelled}.
+#' @examples
+#' eurobarometer_labelled(x = c(6,7,8,9,6,7,8),
+#'    labels =c("HI" = 6, "HELLO" = 7,
+#'              "HOWDY" = 8, "BYE" = 9),
+#'    na_values = 9,
+#'    qb = "Example Block",
+#'    conversion = "eurobarometer_labelled"
+#' )
+#' @export
+
+eurobarometer_labelled <- function ( x = NULL,
+                                     value_labels = NULL,
+                                     na_values = NULL,
+                                     na_range = NULL,
+                                     labels_orig = NULL,
+                                     var_name_orig = NULL,
+                                     qb = NULL,
+                                     conversion = NULL) {
+
+  if ( is.null(x) ) {
+    stop ( "Parameter x must be given.")
+  }
+
+  if (class(x)[1] =='logical' ) {
+    warning ("class <logical> is converted to <integer>.")
+    x <- as.integer(x)
+  }
+  if ( any( class(x) %in% c("integer", "numeric")) ) {
+    ebl <- haven::labelled_spss( x = x,
+                          labels = value_labels,
+                          na_values = na_values,
+                          na_range = na_range )
+  } else if ( any(class(x) %in% "haven_labelled_spss") ) {
+    ebl <- x
[…]

### 2020-07-01 tests/testthat/test-eurobarometer_labelled.R

@@ -1,0 +1,58 @@
+
+### Test wrong parameter classes ---------------------------------
+test_that("Appropriate warnings are given", {
+  expect_warning(eurobarometer_labelled (x = c(TRUE, TRUE, FALSE, NA)))
+  expect_warning(eurobarometer_labelled ( x = c("hello", "hi", "hey"),
+                                          labels_orig = c(1:2) )
+  )
+})
+
+test_that("Errors are found", {
+  expect_error(eurobarometer_labelled ( x = data.frame ( x = 1:5)))
+})
+
+### Test if the constructor works as expected
+
+v <- labelled::labelled(c(3,4,4,3,8, 9),
+                        c(YES = 3, NO = 4, `WRONG LABEL` = 8, REFUSED = 9)
+)
+correct_labels <- c(3,4,8,9)
+names(correct_labels) <- c("YES", "NO", "WRONG LABEL", "REFUSED")
+correct_yes_no <- c(1,0)
+names(correct_yes_no) <- c("yes", "no")
+
+
+eurobarometer_labelled(x = c(6,7,8,9, 6,7,8),
+                       labels =c("HI" = 6, "HELLO" = 7,
+                                 "HOWDY" = 8, "BYE" = 9),
+                       na_values = 9,
+                       qb = "Example Block",
+                       conversion = "eurobarometer_labelled")
+
+x <-haven::labelled_spss(
+  x = c(1,2,2,1,99998,99999),
+  labels =  c(yes = 1, no = 2, wrong_label = 99998, refused = 99999),
+  na_values = NULL,
+  na_range = 99998:99999,
+  label = "Variable Example"
+)
+
+test_that("constructor works", {
+  expect_equal(labelled::val_labels ( eurobarometer_labelled(v)),
+               correct_labels)
+  expect_equal(labelled::val_labels (
+    eurobarometer_labelled(x = c(0,1),
+                           value_labels = c( 'yes' = 1, 'no' = 0)
+                           )),
+    correct_yes_no)
+  expect_equal(labelled::na_range (eurobarometer_labelled(x)),
+               c(99998,99999))
+  expect_equal(class ( eurobarometer_labelled(x)),
+               c("eurobarometer_labelled", class(x)))
+})
+
+
+
+
+
+

### 2020-07-01 README.Rmd
update README

@@
 ```{r, include = FALSE}
+figures_rel_path_from_root <- rprojroot::find_root_file(
+  "man", "figures", "README", 
+  criterion = rprojroot::has_file("DESCRIPTION"))
+
 knitr::opts_chunk$set(
@@
   comment = "#>",
-  fig.path = "man/figures/README-",
+  fig.path = figures_rel_path_from_root,
   out.width = "100%"
@@
 ```
-## Workflow
-
-Not all elements of the [Workflow](http://eurobarometer.danielantal.eu/articles/workflow.html) are implemented yet.
 
@@
 
-The SPSS files in GESIS are not ready for programmatic use in R, because the variable names often contain reserved special characters.
-You should use `gesis_metadata_create()` to understand what are the possible problems in the SPSS file and how to resolve them.
-(See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html))
+The SPSS files in GESIS are not ready for programmatic use in R or for joining into longitudinal panels.
 
-* The naming functions `normalize_names()` removes regex symbols, whitespace, and basic inconsistencies in abbreviations.
-The function `canonical_names()`harmonizes the names across various GESIS files, so that they can be joined into panels across time.
+1.
+We import the GESIS SPSS files with the help of the [haven](https://haven.tidyverse.org/) package.
+In order to preserve the metadata in the files (particularly variable- and value labels, and user-defined missing values) these are converted to labelled classes.
 
-* You can read in the SPSS file to R in [haven](https://haven.tidyverse.org/), which will result in many `labelled` variables.
-The metadata file contains a suggestion of the most practical class conversion for data analysis.
-You can review and change this suggestion and then make the conversions at once with `convert_class()`
+2.
+We analyse the files with `gesis_metadata_create()` to understand what are the possible problems in the SPSS file and how to resolve them.
+(See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html) or `vignette("metadata")`)
 
-See `vignette("metadata")`
+3.
+We suggest various [naming](http://eurobarometer.danielantal.eu/articles/variable_names.html) and [harmonization](http://eurobarometer.danielantal.eu/articles/workflow.html) changes in the data as a data processing step.
+These suggested changes can be modified by the user.
+
+4.
+We convert the variables to a modified version of the `haven_labelled_spss` class that contains the harmonized values, harmonized labels, harmonized missing values codes of the variable with their history (original values, labels, and function applied to change them)  for reproducibility.
+The [eurobarometer_labelled](http://eurobarometer.danielantal.eu/articles/eurobarometer_class.html) class inherits most of the functionality of the packages [haven](https://haven.tidyverse.org/),  [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) and are expected to work well with [sjlabelled](https://strengejacke.github.io/sjlabelled/articles/labelleddata.html).
+
+```{r, echo=FALSE, fig.align='center', out.width='90%', fig.cap="Workflow from 'Introduction to labelled by Joseph Larmaranged'"}
+knitr::include_graphics(file.path(figures_rel_path_from_root, 
+                                  'labelled_workflow.png'))
+```
+
+Our approach follows `Approach B` to allow the conversion of harmonized and joined files back into SPSS.
+
+5.
+The [eurobarometer_labelled](http://eurobarometer.danielantal.eu/articles/eurobarometer_class.html) class comes with simple methods that can convert all variables to the base R types of numeric or factor, and they are ready to be used with any R statistical packages and tidyverse data processing tools.
+Alternatively, they can be exported back to [SPSS](https://haven.tidyverse.org/reference/read_spss.html) or [Stata](https://haven.tidyverse.org/reference/read_dta.html) files with the functions of the [haven](https://haven.tidyverse.org/) package.
 
@@
 
-* Currently `code_nuts1()` and `code_nuts2()` is implemented but this will be replaced with a more comprehensive `NUTS0`-`NUTS3` level helper function to make the regional coding consistent.
-
 ## Code of Conduct

### 2020-07-01 README.Rmd
documentation changes

@@
 
-The SPSS files in GESIS are not ready for programmatic use in R or for joining into longitudinal panels.
+The microdata of the European Commissions Eurobarometer surveys are stored and accessible in [GESIS](https://www.gesis.org/home).
+They are not ready for programmatic use in R or for joining into longitudinal panels.
+We created this package to help this procedure following the principles of reproducible research.
 
@@
 By contributing to this project, you agree to its terms.
+
+## Disclaimer
+
+The authors of this package are not affiliated with [GESIS](https://www.gesis.org/home), or the producer of the Eurobarometer surveys, or the European Commission.
+For more information about the European Commission's Eurobarometer history, visit the European Commission's [Public Opinion](https://ec.europa.eu/commfrontoffice/publicopinion/index.cfm) website.

### 2020-07-03 data-raw/variable-keywords.Rmd
draft variable-keywords vignette

---
title: "Tagging variables with keywords"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vocabulary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
options(width = 160)
```

```{r setup}
library(eurobarometer)
library(tidyverse)
library(tibble)
library(knitr)
library(kableExtra)
library(tidytext)
library(janitor)
```

An important aspect of survey data harmonization - in addition to standardizing response coding and value labels - is the identification of variables that measure the same thing, and standardizing the name of these variables. The process is faciliated by the fact that variables typically have labels that succintly describe the content of the variable, and sometimes also contain information about the coding of responses (e.g., length of scale), the unit (e.g., age in years or categories) and whether the variable contains "original" respondent responses or some transformation of responses (e.g. recoding or index).

Fortunately, in many multi-wave surveys, such as the Eurobarometer, there is a degree of consistency in how variables are labelled in different editions of the survey, and most survey project try to follow some labelling conventions.

Unfortunately, consistency is limited and conventions are not followed strictly, which makes standardizing labels a semi-manual task. Some of these issues stem from the limits in the number of characters of variable labels in common statistical software packages (e.g., SPSS), which leads to non-standard abbreviations.

This vignette outlines a procedure for arriving at standardized variable names via tokenization of normalized variable labels (`var_label_norm`), which are created by [`label_normalize`](http://eurobarometer.danielantal.eu/reference/label_normalize.html) as part of [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html). The logic is similar to text analysis, in that a document-term matrix is created with each variable label treated as a (very short) document, and terms treated as keywords. variables that contain the same sets of keywords can be - with some margin of error - assumed to correspond to the same or equivalent survey questions.

## Procedure

0. Read metadata table (created by [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html)),

1. Standardize the spelling of certain often abbreviated words (e.g., 'gov', 'govnmt', and 'govmnt' for 'government'),

2. Tokenize,

3. Eliminate stop words (but be careful to not eliminate too many, e.g. 'right' or 'working),

4. Create a document-term matrix (actually, variable label - keyword matrix),

5. Sort the columns according to the declining frequency of the keywords,

6. Create standardized variable labels by concatenating keywords.

### 0. Metadata table

The table contains metadata from 97 Eurobarometer files (each file corresponding to one EB round).

```{r}
metadata_database <- readRDS( file.path('../data-raw', 'eb_metadata_database_20200628.rds'))
```

### 1. Standardize the spelling

This would perhaps look better as a translation table instead of regex.

Maybe the recoding of "leftover question numbers at the beginning of value labels" can become part of `label_normalize`?

```{r}

dat <- metadata_database %>%
  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm) %>%

  # eliminate variables that look like region identifiers, weights, archive IDs, etc.
  filter(!grepl("recoded|(rec)|nuts|iso_3166|iso3166|region_ii", var_label_norm)) %>%
  filter(!grepl("^region_|sample_id|(sum)|(index)|filename", var_label_norm)) %>%
  filter(!grepl("icpsr|gesis|zacat|weight_", var_label_norm)) %>%

  # clean up variable labels to unify the spelling of certain common words and abbreviation
  mutate(var_id = paste0(filename, " ", var_name_orig),

         var_label_norm1 = gsub("^cap_", "common-agricultural-policy_", var_label_norm),
         var_label_norm1 = gsub("_cap_", "_common-agricultural-policy_", var_label_norm),

         # leftover question numbers at the beginning of value labels
         var_label_norm1 = gsub("^w[0-9]{1,2}", "", var_label_norm1),
         var_label_norm1 = gsub("^[0-9]{1-3}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[0-9]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", var_label_norm1),

         # country names
         var_label_norm1 = gsub("united_kingdom", "united-kingdom", var_label_norm1),
         var_label_norm1 = gsub("czech_republic", "czech-republic", var_label_norm1),
         var_label_norm1 = gsub("great_britain", "great-britain", var_label_norm1),
         var_label_norm1 = gsub("united_germany", "united-germany", var_label_norm1),
         var_label_norm1 = gsub("united_nations", "united-nations", var_label_norm1),
         var_label_norm1 = gsub("ivory_coast", "ivory-coast", var_label_norm1),

         # commonly abbreviated words
         var_label_norm1 = gsub("particip_", "participation_", var_label_norm1),
         var_label_norm1 = gsub("environmt_|environm_", "environment_", var_label_norm1),
         var_label_norm1 = gsub("citizenshp_", "citizenship_", var_label_norm1),
         var_label_norm1 = gsub("_rght", "_right", var_label_norm1),
         var_label_norm1 = gsub("_fulfld", "_fullfilled", var_label_norm1),
         var_label_norm1 = gsub("_pessim", "_pessimistic", var_label_norm1),
         var_label_norm1 = gsub("_convcd", "_convinced", var_label_norm1),
         var_label_norm1 = gsub("not_invo", "not_involved", var_label_norm1),
         var_label_norm1 = gsub("intelligenc$", "intelligence", var_label_norm1),
         var_label_norm1 = gsub("gvrnm", "government", var_label_norm1),
         var_label_norm1 = gsub("ctry|cntry|\\[country\\]", "country", var_label_norm1),
         var_label_norm1 = gsub("_accom$|_accomod$", "_accomodation", var_label_norm1),
         var_label_norm1 = gsub("_activit", "_activity", var_label_norm1),
         var_label_norm1 = gsub("_cig|_cigs", "_cigarettes", var_label_norm1),
         var_label_norm1 = gsub("cildren", "children", var_label_norm1),
         var_label_norm1 = gsub("citizenshp", "citizenship", var_label_norm1),
         var_label_norm1 = gsub("collgs|collegues", "colleagues", var_label_norm1),
         var_label_norm1 = gsub("competitivn", "competition", var_label_norm1),
         var_label_norm1 = gsub("comptrs", "computers", var_label_norm1),
         var_label_norm1 = gsub("confidfence", "confidence", var_label_norm1),
         var_label_norm1 = gsub("conflicte$", "conflicts", var_label_norm1),
         var_label_norm1 = gsub("conn$|connec$", "connection", var_label_norm1),
         var_label_norm1 = gsub("contrls", "controls", var_label_norm1),
         var_label_norm1 = gsub("communit$", "community", var_label_norm1),
         var_label_norm1 = gsub("situatn_", "situation", var_label_norm1),
         var_label_norm1 = gsub("freq_", "frequency_", var_label_norm1),
         var_label_norm1 = gsub("hh", "household", var_label_norm1),
         var_label_norm1 = gsub("_pers_", "_personal_", var_label_norm1),
         var_label_norm1 = gsub("prov_", "provider_", var_label_norm1),
         var_label_norm1 = gsub("info_", "information_", var_label_norm1),
         var_label_norm1 = gsub("newspv|newspapers", "newspaper", var_label_norm1),
         var_label_norm1 = gsub("_newsp_|_newspap_", "_newspaper_", var_label_norm1),
         var_label_norm1 = gsub("_newsp$|_newspap$", "_newspaper_", var_label_norm1),
         var_label_norm1 = gsub("serv_", "services_", var_label_norm1),
         var_label_norm1 = gsub("financ_", "financial_", var_label_norm1),
         var_label_norm1 = gsub("_op_", "_operator_", var_label_norm1),

         # two- or three-word phrases to be kept as keywords
         var_label_norm1 = gsub("tv_channels", "tv-channels", var_label_norm1),
         var_label_norm1 = gsub("radio_stations", "radio-stations", var_label_norm1),
         var_label_norm1 = gsub("regional_local", "regional-local", var_label_norm1),
         var_label_norm1 = gsub("air_conditioning", "air-conditioning", var_label_norm1),
         var_label_norm1 = gsub("pol_party|political_party", "political-party", var_label_norm1),
         var_label_norm1 = gsub("electoral_particip_", "electoral-participation_", var_label_norm1),
         var_label_norm1 = gsub("electoral_particip$", "electoral-participation$", var_label_norm1),
         var_label_norm1 = gsub("trust_in_institutions", "trust", var_label_norm1),
         var_label_norm1 = gsub("justice_legal_system|justice_nat_legal_system", "justice-system", var_label_norm1),
         var_label_norm1 = gsub("non_govmnt_org|non_govnmt_org", "ngo", var_label_norm1),
         var_label_norm1 = gsub("polit_parties|political_parties", "political-parties", var_label_norm1),
         var_label_norm1 = gsub("reg_loc_public_authorities|reg_local_authorities|reg_loc_authorities|reg_local_public_authorities|rg_lc_authorities", "regional-local-authorities", var_label_norm1),
         var_label_norm1 = gsub("written_press", "press", var_label_norm1),
         var_label_norm1 = gsub("pers_influence", "personal-influence", var_label_norm1),
         var_label_norm1 = gsub("job_situation", "job-situation", var_label_norm1),
         var_label_norm1 = gsub("religious_inst$", "religious-institutions", var_label_norm1),
         var_label_norm1 = gsub("polit_matters|political_matters", "political-matters", var_label_norm1),
         var_label_norm1 = gsub("pol_discussion|polit_discussion|political_discussion", "political-discussion", var_label_norm1),
         var_label_norm1 = gsub("nat_resources", "natural-resources", var_label_norm1),
         var_label_norm1 = gsub("non_euro_zone", "non-euro-zone", var_label_norm1),
         var_label_norm1 = gsub("party_attachment", "party-attachment", var_label_norm1),
         var_label_norm1 = gsub("which_issue", "which-issue", var_label_norm1),
         var_label_norm1 = gsub("economic_situation|economic_sit", "economic-situation", var_label_norm1),
         var_label_norm1 = gsub("public_debt", "public-debt", var_label_norm1),
         var_label_norm1 = gsub("left_right", "left-right", var_label_norm1),
         var_label_norm1 = gsub("important_issues_cntry", "important-issues-cntry", var_label_norm1),
         var_label_norm1 = gsub("important_issues|import_issues", "important-issues", var_label_norm1),
         var_label_norm1 = gsub("important_values_eu", "important-values-eu", var_label_norm1),
         var_label_norm1 = gsub("important_values_pers", "important-values-personal", var_label_norm1),
         var_label_norm1 = gsub("important_life_domains", "important-life-domains", var_label_norm1),
         var_label_norm1 = gsub("quality_of_life|life_quality", "life-in-general", var_label_norm1),
         var_label_norm1 = gsub("household_composition", "household-composition", var_label_norm1),
         var_label_norm1 = gsub("size_of_community", "size-of-community", var_label_norm1),
         var_label_norm1 = gsub("head_of_hh", "head-of-hh", var_label_norm1),
         var_label_norm1 = gsub("last_job", "last-job", var_label_norm1),
         var_label_norm1 = gsub("level_in_society", "level-in-society", var_label_norm1),
         var_label_norm1 = gsub("self_placement", "self-placement", var_label_norm1),
         var_label_norm1 = gsub("social_class|soc_class", "social-class", var_label_norm1),
         var_label_norm1 = gsub("education_level", "education-level", var_label_norm1),
         var_label_norm1 = gsub("living_conditions", "living-conditions", var_label_norm1),
         var_label_norm1 = gsub("_nat_", "_national_", var_label_norm1),
         var_label_norm1 = gsub("^nat_", "national_", var_label_norm1),
         var_label_norm1 = gsub("_nat$", "_national", var_label_norm1),
         var_label_norm1 = gsub("europ_", "european_", var_label_norm1),
         var_label_norm1 = gsub("12_months_|12_mo_", "12-months_", var_label_norm1),
         var_label_norm1 = gsub("national_parliament", "national-parliament", var_label_norm1),
         var_label_norm1 = gsub("national_government", "national-government", var_label_norm1),
         var_label_norm1 = gsub("national_economy", "national-economy", var_label_norm1),
         var_label_norm1 = gsub("good_bad", "good-bad", var_label_norm1),
         var_label_norm1 = gsub("court_of_auditors", "court-of-auditors", var_label_norm1),
         var_label_norm1 = gsub("council_of_ministers", "council-of-ministers", var_label_norm1),

         # EU-related abbreviations and phrases
         var_label_norm1 = gsub("euro_12", "euro-12", var_label_norm1),
         var_label_norm1 = gsub("european_parliament|eu_parl", "european-parliament", var_label_norm1),
         var_label_norm1 = gsub("european_union", "european-union", var_label_norm1),
         var_label_norm1 = gsub("european_unification|europ_unification|europ_unif", "european-unification", var_label_norm1),
         var_label_norm1 = gsub("european_economy|europ_economy", "european-economy", var_label_norm1),
         var_label_norm1 = gsub("european_ombudsman", "european-ombudsman", var_label_norm1),
         var_label_norm1 = gsub("european_elections|ep_election", "european-elections", var_label_norm1),
         var_label_norm1 = gsub("european_elec|european_elect|europ_elec", "european-elections", var_label_norm1),
         var_label_norm1 = gsub("european_economy", "european-economy", var_label_norm1),
         var_label_norm1 = gsub("european_currency|europ_currency", "european-currency", var_label_norm1),
         var_label_norm1 = gsub("europ_court_of_auditors|eur_court_of_auditors", "eu-court-of-auditors", var_label_norm1),
         var_label_norm1 = gsub("europ_court_of_justice|european_court_of_justice", "eu-court-of-justice", var_label_norm1),
         var_label_norm1 = gsub("econ_and_social_committee|economic_and_soc_committee|econ_and_soc_committee", "econ-and-soc-committee", var_label_norm1),
         var_label_norm1 = gsub("europ_polit_matter|europ_political_matters", "european-political-matters", var_label_norm1),
         var_label_norm1 = gsub("euro_zone", "euro-zone", var_label_norm1),
         var_label_norm1 = gsub("eu_issues", "eu-issues", var_label_norm1),
         var_label_norm1 = gsub("eu_budget", "eu-budget", var_label_norm1),
         var_label_norm1 = gsub("eu_statements", "eu-statements", var_label_norm1),
         var_label_norm1 = gsub("eu_citizenship", "eu-citizenship", var_label_norm1),
         var_label_norm1 = gsub("eu_membership|eu_membersh", "eu-membership", var_label_norm1),
         var_label_norm1 = gsub("eu_image", "eu-image", var_label_norm1),
         var_label_norm1 = gsub("eu_level", "eu-level", var_label_norm1),
         var_label_norm1 = gsub("ecb|european_central_bank", "european-central-bank", var_label_norm1),
         var_label_norm1 = gsub("european_commission", "european-commission", var_label_norm1),
         var_label_norm1 = gsub("nms", "new-member-states", var_label_norm1),
         var_label_norm1 = gsub("council_of_the_eu", "council-of-the-eu", var_label_norm1),

         # replace underscore with space
         text = gsub("_", " ", var_label_norm1)) %>%
  ungroup() %>%
  # select the needed variables
  select(filename, qb, var_label_norm1, var_label_orig, var_id, text)
```

### 2 & 3. Tokenize & Eliminate stopwords

```{r}
# list of country codes to be converted to country names
eu_countries <- c("at", "be", "bg", "cy", "cz", "dk", "de", "ee", "es", "fr",
                  "fi", "gb", "gr", "hr", "hu", "ie", "it", "lt", "lv", "lu",
                  "mt", "nl", "pl", "pt", "ro", "se", "si", "sk", "tr")

# !!! the problem here is that dk sometimes means "don't know" and sometimes Demnark !!!

tidy <- dat %>%
  # tokenize
  unnest_tokens(word, text, token = "regex", pattern = " |  ") %>%
  # eliminate stop words, with some exceptions
  anti_join(stop_words %>% filter(!word %in% c("working", "right", "other"))) %>%
  # eliminate numbers
  filter(!grepl("\\b\\d+\\b", word)) %>%
  # convert country codes to country names
  mutate(country = ifelse(word %in% eu_countries, word, NA),
         country = countrycode::countrycode(toupper(country), "iso2c", "country.name"),
         country = tolower(country),
         word = ifelse(!is.na(country), country, word),
         word = ifelse(word == "united kingdom", "united-kingdom", word),
         word = ifelse(word == "tcc", "turkish-cypriot-community", word)) %>%
  select(-country)

head(tidy)
```

```{r}
tidy %>%
  count(word) %>%
  arrange(desc(n)) %>% 
  print(n = 50)
```

### 4. Create a document-term matrix

```{r}
# select words with > 5 occurrences
tidy_small <- tidy %>%
  group_by(word) %>%
  mutate(count_of_word = n()) %>%
  filter(count_of_word > 5)

dtm <- tidy_small %>%
  distinct(var_id, var_label_norm1, word) %>%
  mutate(n = word) %>%
  spread(word, n)

dtm[1100:1110, c(1,2,110:112)]
```

### 5. Sort columns according to the declining frequency of the keywords

```{r}
# get the order of variables according to declining proportion of missing value
# (declining popularity of keywords)
order <- dtm %>%
  summarise_each(funs(sum(is.na(.))/length(.))) %>%
  t() %>% as.vector()

names(order) <- names(dtm)
```

### 6. Create standardized variable labels by concatenating keywords

Standardized variable labels `var_label_std` are concatenated keywords, where the keywords occur in the order of their frequency in the entire corpus of variable labels. As a consequence, more common - likely broader - terms will come first, and less common - more specific - terms will come later.

```{r}
dtm2 <- dtm %>% select(names(sort(order))) %>%
  mutate_at(vars(3:ncol(dtm)), funs(ifelse(is.na(.), "", .))) %>%
  unite("var_label_std", 3:ncol(dtm)) %>%
  mutate(var_label_std = gsub("(\\_)\\1+", "\\1", var_label_std)) %>%
  mutate(var_label_std = gsub("^_|_$", "", var_label_std))

dtm2 %>% 
  print(n = 50)
```

Print the most common standardized variable labels.

```{r}
dtm2 %>%
  count(var_label_std) %>%
  # exclude nationality, country (nation) codes, size of community (country-specific), language of interview
  filter(!grepl("nationality|nation_|size-of-community|language_interview", var_label_std)) %>%
  filter(n > 5) %>%
  arrange(desc(n)) %>% 
  print(n = 50)
```

## Application: political attitudes (1)

Filter on substrings of standardized variable labels `var_label_std`, and count the repetitions across datasets.

```{r}
# filter on var_label_std
dtm2 %>%
  filter(grepl("trust|satisf|left-right|interest|class", var_label_std)) %>% 
  count(var_label_std) %>%
  arrange(desc(n)) %>%
  print(n = 50)
```

## Application: political attitudes (2)

Filter on keywords in the document-term matrix, e.g. `left-right` to see filenames and variables corresponding to variables measuring the respondent's left-right placement on the spectrum of political opinions. This is an easy variable, because all standardized variable labels look pretty much the same, only those with "harmonized" or the number of categories added would require a closer look - they are likely tranformed versions of original variables.

Another thing that is noticable is that variable names are standardized starting with [ZA5913: Eurobarometer 81.2 (March 2014)](https://dbk.gesis.org/dbksearch/sdesc2.asp?no=5913&db=e&doi=10.4232/1.12172), where all left-right scale variables have the name `d1`. Prior to that, each EB wave has its own variable name for the left-right scale variable.

Altogether there are 92 such variables, and eliminating the transformed ones (harmonized or recoded), we are left with 85 variables in 85 different EB waves.

```{r}
# search document-term matrix by keyword
dtm %>%
  filter(!is.na(`left-right`)) %>%
  janitor::remove_empty(c("rows", "cols")) %>% 
  print(n = 100)
```

### 2020-07-03 data-raw/variable-keywords.Rmd
0.1.7

@@
 
-3. Eliminate stop words (but be careful to not eliminate too many, e.g. 'right' or 'working),
+3. Eliminate stop words (but be careful to not eliminate too many, e.g. 'right' or 'working'),
 
@@
 
-Maybe the recoding of "leftover question numbers at the beginning of value labels" can become part of `label_normalize`?
+> Maybe the recoding of "leftover question numbers at the beginning of value labels" can become part of `label_normalize`?
+reply in email
 
 ```{r}
-
 dat <- metadata_database %>%
-  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm) %>%
+  distinct(filename, qb, var_name_orig, 
+           var_name_orig, var_label_orig, var_label_norm) %>%
 
   # eliminate variables that look like region identifiers, weights, archive IDs, etc.
-  filter(!grepl("recoded|(rec)|nuts|iso_3166|iso3166|region_ii", var_label_norm)) %>%
-  filter(!grepl("^region_|sample_id|(sum)|(index)|filename", var_label_norm)) %>%
+  # I think that this should be the task of gesis_metadata_create, 
+  # which should identify ID, protocol, metadata special wars. 
+  filter(
+    !grepl("recoded|(rec)|nuts|iso_3166|iso3166|region_ii", 
+           var_label_norm)) %>%
+  filter(
+    !grepl("^region_|sample_id|(sum)|(index)|filename",
+           var_label_norm)) %>%
   filter(!grepl("icpsr|gesis|zacat|weight_", var_label_norm)) %>%
@@
   # clean up variable labels to unify the spelling of certain common words and abbreviation
-  mutate(var_id = paste0(filename, " ", var_name_orig),
-
-         var_label_norm1 = gsub("^cap_", "common-agricultural-policy_", var_label_norm),
-         var_label_norm1 = gsub("_cap_", "_common-agricultural-policy_", var_label_norm),
+  mutate(
+    var_id = paste0(filename, " ", var_name_orig),
+    var_label_norm1 = gsub(
+      "^cap_", "common-agricultural-policy_", var_label_norm),
+    var_label_norm1 = gsub(
+      "_cap_", "_common-agricultural-policy_", var_label_norm),
 
     # leftover question numbers at the beginning of value labels
-         var_label_norm1 = gsub("^w[0-9]{1,2}", "", var_label_norm1),
-         var_label_norm1 = gsub("^[0-9]{1-3}_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z]{1,2}_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[0-9]{1,2}_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^w[0-9]{1,2}", "", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^[0-9]{1-3}_", "", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^[a-z]{1,2}_", "", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^[a-z][0-9]{2}[a-z]_", "", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^[0-9]{1,2}_", "", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^[a-z][0-9]_[0-9]{1,2}_", "", var_label_norm1),
 
     # country names
-         var_label_norm1 = gsub("united_kingdom", "united-kingdom", var_label_norm1),
-         var_label_norm1 = gsub("czech_republic", "czech-republic", var_label_norm1),
-         var_label_norm1 = gsub("great_britain", "great-britain", var_label_norm1),
-         var_label_norm1 = gsub("united_germany", "united-germany", var_label_norm1),
-         var_label_norm1 = gsub("united_nations", "united-nations", var_label_norm1),
-         var_label_norm1 = gsub("ivory_coast", "ivory-coast", var_label_norm1),
+    var_label_norm1 = gsub(
+      "united_kingdom", "united-kingdom", var_label_norm1),
+    var_label_norm1 = gsub(
+      # should not we use the new official name czechia (shorter?)
+      "czech_republic", "czech-republic", var_label_norm1), 
+    var_label_norm1 = gsub(
+      "great_britain", "great-britain", var_label_norm1),
+    var_label_norm1 = gsub(
+      "united_germany", "united-germany", var_label_norm1),
+    var_label_norm1 = gsub(
+      "united_nations", "united-nations", var_label_norm1),
+    var_label_norm1 = gsub(
+      "ivory_coast", "ivory-coast", var_label_norm1),
 
     # commonly abbreviated words
-         var_label_norm1 = gsub("particip_", "participation_", var_label_norm1),
-         var_label_norm1 = gsub("environmt_|environm_", "environment_", var_label_norm1),
-         var_label_norm1 = gsub("citizenshp_", "citizenship_", var_label_norm1),
-         var_label_norm1 = gsub("_rght", "_right", var_label_norm1),
-         var_label_norm1 = gsub("_fulfld", "_fullfilled", var_label_norm1),
-         var_label_norm1 = gsub("_pessim", "_pessimistic", var_label_norm1),
-         var_label_norm1 = gsub("_convcd", "_convinced", var_label_norm1),
-         var_label_norm1 = gsub("not_invo", "not_involved", var_label_norm1),
-         var_label_norm1 = gsub("intelligenc$", "intelligence", var_label_norm1),
-         var_label_norm1 = gsub("gvrnm", "government", var_label_norm1),
-         var_label_norm1 = gsub("ctry|cntry|\\[country\\]", "country", var_label_norm1),
-         var_label_norm1 = gsub("_accom$|_accomod$", "_accomodation", var_label_norm1),
-         var_label_norm1 = gsub("_activit", "_activity", var_label_norm1),
-         var_label_norm1 = gsub("_cig|_cigs", "_cigarettes", var_label_norm1),
-         var_label_norm1 = gsub("cildren", "children", var_label_norm1),
-         var_label_norm1 = gsub("citizenshp", "citizenship", var_label_norm1),
-         var_label_norm1 = gsub("collgs|collegues", "colleagues", var_label_norm1),
-         var_label_norm1 = gsub("competitivn", "competition", var_label_norm1),
-         var_label_norm1 = gsub("comptrs", "computers", var_label_norm1),
-         var_label_norm1 = gsub("confidfence", "confidence", var_label_norm1),
-         var_label_norm1 = gsub("conflicte$", "conflicts", var_label_norm1),
-         var_label_norm1 = gsub("conn$|connec$", "connection", var_label_norm1),
-         var_label_norm1 = gsub("contrls", "controls", var_label_norm1),
-         var_label_norm1 = gsub("communit$", "community", var_label_norm1),
-         var_label_norm1 = gsub("situatn_", "situation", var_label_norm1),
-         var_label_norm1 = gsub("freq_", "frequency_", var_label_norm1),
-         var_label_norm1 = gsub("hh", "household", var_label_norm1),
-         var_label_norm1 = gsub("_pers_", "_personal_", var_label_norm1),
-         var_label_norm1 = gsub("prov_", "provider_", var_label_norm1),
-         var_label_norm1 = gsub("info_", "information_", var_label_norm1),
-         var_label_norm1 = gsub("newspv|newspapers", "newspaper", var_label_norm1),
-         var_label_norm1 = gsub("_newsp_|_newspap_", "_newspaper_", var_label_norm1),
-         var_label_norm1 = gsub("_newsp$|_newspap$", "_newspaper_", var_label_norm1),
-         var_label_norm1 = gsub("serv_", "services_", var_label_norm1),
-         var_label_norm1 = gsub("financ_", "financial_", var_label_norm1),
-         var_label_norm1 = gsub("_op_", "_operator_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "particip_", "participation_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "environmt_|environm_", "environment_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "citizenshp_", "citizenship_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_rght", "_right", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_fulfld", "_fullfilled", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_pessim", "_pessimistic", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_convcd", "_convinced", var_label_norm1),
+    var_label_norm1 = gsub(
+      "not_invo", "not_involved", var_label_norm1),
+    var_label_norm1 = gsub(
+      "intelligenc$", "intelligence", var_label_norm1),
+    var_label_norm1 = gsub(
+      "gvrnm", "government", var_label_norm1),
+    var_label_norm1 = gsub(
+      "ctry|cntry|\\[country\\]", "country", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_accom$|_accomod$", "_accomodation", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_activit", "_activity", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_cig|_cigs", "_cigarettes", var_label_norm1),
+    var_label_norm1 = gsub(
+      "cildren", "children", var_label_norm1),
+    var_label_norm1 = gsub(
+      "citizenshp", "citizenship", var_label_norm1),
+    var_label_norm1 = gsub(
+      "collgs|collegues", "colleagues", var_label_norm1),
+    var_label_norm1 = gsub(
+      "competitivn", "competition", var_label_norm1),
+    var_label_norm1 = gsub(
+      "comptrs", "computers", var_label_norm1),
+    var_label_norm1 = gsub(
+      "confidfence", "confidence", var_label_norm1),
+    var_label_norm1 = gsub(
+      "conflicte$", "conflicts", var_label_norm1),
+    var_label_norm1 = gsub(
+      "conn$|connec$", "connection", var_label_norm1),
+    var_label_norm1 = gsub(
+      "contrls", "controls", var_label_norm1),
+    var_label_norm1 = gsub(
+      "communit$", "community", var_label_norm1),
+    var_label_norm1 = gsub(
+      "situatn_", "situation", var_label_norm1),
+    var_label_norm1 = gsub(
+      "freq_", "frequency_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "hh", "household", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_pers_", "_personal_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "prov_", "provider_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "info_", "information_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "newspv|newspapers", "newspaper", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_newsp_|_newspap_", "_newspaper_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_newsp$|_newspap$", "_newspaper_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "serv_", "services_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "financ_", "financial_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_op_", "_operator_", var_label_norm1),
 
     # two- or three-word phrases to be kept as keywords
-         var_label_norm1 = gsub("tv_channels", "tv-channels", var_label_norm1),
-         var_label_norm1 = gsub("radio_stations", "radio-stations", var_label_norm1),
-         var_label_norm1 = gsub("regional_local", "regional-local", var_label_norm1),
-         var_label_norm1 = gsub("air_conditioning", "air-conditioning", var_label_norm1),
-         var_label_norm1 = gsub("pol_party|political_party", "political-party", var_label_norm1),
-         var_label_norm1 = gsub("electoral_particip_", "electoral-participation_", var_label_norm1),
-         var_label_norm1 = gsub("electoral_particip$", "electoral-participation$", var_label_norm1),
-         var_label_norm1 = gsub("trust_in_institutions", "trust", var_label_norm1),
-         var_label_norm1 = gsub("justice_legal_system|justice_nat_legal_system", "justice-system", var_label_norm1),
-         var_label_norm1 = gsub("non_govmnt_org|non_govnmt_org", "ngo", var_label_norm1),
-         var_label_norm1 = gsub("polit_parties|political_parties", "political-parties", var_label_norm1),
-         var_label_norm1 = gsub("reg_loc_public_authorities|reg_local_authorities|reg_loc_authorities|reg_local_public_authorities|rg_lc_authorities", "regional-local-authorities", var_label_norm1),
-         var_label_norm1 = gsub("written_press", "press", var_label_norm1),
-         var_label_norm1 = gsub("pers_influence", "personal-influence", var_label_norm1),
-         var_label_norm1 = gsub("job_situation", "job-situation", var_label_norm1),
-         var_label_norm1 = gsub("religious_inst$", "religious-institutions", var_label_norm1),
-         var_label_norm1 = gsub("polit_matters|political_matters", "political-matters", var_label_norm1),
-         var_label_norm1 = gsub("pol_discussion|polit_discussion|political_discussion", "political-discussion", var_label_norm1),
-         var_label_norm1 = gsub("nat_resources", "natural-resources", var_label_norm1),
-         var_label_norm1 = gsub("non_euro_zone", "non-euro-zone", var_label_norm1),
-         var_label_norm1 = gsub("party_attachment", "party-attachment", var_label_norm1),
-         var_label_norm1 = gsub("which_issue", "which-issue", var_label_norm1),
-         var_label_norm1 = gsub("economic_situation|economic_sit", "economic-situation", var_label_norm1),
-         var_label_norm1 = gsub("public_debt", "public-debt", var_label_norm1),
-         var_label_norm1 = gsub("left_right", "left-right", var_label_norm1),
-         var_label_norm1 = gsub("important_issues_cntry", "important-issues-cntry", var_label_norm1),
-         var_label_norm1 = gsub("important_issues|import_issues", "important-issues", var_label_norm1),
-         var_label_norm1 = gsub("important_values_eu", "important-values-eu", var_label_norm1),
-         var_label_norm1 = gsub("important_values_pers", "important-values-personal", var_label_norm1),
-         var_label_norm1 = gsub("important_life_domains", "important-life-domains", var_label_norm1),
-         var_label_norm1 = gsub("quality_of_life|life_quality", "life-in-general", var_label_norm1),
-         var_label_norm1 = gsub("household_composition", "household-composition", var_label_norm1),
-         var_label_norm1 = gsub("size_of_community", "size-of-community", var_label_norm1),
-         var_label_norm1 = gsub("head_of_hh", "head-of-hh", var_label_norm1),
-         var_label_norm1 = gsub("last_job", "last-job", var_label_norm1),
-         var_label_norm1 = gsub("level_in_society", "level-in-society", var_label_norm1),
-         var_label_norm1 = gsub("self_placement", "self-placement", var_label_norm1),
-         var_label_norm1 = gsub("social_class|soc_class", "social-class", var_label_norm1),
-         var_label_norm1 = gsub("education_level", "education-level", var_label_norm1),
-         var_label_norm1 = gsub("living_conditions", "living-conditions", var_label_norm1),
-         var_label_norm1 = gsub("_nat_", "_national_", var_label_norm1),
-         var_label_norm1 = gsub("^nat_", "national_", var_label_norm1),
-         var_label_norm1 = gsub("_nat$", "_national", var_label_norm1),
-         var_label_norm1 = gsub("europ_", "european_", var_label_norm1),
-         var_label_norm1 = gsub("12_months_|12_mo_", "12-months_", var_label_norm1),
-         var_label_norm1 = gsub("national_parliament", "national-parliament", var_label_norm1),
-         var_label_norm1 = gsub("national_government", "national-government", var_label_norm1),
-         var_label_norm1 = gsub("national_economy", "national-economy", var_label_norm1),
-         var_label_norm1 = gsub("good_bad", "good-bad", var_label_norm1),
-         var_label_norm1 = gsub("court_of_auditors", "court-of-auditors", var_label_norm1),
-         var_label_norm1 = gsub("council_of_ministers", "council-of-ministers", var_label_norm1),
+    var_label_norm1 = gsub(
+      "tv_channels", "tv-channels", var_label_norm1),
+    var_label_norm1 = gsub(
+      "radio_stations", "radio-stations", var_label_norm1),
+    var_label_norm1 = gsub(
+      "regional_local", "regional-local", var_label_norm1),
+    var_label_norm1 = gsub(
+      "air_conditioning", "air-conditioning", var_label_norm1),
+    var_label_norm1 = gsub(
+      "pol_party|political_party", "political-party",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "electoral_particip_", "electoral-participation_",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "electoral_particip$", "electoral-participation$",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "trust_in_institutions", "trust",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "justice_legal_system|justice_nat_legal_system",
+      "justice-system", var_label_norm1),
+    var_label_norm1 = gsub(
+      "non_govmnt_org|non_govnmt_org", "ngo", var_label_norm1),
+    var_label_norm1 = gsub(
+      "polit_parties|political_parties", "political-parties", var_label_norm1),
+    var_label_norm1 = gsub(
+      "reg_loc_public_authorities|reg_local_authorities|reg_loc_authorities|reg_local_public_authorities|rg_lc_authorities", "regional-local-authorities", var_label_norm1),
+    var_label_norm1 = gsub(
+      "written_press", "press", var_label_norm1),
+    var_label_norm1 = gsub(
+      "pers_influence", "personal-influence", var_label_norm1),
+    var_label_norm1 = gsub(
+      "job_situation", "job-situation", var_label_norm1),
+    var_label_norm1 = gsub(
+      "religious_inst$", "religious-institutions",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "polit_matters|political_matters", "political-matters", var_label_norm1),
+    var_label_norm1 = gsub(
+      "pol_discussion|polit_discussion|political_discussion", "political-discussion", var_label_norm1),
+    var_label_norm1 = gsub(
+      "nat_resources", "natural-resources", var_label_norm1),
+    var_label_norm1 = gsub(
+      "non_euro_zone", "non-euro-zone", var_label_norm1),
+    var_label_norm1 = gsub(
+      "party_attachment", "party-attachment", var_label_norm1),
+    var_label_norm1 = gsub(
+      "which_issue", "which-issue", var_label_norm1),
+    var_label_norm1 = gsub(
+      "economic_situation|economic_sit", "economic-situation", var_label_norm1),
+    var_label_norm1 = gsub(
+      "public_debt", "public-debt", var_label_norm1),
+    var_label_norm1 = gsub(
+      "left_right", "left-right", var_label_norm1),
+    var_label_norm1 = gsub(
+      "important_issues_cntry", "important-issues-cntry",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "important_issues|import_issues", "important-issues",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "important_values_eu", "important-values-eu", var_label_norm1),
+    var_label_norm1 = gsub(
+      "important_values_pers", "important-values-personal",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "important_life_domains", "important-life-domains",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "quality_of_life|life_quality", "life-in-general",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "household_composition", "household-composition",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "size_of_community", "size-of-community", var_label_norm1),
+    var_label_norm1 = gsub(
+      "head_of_hh", "head-of-hh", var_label_norm1),
+    var_label_norm1 = gsub(
+      "last_job", "last-job", var_label_norm1),
+    var_label_norm1 = gsub(
+      "level_in_society", "level-in-society", var_label_norm1),
+    var_label_norm1 = gsub(
+      "self_placement", "self-placement", var_label_norm1),
+    var_label_norm1 = gsub(
+      "social_class|soc_class", "social-class", var_label_norm1),
+    var_label_norm1 = gsub(
+      "education_level", "education-level", var_label_norm1),
+    var_label_norm1 = gsub(
+      "living_conditions", "living-conditions", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_nat_", "_national_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "^nat_", "national_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "_nat$", "_national", var_label_norm1),
+    var_label_norm1 = gsub(
+      "europ_", "european_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "12_months_|12_mo_", "12-months_", var_label_norm1),
+    var_label_norm1 = gsub(
+      "national_parliament", "national-parliament", var_label_norm1),
+    var_label_norm1 = gsub(
+      "national_government", "national-government", var_label_norm1),
+    var_label_norm1 = gsub(
+      "national_economy", "national-economy", var_label_norm1),
+    var_label_norm1 = gsub(
+      "good_bad", "good-bad", var_label_norm1),
+    var_label_norm1 = gsub(
+      "court_of_auditors", "court-of-auditors", var_label_norm1),
+    var_label_norm1 = gsub(
+      "council_of_ministers", "council-of-ministers",
+      var_label_norm1),
 
     # EU-related abbreviations and phrases
-         var_label_norm1 = gsub("euro_12", "euro-12", var_label_norm1),
-         var_label_norm1 = gsub("european_parliament|eu_parl", "european-parliament", var_label_norm1),
-         var_label_norm1 = gsub("european_union", "european-union", var_label_norm1),
-         var_label_norm1 = gsub("european_unification|europ_unification|europ_unif", "european-unification", var_label_norm1),
-         var_label_norm1 = gsub("european_economy|europ_economy", "european-economy", var_label_norm1),
-         var_label_norm1 = gsub("european_ombudsman", "european-ombudsman", var_label_norm1),
-         var_label_norm1 = gsub("european_elections|ep_election", "european-elections", var_label_norm1),
+    var_label_norm1 = gsub(
+      "euro_12", "euro-12", var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_parliament|eu_parl", "european-parliament",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_union", "european-union", var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_unification|europ_unification|europ_unif",
+      "european-unification", var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_economy|europ_economy", "european-economy",
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_ombudsman", "european-ombudsman", var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_elections|ep_election", "european-elections",
+      var_label_norm1),
     var_label_norm1 = gsub("european_elec|european_elect|europ_elec", "european-elections", var_label_norm1),
-         var_label_norm1 = gsub("european_economy", "european-economy", var_label_norm1),
-         var_label_norm1 = gsub("european_currency|europ_currency", "european-currency", var_label_norm1),
-         var_label_norm1 = gsub("europ_court_of_auditors|eur_court_of_auditors", "eu-court-of-auditors", var_label_norm1),
-         var_label_norm1 = gsub("europ_court_of_justice|european_court_of_justice", "eu-court-of-justice", var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_economy", "european-economy", var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_currency|europ_currency", "european-currency", 
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "europ_court_of_auditors|eur_court_of_auditors", "eu-court-of-auditors", 
+      var_label_norm1),
+    var_label_norm1 = gsub(
+      "europ_court_of_justice|european_court_of_justice", "eu-court-of-justice", 
+      var_label_norm1),
     var_label_norm1 = gsub("econ_and_social_committee|economic_and_soc_committee|econ_and_soc_committee", "econ-and-soc-committee", var_label_norm1),
-         var_label_norm1 = gsub("europ_polit_matter|europ_political_matters", "european-political-matters", var_label_norm1),
-         var_label_norm1 = gsub("euro_zone", "euro-zone", var_label_norm1),
-         var_label_norm1 = gsub("eu_issues", "eu-issues", var_label_norm1),
-         var_label_norm1 = gsub("eu_budget", "eu-budget", var_label_norm1),
-         var_label_norm1 = gsub("eu_statements", "eu-statements", var_label_norm1),
-         var_label_norm1 = gsub("eu_citizenship", "eu-citizenship", var_label_norm1),
-         var_label_norm1 = gsub("eu_membership|eu_membersh", "eu-membership", var_label_norm1),
-         var_label_norm1 = gsub("eu_image", "eu-image", var_label_norm1),
-         var_label_norm1 = gsub("eu_level", "eu-level", var_label_norm1),
-         var_label_norm1 = gsub("ecb|european_central_bank", "european-central-bank", var_label_norm1),
-         var_label_norm1 = gsub("european_commission", "european-commission", var_label_norm1),
-         var_label_norm1 = gsub("nms", "new-member-states", var_label_norm1),
-         var_label_norm1 = gsub("council_of_the_eu", "council-of-the-eu", var_label_norm1),
+    var_label_norm1 = gsub("europ_polit_matter|europ_political_matters", "european-political-matters", 
+                           var_label_norm1),
+    var_label_norm1 = gsub(
+      "euro_zone", "euro-zone", var_label_norm1),
+    var_label_norm1 = gsub(
+      "eu_issues", "eu-issues", var_label_norm1),
+    var_label_norm1 = gsub(
+      "eu_budget", "eu-budget", var_label_norm1),
+    var_label_norm1 = gsub(
+      "eu_statements", "eu-statements", var_label_norm1),
+    var_label_norm1 = gsub(
+      "eu_citizenship", "eu-citizenship", var_label_norm1),
+    var_label_norm1 = gsub(
+      "eu_membership|eu_membersh", "eu-membership", var_label_norm1),
+    var_label_norm1 = gsub(
+      "eu_image", "eu-image", var_label_norm1),
+    var_label_norm1 = gsub(
+      "eu_level", "eu-level", var_label_norm1),
+    var_label_norm1 = gsub(
+      "ecb|european_central_bank", "european-central-bank", var_label_norm1),
+    var_label_norm1 = gsub(
+      "european_commission", "european-commission", var_label_norm1),
+    var_label_norm1 = gsub(
+      "nms", "new-member-states", var_label_norm1),
+    var_label_norm1 = gsub(
+      "council_of_the_eu", "council-of-the-eu", var_label_norm1),
 
@@
 # list of country codes to be converted to country names
-eu_countries <- c("at", "be", "bg", "cy", "cz", "dk", "de", "ee", "es", "fr",
+eu_countries <- c(
+  "at", "be", "bg", "cy", "cz", "dk", "de", "ee", "es", "fr",
    "fi", "gb", "gr", "hr", "hu", "ie", "it", "lt", "lv", "lu",
@@
 # !!! the problem here is that dk sometimes means "don't know" and sometimes Demnark !!!
+# this needs to be treated in gesis_metadata_create, where the
+# user defined missings are separately analyzed, they dk should 
+# be converted there to do_not_know.
 
@@
   # tokenize
-  unnest_tokens(word, text, token = "regex", pattern = " |  ") %>%
+  unnest_tokens(word, text, 
+                token = "regex", pattern = " |  ") %>%
   # eliminate stop words, with some exceptions
-  anti_join(stop_words %>% filter(!word %in% c("working", "right", "other"))) %>%
+  anti_join(stop_words %>%
+              filter(!word %in% c("working", "right", "other")
+                     )) %>%
   # eliminate numbers
@@
   # convert country codes to country names
-  mutate(country = ifelse(word %in% eu_countries, word, NA),
-         country = countrycode::countrycode(toupper(country), "iso2c", "country.name"),
+  mutate(
+    country = ifelse(word %in% eu_countries, word, NA),
+         country = countrycode::countrycode(
+           toupper(country), "iso2c", "country.name"),
     country = tolower(country),
@@
     word = ifelse(word == "united kingdom", "united-kingdom", word),
-         word = ifelse(word == "tcc", "turkish-cypriot-community", word)) %>%
+    word = ifelse(word == "tcc", "turkish-cypriot-community", word)
+    ) %>%
   select(-country)
-
+```
+```{r, eval=FALSE}
 head(tidy)
 ```
+```{r, echo=FALSE, results='asis'}
+head(tidy) %>%
+  kable () %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 ) %>%
+  column_spec(1, width =  "4em") %>%
+  column_spec(2:5, width = "3em") %>%
+  column_spec(6, width = "3em", bold = T, border_left = T)
 
-```{r}
+```
+
+```{r, eval=FALSE}
 tidy %>%
@@
   arrange(desc(n)) %>% 
-  print(n = 50)
+  head(., 50)
+```
+
+```{r, echo=FALSE}
+kable(
+  tidy %>%
+  count(word) %>%
+  arrange(desc(n)) %>%
+  head(., 50)
+  ) %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 12 ) %>%
+  column_spec(1, width =  "5em", bold = T) %>%
+  column_spec(2, width = "3em")
 ```
@@
   spread(word, n)
-
+```
+```{r, eval=FALSE}
 dtm[1100:1110, c(1,2,110:112)]
+```
+```{r, echo=FALSE, results='asis'}
+dtm[1100:1110, c(1,2,110:112)] %>%
+  kable () %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 )
 ```
@@
 order <- dtm %>%
-  summarise_each(funs(sum(is.na(.))/length(.))) %>%
+  summarise_all(funs(sum(is.na(.))/length(.))) %>%
   t() %>% as.vector()
@@
 ```{r}
-dtm2 <- dtm %>% select(names(sort(order))) %>%
-  mutate_at(vars(3:ncol(dtm)), funs(ifelse(is.na(.), "", .))) %>%
+dtm2 <- dtm %>%
+  select( names(sort(order)) ) %>%
+  mutate_at(vars(3:ncol(dtm)), 
+            funs(ifelse(is.na(.), "", .))
+            ) %>%
   unite("var_label_std", 3:ncol(dtm)) %>%
-  mutate(var_label_std = gsub("(\\_)\\1+", "\\1", var_label_std)) %>%
-  mutate(var_label_std = gsub("^_|_$", "", var_label_std))
+  mutate(var_label_std = gsub(
+    "(\\_)\\1+", "\\1", var_label_std) ) %>%
+  mutate(var_label_std = gsub(
+    "^_|_$", "", var_label_std) )
+```
 
+```{r}
 dtm2 %>% 
-  print(n = 50)
+  print(n = 50) %>%
+  kable() %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 12 ) %>%
+  column_spec(1, width = "5em", bold = T) %>%
+  column_spec(2, width = "3em")
 ```
@@
   count(var_label_std) %>%
-  # exclude nationality, country (nation) codes, size of community (country-specific), language of interview
-  filter(!grepl("nationality|nation_|size-of-community|language_interview", var_label_std)) %>%
+  # exclude nationality, country (nation) codes, 
+  # size of community (country-specific), language of interview
+  filter(
+    !grepl("nationality|nation_|size-of-community|language_interview",
+           var_label_std)) %>%
   filter(n > 5) %>%
@@
 
-```{r}
+```{r, eval=FALSE}
 # filter on var_label_std
@@
   print(n = 50)
+```
+```{r, echo=FALSE}
+dtm2 %>%
+  filter(grepl("trust|satisf|left-right|interest|class", var_label_std)) %>% 
+  count(var_label_std) %>%
+  arrange(desc(n)) %>%
+  head(., n = 50) %>%
+  kable () %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 12 ) %>%
+  column_spec(1, width =  "8em", bold = T ) %>%
+  column_spec(2, width = "3em")
+
 ```
@@
 
-```{r}
+```{r, eval=FALSE}
 # search document-term matrix by keyword
@@
 ```
+
+```{r, echo=FALSE, results='asis'}
+# search document-term matrix by keyword
+dtm %>%
+  filter(!is.na(`left-right`)) %>%
+  janitor::remove_empty(c("rows", "cols")) %>% 
+  print(n = 100) %>%
+  kable() %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 ) %>%
+  column_spec(1,  bold = T, width = "4em") %>%
+  column_spec(3:4, width = "2em") %>%
+  column_spec(5:6, width = "4em")
+```

### 2020-07-03 R/data-eurobarometer_dtm.R

@@ -1,0 +1,11 @@
+#' Eurobarometer DTM
+#'
+#' Please give a bit more information about this.
+#'
+#' @format A data frame with 60671 rows in 3  variables:
+#' \describe{
+#'   \item{var_label_norm1}{please describe}
+#'   \item{var_id}{please describe}
+#'   \item{var_label_std}{please describe}
+#'   }
+"eurobarometer_dtm"

### 2020-07-03 data-raw/create_eurobarometer_dtm.R

@@ -1,0 +1,6 @@
+
+# you can have a copy of the code here.
+# I chose to save dtm2 (10.6 mbyte), because dtm is too big (1421.365 mbyte).
+
+eurobarometer_dtm <- dtm2
+usethis::use_data(eurobarometer_dtm, internal = F, overwrite =T )

### 2020-07-06 data-raw/article.Rmd
article , etc

---
title: "Article"
author: "Daniel Antal, Marta Kolczynska"
date: "7/6/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The increasing availability of microdata available as open data creates new opportunities for ex-post harmonization and joint exploitation of these resources. We introduce the surveyharmonization and eurobarometer packages that help the reproducible ex-post harmonization of survey data stored in various, individual files with rich metadata.  Our examples are related to joining data from SPSS files of various Eurobarometer survey responses spanning over 25 years, but our tools are useful to make this work.

The Eurobarometer case is particularly interesting, because each wave of the Eurobarometer survey is n fact a combination of 27-30 nationally representative survey conducted in about 20 languages based on a generic English and French questionnaire. They contain trend variables, which are asked repeatedly in many waves, so only joining the same variables of two Eurobarometer waves, for example, on attitudes to trust in institutions, requires a harmonization of about 60 complex data assets.  Much of this harmonization work is done ex ante, and it is mainly outside of the scope of our interest.  While conducting the fieldwork of these surveys and processing the interview data is fully harmonized within each wave, there are various ex ante harmonization solutions are applied in various waves, especially ones that took place far away in time, for example, in 1995 and 2019. When we are ex post joining data from different survey waves, we have to understand and partly reconstruct ex ante harmonization steps.

Ex-post (or retrospective) data harmonization refers to procedures applied to already collected data to improve the comparability and inferential equivalence of measures collected by different studies. In practice, ex-post harmonization means a lot of recoding, i.e. to make sure that codes are understood exactly the same way across all data resources.

Reproducible recoding can be in form of code in some programming language, where running the code performs all the data transformations. When the number of recodes is large, the code quickly becomes long and complicated and difficult to follow for anyone but the code’s creator. We tried to create a workflow that is fully reproducible and automatically documented.

In simple words, when we are creating a data panel from two (or more) Eurobarometer surveys stored in separate SPSS files, we are aiming to solve the following problems:

--	[Importing data](#import) from proprietary statistical software and creating a faithful, practical representation of their data and metadata contents for later harmonization and join.

-- [Ex-post harmonization](#expost)

* Define a common [taxonomy](#taxonomy), probably based on a thesaurus for consistentcy.

* [harmonizing variable names](#harmonize-var-names), or “variable labels” in an SPSS context, making sure that variables that represent exactly the same information have exactly the same, programmatically usable variable names; dissimilar variables have dissimilar names.

* [Harmonizing value labels](#harmonize-labels) of categorical variables, making sure that the same labels, such as ‘good’, ‘mediocre’, ‘bad’ always have the same representation in our new data object.

* Harmonizing the treatment of [missing values](#harmonize-missing), carefully documenting the differences of ‘missingness’ (inapplicable items, items that were not collected, etc.), because during analysis they may require different interpretation.

--	Creating a [joint data table](#datajoin) where each variable that represents the same class and metadata attributes; documenting all the changes and 	Providing methods that re-export the data files to SPSS, Stata, SAS.

* Creating a consistent, [unique identifier](#uniqid) for each observation.

* Create a full [documentation](#documentation) of conversions, recoding for reproducability.

* Provide methods to export the harmonized, joined data to proprietary formats for [analysis outside R](#exportmethods)

* If the user wishes to prepare reproducible analysis in R, provide [methods](#rmethods) to convert the data to base types of numeric and factor, the usual inputs of statistical functions.

Our workflow generally fits into the tidyverse.

## Import files {#import}
Complex microdata, such as primary survey data, is often stored in some statistical file format. In our article, we use examples for SPSS’s .sav format, but similar problems are present with other proprietary file formats.  They usually do not only represent the data, but rich metadata that provides guidance for the analyst who wants to use them – for example, which observations should be treated as missing, or what is the valid range of a variable.

In our case, we are dealing with highly structured survey data that contain a lot of metadata, partly about the ex ante harmonization process that is vital for using the data.  For example, an archived Eurobarometer wave file contains recently more than 28 subsamples (for each EU member state, two for Germany, two for the former member state United Kingdom, and often for EEA or prospective member state countries.)  The way these subsamples were merged together and handled must be preserved.

We fully rely in importing on two packages, [haven](https://haven.tidyverse.org/) and [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) from the tidyverse.  They have created classes and methods that give a faithful representation of imported statistical software files, but they need to be amended for reproducibility and harmonization information for later joins.  These steps are likely to be present when a researcher is joining any two statistical software files, not only survey microdata files.

## Ex-post harmonization {#expost}

> maybe you code write a bit about ex-post harmonizaton here. This can also lead out to your other work or to a later article about describing a trend file that we create.

### Thesaurus, taxonomy {#taxonomy}

### Harmonization of variable names {#harmonize-var-names}

We should keep an eye on [snakecase](#https://github.com/Tazinho/snakecase) and maybe janitor, too.

### Harmonization of categories {#harmonize-labels}

We should keep an eye on [forcats](#https://forcats.tidyverse.org/)

### Harmonization of missing values {#harmonize-missing}

## Joining the data {#datajoin}

At this stage, we are ready to have rich data tables with many metadata attributes that are ready to be joined in R. We have kept all information that is necessary for the later step, i.e. the actual statistical analysis of the data in a way that it still can be done in SPSS, Stata, or in R.

* attributes are harmonized and joined
* data is harmonized and join

At this point the new data table is ready for analysis.  We provide some methods for exporting them (back) to SPSS or Stata, or flat tables, before turning to analyzing the data in R.

<this should be more data scientific >

### Unique identification {#uniqid}

### Documenting the process {#documentation}

<probably this should be a heavier part>

### Methods for exporting {#exportmethods}
<just simple description and code links>

* exporting to SPSS
* exporting to Stata
* exporting to flat files

### Methods for analyzing in R {#rmethods}

<just simple description and code links>

### 2020-07-06 data-raw/working_with_stringi.R

@@ -1,0 +1,32 @@
+library(stringi)
+
+replacements_table <- data.frame(
+  pattern = c("%", "europ|eu", "dk"),
+  x = c("_pct", "european", "do_not_know")
+)
+
+
+replacement_vector <- c("^cap_", "^[a-z]{1,2}_", "czech_republic|czech_rep")
+names (replacement_vector) <- c("common-agricultural-policy_", "", "chechia")
+
+str_replace_all(s = c("europ-council", "dk",
+                      "100%", "less than 25%", "eu parliament"),
+                replacement_vector)
+
+
+stri_replace_all_fixed("The quicker brown fox jumped over the lazy dog.",
+                       c("quick", "brown", "fox"),
+                       c("slow",  "black", "bear"), vectorize_all=FALSE)
+
+
+stri_replace_all_regex("The quicker brown fox jumped over the lazy dog.",
+                       "\\b"%s+%c("quick", "brown", "fox")%s+%"\\b", c("slow",  "black", "bear"),
+                       vectorize_all=FALSE)
+
+
+
+stri_replace_all_regex(tolower(c("D12_cap_in_czech_republic",
+                                 "cap_in_czech_rep")),
+                       pattern = replacement_table$pattern,
+                       replacement = replacement_table$replacement,
+                       vectorize_all=FALSE)

### 2020-07-06 tests/testthat/test-utils.R

@@ -13,3 +13,5 @@ test_that("to_survey_list handles exceptions", {
 ## Test class suggest ----------------------------------------------
 
 #metadata <- gesis_metadata_create(ZA5913_sample)
+
+

### 2020-07-10 README.Rmd
transition to 0.1.8 and moving functionality to retroharmonize

@@
 The goal of `eurobarometer` is converting Eurobarometer microdata files, as stored by GESIS, into tidy R data frames and help common pre-processing problems.
+The more generic function for retrospecitve (ex post) survey harmonization can be found in the package [retroharmonize](http://retroharmonize.satellitereport.com/).
 
@@
 
-5.
-The [eurobarometer_labelled](http://eurobarometer.danielantal.eu/articles/eurobarometer_class.html) class comes with simple methods that can convert all variables to the base R types of numeric or factor, and they are ready to be used with any R statistical packages and tidyverse data processing tools.
-Alternatively, they can be exported back to [SPSS](https://haven.tidyverse.org/reference/read_spss.html) or [Stata](https://haven.tidyverse.org/reference/read_dta.html) files with the functions of the [haven](https://haven.tidyverse.org/) package.
-
 ## Joining With Eurostat & Other Data Tables

### 2020-07-10 R/data-vocabulary_nuts1.R

@@ -1,10 +1,0 @@
-#' Vocabulary of NUTS1 regions.
-#'
-#' @format A data frame with 49 observations and four variables:
-#' \describe{
-#'   \item{country_code}{ISO 3166 country codes}
-#'   \item{region_nuts_names}{Region names in Eurobarometer files.}
-#'   \item{code10}{NUTS2010 regional codes}
-#'   \item{code13}{NUTS2013 regional codes}
-#'   }
-"vocabulary_nuts1"

### 2020-07-10 R/eurobarometer_labelled.R

@@ -1,95 +1,0 @@
-#' Construct A Variable Of Type 'eurobarometer_labelled'
-#'
-#' @param x A vector of numeric values of class integer, numeric,
-#' haven_labelled or haven_labelled_spss.
-#' @param value_labels The \code{labels} attribute, defaults to \code{NULL}.
-#' @param na_values The \code{na_values} attribute, defaults to \code{NULL}.
-#' @param na_range The \code{na_range} attribute, defaults to \code{NULL}.
-#' @param labels_orig The \code{labels_orig} attribute which contains the original
-#' value_labels before harmonization or normalization, defaults to \code{NULL}.
-#' @param var_name_orig The \code{var_name_orig} attribute, which
-#' contains the original variable name,  defaults to \code{NULL}.
-#' @param qb The \code{qb} attribute, which contains the identifier of the
-#' question block, defaults to \code{NULL}.
-#' @param conversion The \code{conversion} attribute which contains the
-#' name of the function that created the variable, defaults to \code{NULL}.
-#' @return A vector with metadata attributes of class \code{eurobarometer_labelled}.
-#' @examples
-#' eurobarometer_labelled(x = c(6,7,8,9,6,7,8),
-#'    labels =c("HI" = 6, "HELLO" = 7,
-#'              "HOWDY" = 8, "BYE" = 9),
-#'    na_values = 9,
-#'    qb = "Example Block",
-#'    conversion = "eurobarometer_labelled"
-#' )
-#' @export
-
-eurobarometer_labelled <- function ( x = NULL,
-                                     value_labels = NULL,
-                                     na_values = NULL,
-                                     na_range = NULL,
-                                     labels_orig = NULL,
-                                     var_name_orig = NULL,
-                                     qb = NULL,
-                                     conversion = NULL) {
-
-  if ( is.null(x) ) {
-    stop ( "Parameter x must be given.")
-  }
-
-  if (class(x)[1] =='logical' ) {
-    warning ("class <logical> is converted to <integer>.")
-    x <- as.integer(x)
-  }
-  if ( any( class(x) %in% c("integer", "numeric")) ) {
-    ebl <- haven::labelled_spss( x = x,
-                          labels = value_labels,
-                          na_values = na_values,
-                          na_range = na_range )
-  } else if ( any(class(x) %in% "haven_labelled_spss") ) {
-    ebl <- x
[…]

### 2020-07-10 R/panel_create.R

@@ -1,39 +1,0 @@
-#' @title Create A Skeleton Survey
-#'
-#' @param survey_list A list of data frames containing surveys.
-#' @param id_vars A vector of ID variables to form a panel ID var,
-#' defaults to \code{c("uniqid", "doi")}.
-#' @return A tibble with the number of original ID vars and their
-#' \code{panel_id} as their concatenation for several surveys.
-#' @family harmonization functions
-#' @examples
-#' data (ZA7576_sample)
-#' data (ZA6863_sample)
-#' data (ZA5913_sample)
-#'
-#' survey_list <- list ( 'ZA7576_sample' = ZA7576_sample,
-#'                       'ZA6863_sample' = ZA6863_sample,
-#'                       'ZA5913_sample' = ZA5913_sample)
-#'
-#' panel_create (survey_list, id_vars =c("uniqid", "doi") )
-#' @export
-
-
-panel_create <- function (survey_list,
-                          id_vars = c("uniqid", "doi") ) {
-
-  input_id_vars <- id_vars
-  panel_id <- NULL
-  id_vars <- NULL
-
-  # in case of unexpected results, after fixing add a unique test
-  # to tests/testthat/test-panel_create.R
-
-  results_id <- lapply (
-    survey_list,
-    function(x) id_create (dat = x, id_vars = input_id_vars)
-    )
-
-  do.call(rbind,results_id )
-
-}

### 2020-07-10 tests/testthat/test-eurobarometer_labelled.R

@@ -1,53 +1,0 @@
-library(testthat)
-
-### Test wrong parameter classes ---------------------------------
-
-# "Hello, Hi, Hey":
-# https://open.spotify.com/track/35J5fWBYFx1LTut0JwXxgh?si=pspMB7AfT2G808_wP3Ysbw
-
-test_that("Appropriate warnings are given", {
-  expect_warning(eurobarometer_labelled (x = c(TRUE, TRUE, FALSE, NA))  )
-})
-
-test_that("Errors are found", {
-  expect_error(eurobarometer_labelled ( x = data.frame ( x = 1:5)))
-})
-
-### Test if the constructor works as expected -----------
-v <- labelled::labelled(c(3,4,4,3,8, 9),
-                        c(YES = 3, NO = 4, `WRONG LABEL` = 8, REFUSED = 9)
-)
-correct_labels <- c(3,4,8,9)
-names(correct_labels) <- c("YES", "NO", "WRONG LABEL", "REFUSED")
-correct_yes_no <- c(1,0)
-names(correct_yes_no) <- c("yes", "no")
-
-
-eurobarometer_labelled(x = c(6,7,8,9, 6,7,8),
-                       labels =c("HI" = 6, "HELLO" = 7,
-                                 "HOWDY" = 8, "BYE" = 9),
-                       na_values = 9,
-                       qb = "Example Block",
-                       conversion = "eurobarometer_labelled")
-
-x <-haven::labelled_spss(
-  x = c(1,2,2,1,99998,99999),
-  labels =  c(yes = 1, no = 2, wrong_label = 99998, refused = 99999),
-  na_values = NULL,
-  na_range = 99998:99999,
-  label = "Variable Example"
-)
-
-test_that("constructor works", {
-  expect_equal(labelled::val_labels ( eurobarometer_labelled(v)),
-               correct_labels)
-  expect_equal(labelled::val_labels (
-    eurobarometer_labelled(x = c(0,1),
-                           value_labels = c( 'yes' = 1, 'no' = 0)
-                           )),
-    correct_yes_no)
-  expect_equal(labelled::na_range (eurobarometer_labelled(x)),
-               c(99998,99999))
-  expect_equal(class ( eurobarometer_labelled(x)),
-               c("eurobarometer_labelled", class(x)))
-})

### 2020-07-10 README.Rmd
remove "

@@
 [![Project Status: Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
-"[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)"
-"[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/eurobarometer)](https://cran.r-project.org/package=eurobarometer)"
+[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
+[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/eurobarometer)](https://cran.r-project.org/package=eurobarometer)
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3901724.svg)](https://doi.org/10.5281/zenodo.3901724)

### 2020-07-10 data-raw/article.Rmd
Harmonization description, edits, references to article draft

@@
 ---
-title: "Article"
+title: "Tools for reproducible ex-post survey data harmonization"
+subtitle: "The case of the Eurobarometer surveys"
 author: "Daniel Antal, Marta Kolczynska"
 date: "7/6/2020"
+bibliography: library.bib
+biblio-style: apalike
 output:
@@
 
-The increasing availability of microdata available as open data creates new opportunities for ex-post harmonization and joint exploitation of these resources.
+The increasing availability of microdata creates new opportunities for joint exploitation of these resources for comparative research.
+Before joint analysis is possible, the data need to be harmonized, i.e.
+transformed and combined in a single data set in a way that matches the measurement of the same characteristics and standardizes the coding of responses.
 We introduce the surveyharmonization and eurobarometer packages that help the reproducible ex-post harmonization of survey data stored in various, individual files with rich metadata.
-Our examples are related to joining data from SPSS files of various Eurobarometer survey responses spanning over 25 years, but our tools are useful to make this work.
+We illustrate the proposed workflow using the case of the Eurobarometer surveys, which provide rich data on European publics in form of single-wave SPSS files from a period of over 25 years.
+Our tools are flexible and can be applied to other survey projects.
 
-The Eurobarometer case is particularly interesting, because each wave of the Eurobarometer survey is n fact a combination of 27-30 nationally representative survey conducted in about 20 languages based on a generic English and French questionnaire.
-They contain trend variables, which are asked repeatedly in many waves, so only joining the same variables of two Eurobarometer waves, for example, on attitudes to trust in institutions, requires a harmonization of about 60 complex data assets.
-Much of this harmonization work is done ex ante, and it is mainly outside of the scope of our interest.
-While conducting the fieldwork of these surveys and processing the interview data is fully harmonized within each wave, there are various ex ante harmonization solutions are applied in various waves, especially ones that took place far away in time, for example, in 1995 and 2019.
+The Eurobarometer case is particularly interesting, because each wave of the Eurobarometer project includes 27-30 nationally representative surveys conducted in about 20 languages based on a generic English and French questionnaire.
+They contain trend variables, which are asked repeatedly in many waves, so only joining the same variables of two Eurobarometer waves, for example, on trust in institutions, requires a harmonization of about 60 complex data assets.
+Within each wave, the surveys are harmonized ex ante, including fieldwork organization or processing of the interview data and combining them into a single data file.
+These procedures differ between waves, especially ones that took place far away in time, for example, in 1995 and 2019.
 When we are ex post joining data from different survey waves, we have to understand and partly reconstruct ex ante harmonization steps.
 
-Ex-post (or retrospective) data harmonization refers to procedures applied to already collected data to improve the comparability and inferential equivalence of measures collected by different studies.
-In practice, ex-post harmonization means a lot of recoding, i.e.
-to make sure that codes are understood exactly the same way across all data resources.
+Ex-post data harmonization refers to procedures applied to existing data that have not been created in a way that ensures comparability, in a way that makes the resulting, harmonized, data suitable for joint analysis.
+In practice, ex-post harmonization means taking data as they were published by research projects and recording them to ensure that codes have the same meaning across all data sources.
 
-Reproducible recoding can be in form of code in some programming language, where running the code performs all the data transformations.
-When the number of recodes is large, the code quickly becomes long and complicated and difficult to follow for anyone but the code’s creator.
-We tried to create a workflow that is fully reproducible and automatically documented.
+In order for the harmonization to be reproducible, all data transformations and decisions must be documented with procedures that enable the reproduction of the end result.
+Reproducible recoding can be in form of code in some programming language, where running the code performs all the necessary data transformations.
+As the number of recodes and/or different source data sets increases, the code quickly becomes long and complicated and difficult to follow for anyone but the code’s creator.
+Instead, the reproducible workflow we propose has been designed with readability in mind, with documentation built into all steps and all procedures.
 
-In simple words, when we are creating a data panel from two (or more) Eurobarometer surveys stored in separate SPSS files, we are aiming to solve the following problems:
+In simple words, when we are creating a multi-wave integrated data set from two (or more) Eurobarometer surveys stored in separate SPSS files, we are aiming to solve the following problems:
 
---	[Importing data](#import) from proprietary statistical software and creating a faithful, practical representation of their data and metadata contents for later harmonization and join.
+--	[Importing data](#import) from proprietary statistical software formats and creating a faithful, practical representation of the data and metadata contents for later harmonization and join.
 
@@
 
-* [harmonizing variable names](#harmonize-var-names), or “variable labels” in an SPSS context, making sure that variables that represent exactly the same information have exactly the same, programmatically usable variable names; dissimilar variables have dissimilar names.
+* [harmonizing variable names](#harmonize-var-names), or “variable labels” in an SPSS context, making sure that variables that represent exactly the same information (in terms of question content and response measurement) have exactly the same, programmatically usable variable names; dissimilar variables have dissimilar names.
 
@@
 
-* Create a full [documentation](#documentation) of conversions, recoding for reproducability.
+* Creating a full [documentation](#documentation) of conversions, recoding for reproducability.
 
-* Provide methods to export the harmonized, joined data to proprietary formats for [analysis outside R](#exportmethods)
+* Providing methods to export the harmonized data set to proprietary formats for [analysis outside R](#exportmethods).
 
-* If the user wishes to prepare reproducible analysis in R, provide [methods](#rmethods) to convert the data to base types of numeric and factor, the usual inputs of statistical functions.
+* Providing [methods](#rmethods) to convert the data to base types of numeric and factor, the usual inputs of statistical functions, for reproducible analysis in R.
 
@@
 
+# The Eurobarometer surveys
+
+<Here more information about the EUrobarometer, the contents, frequency of waves, amount of data, fieldwork and number of respondents.
+Maybe even the paragraph from the introduction would fit better here than there>
+
 ## Import files {#import}
-Complex microdata, such as primary survey data, is often stored in some statistical file format.
-In our article, we use examples for SPSS’s .sav format, but similar problems are present with other proprietary file formats.
-They usually do not only represent the data, but rich metadata that provides guidance for the analyst who wants to use them – for example, which observations should be treated as missing, or what is the valid range of a variable.
+
+Complex microdata, such as primary survey data, is often stored in a format used by statistical software.
+In our article, we use examples for SPSS’s .sav format, but other proprietary file formats pose similar challenges.
+These formats usually do not only represent the data, but also contain rich metadata that provide additional information to the researcher, such as which observations should be treated as missing, or what is the valid range of a variable's values.
 
 In our case, we are dealing with highly structured survey data that contain a lot of metadata, partly about the ex ante harmonization process that is vital for using the data.
-For example, an archived Eurobarometer wave file contains recently more than 28 subsamples (for each EU member state, two for Germany, two for the former member state United Kingdom, and often for EEA or prospective member state countries.)  The way these subsamples were merged together and handled must be preserved.
+For example, an archived Eurobarometer wave file contains recently more than 28 national samples (for each EU member state, plus two for Germany - East and West, two for the former member state United Kingdom - for Great Britain and Northern Ireland, and often for European Economic Area or candidate countries.)  The way these subsamples were merged together and handled must be preserved.
 
@@
 
-> maybe you code write a bit about ex-post harmonizaton here. This can also lead out to your other work or to a later article about describing a trend file that we create.
+Ex-post (or retrospective) data harmonization refers to procedures applied to existing data to improve the comparability and inferential equivalence of measures collected by different studies [cf.
+@Fortier2017a].
+Ex-post survey data harmonization involves applying these procedures to survey data sets that were not created with comparability in mind or are otherwise not compatible enough to be analyzed as a single data source.
+
+Ex-post harmonization is both theory-informed and data-driven.
+Theory provides concepts and definitions, while data availability to a large extent determines what ends up being harmonized and how.
+Additionally, methodological research provides evidence about limits of ex-post harmonization with regard to achieving data comparability.
+
+While the exact steps vary by application case, a survey data harmonization project can be generally described as consisting of the following steps: (1) definition of the concepts of interest, (2) indentifying and obtaining the data, (3) data transformations, and (4) verification and documentation [cf.
+@Granda2016; @Wolf2016; @Fortier2017a; @Slomczynski2018; @Kolczynska2020a].
+It is worth emphasizing that "[R]etrospective harmonization of individual participant data is an iterative process composed of a series of key closely related, inter-dependant and often integrated steps" [@Fortier2017b: 1].
+
+The scope and complexity of the steps involved in survey data harmonization depends on the extent to which the existing data are already comparable with regard to the (a) sample and (b) measurement.
+The comparability of survey samples includes the definition of the target population, the sampling frame and sample type, as well as the response (or other outcome) rate.
+Measurement deals with the content and design of the survey questions, as well as the recording responses, ata entry and data validation procedures applied following fieldwork.
+The more similar the procedures in both areas, the higher are the chances for a high quality end product.
+
+In the case of the Eurobarometer, the harmonization process in greatly facilitated by the fact that all Eurobarometer waves are part of the same project with largely standardized procedures for many stages of the survey process, including questionnaire design, sampling, fieldwork, data processing and dissemination.
+Additionally, within each Eurobarometer wave, data for all countries are harmonized ex-ante.
+Thus, in this case the harmonization of the data across multiple Eurobarometer surveys boils down to data data transformation and documentation.
+For the purposes of presenting the process of harmonization with surveyharmonization and eurobarometer packages, we use the case of the Eurobarometer without the loss of generality of the described procedures, since it includes all data processing steps involved in harmonization.
 
@@
 <just simple description and code links>
+
+# References

### 2020-07-10 data-raw/article.Rmd
article update

@@
 date: "7/6/2020"
-bibliography: library.bib
+bibliography: [library.bib, packages.bib]
 biblio-style: apalike
 output:
+  html_document: default
   word_document: default
-  html_document: default
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::write_bib(c(
+  .packages(), 'haven', 'labelled', 'vctrs'
+), 'packages.bib')
 ```
+
+[Read the current knitted version on Rpubs](https://rpubs.com/antaldaniel/retroharmonize)
 
@@
 
-In our case, we are dealing with highly structured survey data that contain a lot of metadata, partly about the ex ante harmonization process that is vital for using the data.
-For example, an archived Eurobarometer wave file contains recently more than 28 national samples (for each EU member state, plus two for Germany - East and West, two for the former member state United Kingdom - for Great Britain and Northern Ireland, and often for European Economic Area or candidate countries.)  The way these subsamples were merged together and handled must be preserved.
+In our case, we are dealing with highly structured survey data that contain a lot of metadata, partly about the *ex ante* harmonization process that is vital for using the data.
+For example, an archived Eurobarometer wave file from recent years contains more than 28 national samples (for each EU member state, plus two for Germany - East and West, two for the former member state United Kingdom - for Great Britain and Northern Ireland, and often for European Economic Area or candidate countries, such as North Macedonia or Albania.)  The way these subsamples were merged together and handled must be preserved.
 
-We fully rely in importing on two packages, [haven](https://haven.tidyverse.org/) and [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) from the tidyverse.
+We almost fully rely in importing on two packages, [haven](https://haven.tidyverse.org/) and [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) from the tidyverse [cf.
+@R-haven; @R-labelled; @R-base].
 They have created classes and methods that give a faithful representation of imported statistical software files, but they need to be amended for reproducibility and harmonization information for later joins.
-These steps are likely to be present when a researcher is joining any two statistical software files, not only survey microdata files.
+Our version of the [retroharmonize::read_spss()](http://retroharmonize.satellitereport.com/reference/read_spss.html) function records the parameters of the [haven::read_spss()](https://haven.tidyverse.org/reference/index.html) call for reproducibility.
 
@@
 The comparability of survey samples includes the definition of the target population, the sampling frame and sample type, as well as the response (or other outcome) rate.
-Measurement deals with the content and design of the survey questions, as well as the recording responses, ata entry and data validation procedures applied following fieldwork.
+Measurement deals with the content and design of the survey questions, as well as the recording responses, data entry and data validation procedures applied following fieldwork.
 The more similar the procedures in both areas, the higher are the chances for a high quality end product.
@@
 In the case of the Eurobarometer, the harmonization process in greatly facilitated by the fact that all Eurobarometer waves are part of the same project with largely standardized procedures for many stages of the survey process, including questionnaire design, sampling, fieldwork, data processing and dissemination.
-Additionally, within each Eurobarometer wave, data for all countries are harmonized ex-ante.
+Additionally, within each Eurobarometer wave, data for all countries are harmonized *ex ante*.
 Thus, in this case the harmonization of the data across multiple Eurobarometer surveys boils down to data data transformation and documentation.
@@
 
+### Harmonization of missing values {#harmonize-missing}
+
+The harmonization of the value labels is a three-step process in our case.
+Because we work microdata that is not native R data, we must make sure that we have an  unambiguous type casting of the variables that we want to join, we apply the same range of category or value labels, and we consistently treat missing values.
+
+We start with the missing values, because they are treated differently in our main importing format, SPSS, and the R language.
+In SPSS, the user can defined certain values (and labels) to be treated as if they were missing, for example, in the calculation of an arithmetic mean or a median --- for example, `do not know`  answers.
+
+The haven package has already adopted the importing of user-defined missing values, or the user defined missing value range with the creation of the class [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html).
+While this class generally works well with the older [haven_labelled](https://haven.tidyverse.org/reference/labelled.html) methods, it does not have many implemented methods in haven.
+In our case, we had to make sure that when we concatenate two vectors with user-defined missing values or ranges, they will remain consistent after the join.
+Our class relies on the [vctrs](https://vctrs.r-lib.org/) package, which contains helper methods for creating new vector classes [cf.
+@R-vctrs], and which serves as the foundation of haven's own class system, too.
+
+Our choice in this case is the that we apply integer values, long out of the range of categorical integer values in the range `99900:99999` for a consistent numeric coding of the values to be treated as missing, and we label them consistently.
+The original missing labels and missing values as preserved as an attribute in our [labelled_spss_survey](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) class.
+
+The consistent, harmonized treatment of the missing values allows us to easily implement the [as_numeric](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html), [as_character - *documentation missing*], and [as_factor](http://retroharmonize.satellitereport.com/reference/as_factor.html) methods, by replacing the user-defined missing values with type-specific NA values in the R language (NA_real_ or NA_character_).
+After this conversion, R artithmetic and statistical functions can be applied without compromising reproducability --- all metadata is present for a consistent export to SPSS or STATA, or to document, or even revert to an earlier labelling.
+
 ### Harmonization of categories {#harmonize-labels}
 
-We should keep an eye on [forcats](#https://forcats.tidyverse.org/)
+The function  [harmonize_values()](http://retroharmonize.satellitereport.com/reference/harmonize_values.html) does three critically important steps:
 
-### Harmonization of missing values {#harmonize-missing}
+* changes the numeric (integer) codes of a survey to a pre-defined format, for example, coding all `yes` answers to 1 and all `no` answers to 0.
+
+* applies a consistent labelling, for example, uniformly applying `no` instead of `NO` and `not`.
+
+* as mentioned earlier, removes the user-defined missing values from the category integer range and uses a pre-defined missing labelling.
+
+Working with [harmonize_values()](http://retroharmonize.satellitereport.com/reference/harmonize_values.html) is relatively easy if the researcher wants to join a few variables in a few survey files, but for large-scale, programmatic use it not really suitable.
+
+We create various helper functions that allows a massive relabelling based on pre-defined harmonized value labels. [... later ....]
+
+The `retroharmonize` package only deals with a simple pre-processing step, i.e.
+a soft normalization of the value labels.
+The use of value labelling is very typical for a survey program, such as Eurobarometer, and a data archive, such as [GESIS](https://www.gesis.org/home), so we implement most of the value harmonization (and its helper functions) in the `eurobarometer` package.
+This implementation can be used as a template for other survey programs.
+
+### Documenting the process {#documentation}
+
+Our class definitions partly extended the S3 classes of haven for reproducibility and for automatic documentation.
+The haven package is designed to work with data from a single external resource --- it does not provide function to join data frames from different sources, and it does not record the call parameters of its functions.
+This is not a handicap when the researcher works with a single data file, but in our case, we aim to join data from potentially dozens of SPSS files.
+
+In the first step, we modified [read_spss()](http://retroharmonize.satellitereport.com/reference/read_spss.html) from haven to record some of the call parameters [*not yet all parameters are recorded, but the filename is*], most importantly the source file's name.
+Since we aim to work with `trust_in_eu_parlament` variables from different sources, for harmonization and validation we must be able to reference their source.
+We also create a simple id attribute, which may be the name of the survey, or the file name, a DOI, or any unique identifier of the original file.
+The result is a simple tibble (data.frame-like) object of class [survey *missing from documentation**], which has at least an id attribute.
+This attribute serves only identification and documentation purposes, therefore survey objects work exactly as tibbles in all situations.
+
+In the next step, [retroharmonize::harmonize_labels()](http://retroharmonize.satellitereport.com/reference/harmonize_values.html) changes the labels attribute and the base values of the [labelled_spss()](https://haven.tidyverse.org/reference/labelled_spss.html) vector, and at the same time we record the original coding and labelling of the vector.
+For example, if we created the id `ZA7976` for the respective Eurobarometer survey file, then the original labels will be stored in the attribute `ZA7976_labels`, and the original (character, integer or double) values in `ZA7076_values`.
+
+When this vector is concatenated with the same vector of observations from `ZA6968` survey, then apart from sharing the same labels and missing value ranges, the new vector will carry  `ZA7976_labels`, `ZA6968_labels`, `ZA7976_values` and `ZA6968_values`.
+The flexible metadata system of R allows us to extend this further when we are concatenating further and further observations from new files.
+
+Two simple helper functions support the user to retrieve this information easily --- in fact, to print an entire codebook of the harmonization process.
 
@@
 
-<this should be more data scientific >
+The advantage of defining a new S3 type with rather strict coercion rules is that after harmonization and validation, it allows for a safe and quick joining of the variables as vector with the `vctrs::vec_c()` or even the base R `c()` method, and a fast `rbind()` method.
 
-### Unique identification {#uniqid}
+In our experience, this is usually not possible with SPSS files imported by haven alone.
+Haven's two classes, labelled and [labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) cannot be concatenated, because it is not guaranteed that the missing values are similarly defined.
+If in a Eurobarometer wave the person creating the final SPSS version did not explicitly define missing values, but in a later wave it was defined, the data join will fail even if the integer coding and the labelling is perfectly matching.
+This is the reason of for creating our own coercion and recasting method, which is very strict, and defines a missing range in the labelled variables even if there are no missing values present.
 
-### Documenting the process {#documentation}
+However, in our experience, the lack of user-defined missing values does not mean that they should not be defined.
+There is no reason why certain Eurobarometer surveys indicate `do not know` answer options as a missing value category while others not.
+We implement a subjective solution for their harmonization in eurobarometer.
+Our implementation can be easily overridden.
 
-<probably this should be a heavier part>
+Joining entire survey objects, i.e.
+identified tibbles, would not be feasible, as they may have any kind of variables with unknown joining consequences.
+To avoid concatenating by vectors, i.e.
+simple questionnaire items, we created methods and functions to join question blocks - sufficiently similar parts of survey files.
+
+Naturally, surveys can be binded together if the same variables have the same names, and if they are labelled with categories or missing values, the labelling is consistent.
+This element of the workflow is rather idiosyncratic for the survey program in question.
+We devided the implementation to provide the joining, binding infrastructure in [retroharmonize](http://retroharmonize.satellitereport.com/), and to provide the variable name, label harmonization vocabulary and validation in the survey-specific [eurobarometer](http://eurobarometer.danielantal.eu/).
 
 ### Methods for exporting {#exportmethods}
+
+Our [labelled_spss_survey](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) is a bit extended [labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) class.
+It can be exported to SPSS or Stata with information loss.
+The lost information is the metadata of the harmonization history, which can be separately exported for documentation purposes.
+
 <just simple description and code links>
@@
 
-<just simple description and code links>
+As Joseph Larmarange well described in [Introduction to labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) (see: [cf.
+@R-labelled]), working with non-native microdata may have two logical workflows, out of which we chose workflow B.
+This separates the data processing work from analysis, and therefore allows to re-export the processed data (with some information loss) to SPSS or Stata.
+
+In the R language, almost all arithmetic and statistical functions work with two input types: numeric variables (integers or doubles) and factors (sometimes simplified to character).
+In order to use our harmonized data in R, we must provide methods to bring them to these types in a consistent way --- making sure that factor levels are harmonized and missing values are treated consistently.
+
+Our simple methods, [as_factor()](http://retroharmonize.satellitereport.com/reference/as_factor.html) (imported from haven) and [as_numeric()](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) precisely do this.
+While we would not advocate the use of character vectors for statistical analysis, for data visualization they may be more practical, and as_character() [*not het documented*] will display the value labels as a character vector.
+
+Our package implements simple arithmetic methods, such as median, quantile, mean, sum, weighted mean, for providing useful summary, printing and documentation methods, and for weighting without metadata loss.
+These methods are not exported, because we expect that the R user will analyze the data with their numeric or factor representations.
 

### 2020-07-10 data-raw/article.Rmd
update article

@@
 transformed and combined in a single data set in a way that matches the measurement of the same characteristics and standardizes the coding of responses.
-We introduce the surveyharmonization and eurobarometer packages that help the reproducible ex-post harmonization of survey data stored in various, individual files with rich metadata.
+We introduce the `retroharmonize` and `eurobarometer` packages that help the reproducible ex-post harmonization of survey data stored in various, individual files with rich metadata.
 We illustrate the proposed workflow using the case of the Eurobarometer surveys, which provide rich data on European publics in form of single-wave SPSS files from a period of over 25 years.
@@
 Thus, in this case the harmonization of the data across multiple Eurobarometer surveys boils down to data data transformation and documentation.
-For the purposes of presenting the process of harmonization with surveyharmonization and eurobarometer packages, we use the case of the Eurobarometer without the loss of generality of the described procedures, since it includes all data processing steps involved in harmonization.
+For the purposes of presenting the process of harmonization with `retroharmonize` and eurobarometer packages, we use the case of the Eurobarometer without the loss of generality of the described procedures, since it includes all data processing steps involved in harmonization.
 

### 2020-07-10 data-raw/article.Rmd
update article

@@
 author: "Daniel Antal, Marta Kolczynska"
-date: "7/6/2020"
+date: "7/10/2020"
 bibliography: [library.bib, packages.bib]

### 2020-07-13 data-raw/article.Rmd
update article (just links and a bit of code description, nothing fundamental.)

@@
 
-* Creating a full [documentation](#documentation) of conversions, recoding for reproducability.
+* Creating a full [documentation](#documentation) of conversions, recoding for reproducibility.
 
@@
 
-The consistent, harmonized treatment of the missing values allows us to easily implement the [as_numeric](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html), [as_character - *documentation missing*], and [as_factor](http://retroharmonize.satellitereport.com/reference/as_factor.html) methods, by replacing the user-defined missing values with type-specific NA values in the R language (NA_real_ or NA_character_).
+The consistent, harmonized treatment of the missing values allows us to easily implement the [as_numeric, as_factor and as_character](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) methods, by replacing the user-defined missing values with type-specific `NA` values in the R language (`NA_real_` or `NA_character_`).
 After this conversion, R artithmetic and statistical functions can be applied without compromising reproducability --- all metadata is present for a consistent export to SPSS or STATA, or to document, or even revert to an earlier labelling.
+
+SPSS and the [labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) class allows the parallel use of two attributes, `na_values` for enumeration of all values that should be treated as missing, or `na_range` for declaring a continuous range as missing.
+In surveying, the use of `na_values` is more practical, nevertheless, we must check the consistency of these attributes when we want to bind together several vectors.
+The [na_range_to_values](http://retroharmonize.satellitereport.com/reference/na_range_to_values.html) creates this consistency in case `na_range` is present in a vector.
+We did not implement the reverse `na_values` to `na_range` as it is unlikely to be used for survey harmonization.
+The function gives a warning in case of inconsistency --- in such a case we recommend a manual review of the valid and missing ranges of the vector, as this is more of a logical than a syntactical error in the survey file.
 
@@
 Working with [harmonize_values()](http://retroharmonize.satellitereport.com/reference/harmonize_values.html) is relatively easy if the researcher wants to join a few variables in a few survey files, but for large-scale, programmatic use it not really suitable.
+(See [vignette](http://retroharmonize.satellitereport.com/articles/harmonize_labels.html))
 
@@
 
-<just simple description and code links>
+>Unfortunately, the C++ code that writes from haven to sav (behind write_sav) is not implemented for the haven_labelled_spss class and does not record back user-defined NA values.
+I was not able to amend the code, but I wrote a detailed bug report.
 
-* exporting to SPSS
-* exporting to Stata
-* exporting to flat files
+>Currently, for testing and demonstration purposes, we must rely on the original GESIS files, or, if we want to show smaller or integrated examples, their rds representations.
+
+See [read_rds](http://retroharmonize.satellitereport.com/reference/read_rds.html) for the concept, the *write_rds* pair is not yet implemented, but easy to imagine.
+My [read_spss](http://retroharmonize.satellitereport.com/reference/read_spss.html) does what I imagined, but I cannot modify the haven::write_sav in C++ so until it is fixed, we cannot rely on that.
+
+> End of comments.
 
@@
 Our simple methods, [as_factor()](http://retroharmonize.satellitereport.com/reference/as_factor.html) (imported from haven) and [as_numeric()](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) precisely do this.
-While we would not advocate the use of character vectors for statistical analysis, for data visualization they may be more practical, and as_character() [*not het documented*] will display the value labels as a character vector.
+While we would not advocate the use of character vectors for statistical analysis, for data visualization they may be more practical, and [as_character()](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) will display the value labels as a character vector.
 

### 2020-07-16 data-raw/variable-keywords-table.Rmd
Update to variable names-keywords procedure (adds dictionary and survey metadata tables)

---
title: "Tagging variables with keywords"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vocabulary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
options(width = 160)
```

```{r setup}
library(eurobarometer)
library(tidyverse)
library(tibble)
library(knitr)
library(kableExtra)
library(tidytext)
library(janitor)
library(stringi)
```

An important aspect of survey data harmonization - in addition to standardizing response coding and value labels - is the identification of variables that measure the same thing, and standardizing the names of these variables. The process is faciliated by the fact that variables typically have labels that succintly describe the content of the variable, and sometimes also contain information about the coding of responses (e.g., length of scale), the unit (e.g., age in years or categories) and whether the variable contains "original" respondent responses or some transformation of responses (e.g. recoding or index).

Fortunately, in many multi-wave surveys, such as the Eurobarometer, there is a degree of consistency in how variables are labelled in different editions of the survey, and most survey projects try to follow some labelling conventions.

Unfortunately, consistency is limited and conventions are not followed strictly, which makes standardizing labels a semi-manual task. Some of these issues stem from the limits in the number of characters of variable labels in common statistical software packages (e.g., SPSS), which leads to non-standard abbreviations.

This vignette outlines a procedure for arriving at standardized variable names via tokenization of normalized variable labels (`var_label_norm`), which are created by [`label_normalize`](http://eurobarometer.danielantal.eu/reference/label_normalize.html) as part of [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html). The logic is similar to text analysis, in that a document-term matrix is created with each variable label treated as a (very short) document, and terms treated as keywords. variables that contain the same sets of keywords can be - with some margin of error - assumed to correspond to the same or equivalent survey questions.

## Procedure

0. Read survey and variable metadata, match variables to variable groups as defined by GESIS-ZACAT, filter out technical and protocol variables,

1. Standardize the spelling of certain often abbreviated words (e.g., 'gov', 'govnmt', and 'govmnt' for 'government'),

2. Tokenize,

3. Eliminate stop words (but be careful to not eliminate too many, e.g. 'right' or 'working'),

4. Create a document-term matrix (actually, variable label - keyword matrix),

5. Sort the columns according to the declining frequency of the keywords,

6. Create standardized variable labels by concatenating keywords.

### 0. Survey and variable metadata

#### Variable metadata

The table contains metadata from 97 Eurobarometer files (each file corresponding to one EB round). The metadata are extracted from SPSS files with [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html).

```{r}
metadata_database <- readRDS( file.path('../data-raw', 'eb_metadata_database_20200628.rds')) %>%
  mutate(archive_id = substr(filename, 1, 6))
```

#### Waves - Archive IDs

Mapping of GESIS Archive IDs (as used in file names) to wave numbers, from EB/44.2bis (ZA2828) to EB/92.1 (ZA7579) - 156 EB waves altogether.

```{r, warning = F}
eb_names <- rio::import("../data-raw/eb_waves_filenames.xlsx") %>%
  separate(x, c("archive_id", "x2", "title"), sep = ":") %>%
  separate(x2, c("wave", "dates"), sep = " \\(|\\)") %>%
  mutate(wave = gsub("Eurobarometer | ", "", wave))

eb_names %>%
  select(archive_id, wave, dates) %>%
  head() %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

#### Variable groups

Mapping of variable labels to variable groups, as defined in [GESIS ZACAT](https://zacat.gesis.org/webview/), for 51 waves from EB 79.1 (ZA5687) through EB 92.1 (ZA7579).

```{r}
varlabs <- rio::import("../data-raw/eb_variable_groups_zacat.xlsx", sheet = 4) %>%
  mutate(wave_desc = ifelse(substr(var_label_zacat, 1, 13) == "Eurobarometer", var_label_zacat, NA),
         var_label_zacat = ifelse(substr(var_label_zacat, 1, 13)=="Eurobarometer", NA, var_label_zacat),
         # group names are writte in camel case and != toupper(group names)
         group = ifelse(var_label_zacat != toupper(var_label_zacat), var_label_zacat, NA),
         # exceptions from the rule identifying group names
         group = ifelse(substr(group, 1, 9) == "OWNERSHIP", NA, group), 
         group = ifelse(substr(group, 1, 19) == "POSITIVE EU RESULTS", NA, group),
         group = ifelse(substr(group, 1, 24) == "OPINION LEADERSHIP INDEX", NA, group),
         var_label_zacat = ifelse(!is.na(group), NA, var_label_zacat)) %>%
  fill(wave_desc, group) %>%
  drop_na(group, var_label_zacat) %>%
  mutate(wave = substr(wave_desc, 1, 18),
         wave = gsub("^Eurobarometer ", "", wave),
         var_label_zacat = gsub("^W[0-9]{1,2}[A-Z] |^W3A/W4A ", "", var_label_zacat)) %>%
  left_join(eb_names) %>%
  select(wave, archive_id, group, var_label_zacat, everything())

varlabs %>%
  select(archive_id, wave, group, var_label_zacat) %>%
  head() %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

```{r}
# vector of archive IDs
archive_ids <- varlabs %>% distinct(archive_id) %>% pull(archive_id)

length(archive_ids)
```

### 1. Standardize the spelling

All procedures below applied to 51 waves, for which all metadata are available.

#### 1.1. Translation table ...

```{r}
dict <- rio::import("../data-raw/var_lab_dict.xlsx", sheet = "dict") %>%
  gather(id, from, 2:6) %>%
  select(from, to) %>%
  drop_na(from)

head(dict) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

#### 1.2. ... applied to metadata

```{r}
# this part now became very fast with stri_replace_all_regex

dat2 <- metadata_database %>%
  # filter only waves for which variable groups are available
  filter(archive_id %in% archive_ids,
         # exclude this strange variable
         var_label_orig != "filename") %>%
  # keep one row per variable
  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm, archive_id) %>%
  # original variable labels in the data and in GESIS ZACAT are not exactly the same 
  # --> need some cleaning
  mutate(var_label_orig2 = gsub("^W[0-9]{1,2}", "", var_label_orig),
         var_label_orig2 = gsub("^W[0-9]{1,2}[A-Z]", "", var_label_orig),
         var_label_orig2 = gsub("^[0-9]{1-3} ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z][0-9]{1,2}[A-Z] ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2} ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2}[A-Z] ", "", var_label_orig2),
         var_label_orig2 = gsub("^[0-9]{1,2} ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z][0-9] [0-9]{1,2} ", "", var_label_orig2),
         var_label_orig2 = gsub("^W3A/W4A ", "", var_label_orig2)) %>%
  # join with table with variable groups
  left_join(varlabs, by = c("var_label_orig2" = "var_label_zacat", "archive_id")) %>%
  # remove variables in the following groups
  filter(!group %in% c("Archive and survey ID variables",
                       "Country group variables",
                       "Nationality (Q1)",
                       "Protocol variables (P)",
                       "Standard nation ID variables",
                       "Weight variables (W)",
                       "Nation ID variables",
                       "Protocol variables",
                       "Weight variables",
                       "Archive and ID variables",
                       "Household Weights"),
         # remove recoded variables
         !grepl("recoded|(rec)", var_label_norm)) %>%
  mutate(var_id = paste0(filename, " ", var_name_orig),
         # leftover question numbers at the beginning of value labels
         var_label_norm1 = gsub("^w[0-9]{1,2}", "", var_label_norm),
         var_label_norm1 = gsub("^[0-9]{1-3}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[0-9]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", var_label_norm1)) %>%
  # apply the translation table
  mutate(var_label_norm1 = stri_replace_all_regex(var_label_norm1,
                                                  pattern = dict$from,
                                                  replacement = dict$to,
                                                  vectorize_all = FALSE)) %>%
  # replace underscore with space
  mutate(text = gsub("_", " ", var_label_norm1)) %>%
  ungroup() %>%
  # select the needed variables
  select(filename, qb, var_label_orig, var_label_norm, var_label_norm1, text, var_id, group)

```

### 2 & 3. Tokenize & Eliminate stopwords

```{r}
# list of country codes to be converted to country names
eu_countries <- c(
  "at", "be", "bg", "cy", "cz", "dk", "de", "ee", "es", "fr",
  "fi", "gb", "gr", "hr", "hu", "ie", "it", "lt", "lv", "lu",
  "mt", "nl", "pl", "pt", "ro", "se", "si", "sk", "tr")

# !!! the problem here is that dk sometimes means "don't know" and sometimes Demnark !!!
# this needs to be treated in gesis_metadata_create, where the
# user defined missings are separately analyzed, they dk should 
# be converted there to do_not_know.

# it's not clear hot to do this, because both types of DK appear at the end of labels e.g. in
# mentioned/not mentioned questions, and they look very much the same

tidy <- dat2 %>%
  # tokenize
  unnest_tokens(word, text, 
                token = "regex", pattern = " |  ") %>%
  # eliminate stop words, with some exceptions
  anti_join(stop_words %>%
              filter(!word %in% c("working", "right", "other")
                     )) %>%
  # eliminate numbers
  filter(!grepl("\\b\\d+\\b", word)) %>%
  # convert country codes to country names
  mutate(
    country = ifelse(word %in% eu_countries, word, NA),
         country = countrycode::countrycode(
           toupper(country), "iso2c", "country.name"),
    country = tolower(country),
    word = ifelse(!is.na(country), country, word),
    word = ifelse(word == "united kingdom", "united-kingdom", word),
    word = ifelse(word == "tcc", "turkish-cypriot-community", word)
    ) %>%
  select(-country)
```

```{r}
tidy %>%
  select(var_id, var_label_norm1, word) %>%
  head() %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

```{r}
tidy %>%
  count(word) %>%
  arrange(desc(n)) %>% 
  head(., 10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Note: this Denmark is probably a "don't know"

### 4. Create a document-term matrix

Filter out labels that have words that appear in < 5 different waves. If we want to create a trend file, we're only interested in variables that are repeated often.

```{r}
tidy_small_2 <- tidy %>%
  group_by(word) %>%
  # number of distinct file names (waves) per word
  mutate(n_wave = n_distinct(filename)) %>%
  # flag words that appear in < 5 different files
  mutate(to_drop = as.numeric(n_wave < 10)) %>%
  # group by variable ID (file x variable name)
  group_by(var_id) %>%
  # copy to_drop across words in the same variable label
  mutate(to_drop = max(to_drop, na.rm = TRUE)) %>%
  # drop variable labels with the to_drop identifier
  filter(to_drop != 1) %>%
  # remove auxiliary variables
  select(-to_drop, -n_wave) %>%
  ungroup()

dtm <- tidy_small_2 %>%
  distinct(var_id, var_label_norm1, word) %>%
  mutate(n = word) %>%
  spread(word, n)

dim(dtm)
```

### 5. Sort columns according to the declining frequency of the keywords

```{r}
# get the order of variables according to declining proportion of missing value
# (declining popularity of keywords)
order <- dtm %>%
  summarise_all(funs(sum(is.na(.))/length(.))) %>%
  t() %>% as.vector()

names(order) <- names(dtm)
```

### 6. Create standardized variable labels by concatenating keywords

Standardized variable labels `var_label_std` are concatenated keywords, where the keywords occur in the order of their frequency in the entire corpus of variable labels. As a consequence, more common - likely broader - terms will come first, and less common - more specific - terms will come later.

```{r}
dtm2 <- dtm %>%
  select( names(sort(order)) ) %>%
  mutate_at(vars(3:ncol(dtm)), 
            funs(ifelse(is.na(.), "", .))
            ) %>%
  unite("var_label_std", 3:ncol(dtm)) %>%
  mutate(var_label_std = gsub(
    "(\\_)\\1+", "\\1", var_label_std) ) %>%
  mutate(var_label_std = gsub(
    "^_|_$", "", var_label_std) )

dim(dtm2)
```

These standardized variable names still have problems that need to be fixed.

```{r}
dtm2 %>% 
  head(., 25) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Print the most common standardized variable labels.

```{r}
dtm2 %>%
  count(var_label_std) %>%
  arrange(desc(n)) %>% 
  head(., 25) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

## Application: political attitudes

Filter on substrings of standardized variable labels `var_label_std`, and count the repetitions across datasets.

```{r}
# filter on var_label_std
dtm2 %>%
  filter(grepl("trust|satisf|left-right|interest|class", var_label_std)) %>% 
  count(var_label_std) %>%
  arrange(desc(n)) %>%
  head(., 25) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

### 2020-07-16 data-raw/article.Rmd
Documenting the recent developments

@@
 The haven package has already adopted the importing of user-defined missing values, or the user defined missing value range with the creation of the class [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html).
-While this class generally works well with the older [haven_labelled](https://haven.tidyverse.org/reference/labelled.html) methods, it does not have many implemented methods in haven.
+While this class generally works well with the simpler [haven_labelled](https://haven.tidyverse.org/reference/labelled.html) methods, it does not have many implemented methods in haven.
 In our case, we had to make sure that when we concatenate two vectors with user-defined missing values or ranges, they will remain consistent after the join.
-Our class relies on the [vctrs](https://vctrs.r-lib.org/) package, which contains helper methods for creating new vector classes [cf.
+Our [retroharmonize_labelled_spss_survey](http://retroharmonize.satellitereport.com/articles/labelled_spss_survey.html) class is an extension of the [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html).
+It relies on the [vctrs](https://vctrs.r-lib.org/) package, which contains helper methods for creating new vector classes [cf.
 @R-vctrs], and which serves as the foundation of haven's own class system, too.
 
-Our choice in this case is the that we apply integer values, long out of the range of categorical integer values in the range `99900:99999` for a consistent numeric coding of the values to be treated as missing, and we label them consistently.
+Our choice in this case is the that we apply integer values, long out of the range of categorical integer values in the range `99900:99999` for a consistent numeric coding of the values to be treated as missing, and then we label them consistently.
 The original missing labels and missing values as preserved as an attribute in our [labelled_spss_survey](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) class.
+We add missing labels to data that does not even have missing values for consistency.
 
-The consistent, harmonized treatment of the missing values allows us to easily implement the [as_numeric, as_factor and as_character](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) methods, by replacing the user-defined missing values with type-specific `NA` values in the R language (`NA_real_` or `NA_character_`).
+The [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) does not have many implemented coercion and casting methods, because there is no objective way of handling the different missing values.
+We had to make some subjective and consistent choices to allow a consistent re-casting with [as_numeric, as_factor and as_character](http://retroharmonize.satellitereport.com/reference/labelled_spss_survey.html) methods.
+For full reproducability, before replacing the user-defined missing values with the type-specific `NA` values in the R language (`NA_real_` or `NA_character_`), we carefully record the labels of these conversions, and even their pre-harmonization, historical labelling, too.
 After this conversion, R artithmetic and statistical functions can be applied without compromising reproducability --- all metadata is present for a consistent export to SPSS or STATA, or to document, or even revert to an earlier labelling.
+
+The lack of clear coercion and casting disallows the concatenation or binding of [haven_labelled_spss](https://haven.tidyverse.org/reference/labelled_spss.html) vectors in a data-frame.
+Our subjective choice of pushing all user-defined missing values to the numeric range of range `99900:99999` with the same labelling, and applying exactly the same coding and labelling of the valid range allows us to use [vctrs::vec_c](https://vctrs.r-lib.org/) and [vctrs::vec_rbind](https://vctrs.r-lib.org/) for binding the variables together.
 
@@
 Two simple helper functions support the user to retrieve this information easily --- in fact, to print an entire codebook of the harmonization process.
+
+The [document_survey_item](http://retroharmonize.satellitereport.com/reference/document_survey_item.html) function exports the entire history of the harmonization process along a codebook for the variable.
+...
+the looping version is not yet desinged....
 

### 2020-07-17 data-raw/variable-keywords.Rmd
var_label_normalize

@@
 ```{r}
-metadata_database <- readRDS( file.path('../data-raw', 'eb_metadata_database_20200628.rds'))
+metadata_database <- readRDS( file.path('..', 'data-raw', 'eb_metadata_database_20200628.rds'))
 ```
@@
   # clean up variable labels to unify the spelling of certain common words and abbreviation
+  #
+  #  FROM HERE on this is var_label_normalize()
+  mutate(
+    var_id = paste0(filename, " ", var_name_orig),
+    var_label_norm1 = var_label_normalize(var_label_norm)
+    ) %>%
+  ungroup() %>%
+  # select the needed variables
+  select(filename, qb, var_label_norm1, var_label_orig, var_id, text)
+```
+
+```{r, eval = FALSE}
+## went to var_label_normalize()
   mutate(
@@
     # replace underscore with space
-    text = gsub("_", " ", var_label_norm1)) %>%
-  ungroup() %>%
-  # select the needed variables
-  select(filename, qb, var_label_norm1, var_label_orig, var_id, text)
+    text = gsub("_", " ", var_label_norm1)) 
 ```

### 2020-07-17 R/var_label_normalize.R

@@ -1,0 +1,375 @@
+#' Create Normalized Variable Labels
+#'
+#' Remove relative reference to question blocks, special characters
+#' used in regular expressions, whitespace.
+#'
+#' @param x A vector of the (GESIS) variable or value labels
+#' @importFrom retroharmonize var_label_normalize
+#' @family naming functions
+#' @return A character vector with the normalized labels.
+#' @examples
+#' label_normalize (
+#'     c(QF4C INTERESTS INFO SOURCES: INTERNET',
+#'     'QB15A CHILDREN NEEDS: SOME NEW CLOTHES',
+#'     'QB9B RESEARCH INFO SOURCE - PREFERENCE 2ND',
+#'     'EU MEANING: CULTURAL DIVERSITY',
+#'     'P6 SIZE OF COMMUNITY - AUSTRIA',
+#'     'D49D_CY WEBSITES: OTHER',
+#'     'LOVE RELATIONSHIP OF CHILD - CHRISTIAN PERSON' )
+#' )
+#' @export
+
+var_label_normalize <- function(x) {
+
+  y <- retroharmonize::var_label_normalize(x)
+
+  ##do the abbreviations first that may have a . sign
+  y <- gsub( '\\ss\\.a\\.', '_sa', y)
+
+  ## remove prefix of question block numbers
+  y <- gsub( '^q[abcdef]\\d{1,2}[abcdef]', '', y )  # remove QB9B, QB25 etc
+  y <- gsub( '^q[abcdef]\\d{1,2}_\\d{1,2}', '', y )  # remove QA117_1
+  y <- gsub( '^q[abcdef]\\d{1,2}', '', y )  # remove QA1, QB25 etc
+  y <- gsub( '^q\\d{1,2}[abc]', '', y )  # remove QA1, QB25 etc
+  y <- gsub( '^d\\d{1,2}[abcdef]', '', y )       # removed d26a_ etc
+  y <- gsub( '^d\\d{1,2}', '', y )       # removed d26_ etc
+  y <- gsub( '^c\\d{1,2}', '', y )       # removed c26_ etc
+  y <- gsub ( "^p\\d+{1,}_", "", y)  #remove p6_ like starts
+  y <- gsub ( "^p\\d+{1,}\\s", "", y)  #remove p6  like starts
+  y <- gsub ( "^q\\d+{1,}_", "", y)  #remove q1_ like starts
[…]

### 2020-07-17 R/gesis_metadata_create.R
add new vignette starts example data

@@ -55,11 +55,13 @@ gesis_metadata_create <- function ( survey_list ) {
     mtd <- retroharmonize::metadata_create(dat) %>%
       mutate (
         filename = dat$filename[1],
-        var_label_norm = eurobarometer::var_label_normalize(label_orig),
+        var_label_norm = var_label_normalize(label_orig)
+        ) %>%
+      mutate (
         # do we need this?
         var_name_suggested = label_suggest( var_label_norm,
-                                            names(dat))
-               )
+                                            var_name_orig )
+      )
 
     ## Creating the basic metadata ----
     metadata  <- mtd %>%
@@ -82,9 +84,7 @@ gesis_metadata_create <- function ( survey_list ) {
 
     metadata <- question_block_identify(metadata)
 
-    names (metadata)
-
-   metadata %>%
+    metadata %>%
       select ( all_of(c("filename", "qb", "var_name_orig",
                         "var_label_orig",
                         "var_label_norm", "var_name_suggested",
@@ -93,14 +93,11 @@ gesis_metadata_create <- function ( survey_list ) {
                         "n_categories",
                         "labels","na_levels", "na_range",
                         "class_orig")
-                      )
-               )
+              )
+      )
 
-    ## class_suggest is not neeeded any more  in utils.R
-
+    ## class_suggest is not needed any more  in utils.R
   }
-
-  tmp <- metadata_create( dat = survey_list[[1]])
 
   metadata_list <- lapply ( survey_list, metadata_create )
   do.call(rbind,metadata_list)

### 2020-07-17 old_code/old_vignettes/create_tables.Rmd
new vignettes, old functions in not built directory

---
title: "Creating Variable and Values Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Variable and Values Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warnings=FALSE}
library(eurobarometer)
library(dplyr)
library(tibble)
library(knitr)
library(kableExtra)
library(rprojroot)
# The examples of this vignette can be found run with  
# source(
#  file.path("not_included", "vignette_vocabulary_examples.R")
#  )
```

## Procedure {#procedure}

The procedure involves four steps:

1. `Concept definition`: Trust in institutions: general trust without specifying the concrete domains of activity where the trust applies (these specific questions would require special treatment).

2. `Selection of survey questions` that correspond to the concept by filtering on variable and value labels.

3. `Standardizing value labels` across questions.

4. `Standardizing variable name`s to have the same name for all variables that correspond to the same question.

## Metadata

`metadata_database` is a table with variable metadata created from Eurobarometer SPSS files. It has the following columns:

`var_name_orig`: variable name in the original dataset  
`class_orig`: column class in the original dataset   
`var_label_orig`: variable label in the original dataset  
`var_label_norm`: normalized variable label  
`var_name_suggested`: suggested variable label  
`factor_levels`: vector of value labels as a list (they have no fixed length)  
`n_categories`: total number of response options `= length(factor_levels)`   
`class_suggested`: suggested conversion to an R class, it should correspond later to a conversion function, so that the researcher can just simply approve and get a correct R representation.  
`filename`: original name of file as obtained from GESIS; it contains a disambigous version information, too.
`val_numeric_orig`: numeric value code (if available, empty otherwise)   
`val_label_orig`: value label in the original dataset corresponding to a response option
`val_order_alpha`: alphabetical number (position) of response option in the set of value labels, after sorting with `sort()` (we use alphabetical order instead of `levels()` because levels may not be the same in different survey files, and we revert to a basic but disambigous sorting)   
`val_order_length`:  
`val_label_norm`: value label normalized with `label_normalize()`

```{r, eval=FALSE}
# Avoid changes in the working directory when building vignettes:
metadata_rel_path_from_root <- find_root_file(
  "data-raw", "eb_metadata_database_large.rds", 
  criterion = has_file("DESCRIPTION"))

metadata_database <- readRDS( metadata_rel_path_from_root )
```

```{r}
data("metadata_filter_example")
metadata_database <- metadata_filter_example
rm(metadata_filter_example)
```
## Selection of variables

This is an iterative process.

### Iteration 1

Let's filter out questions with `trust` in their variable labels.

```{r}
select_metadata_vars <- c(
  "filename", "var_name_orig", "var_label_orig",
  "var_label_norm", "val_label_orig", "val_label_norm", 
  "val_numeric_orig", "val_order_alpha", "n_categories")

trust_metadata <- metadata_database  %>%
  filter (
    grepl( "trust", var_label_norm )
  )  %>%
  select ( all_of(select_metadata_vars) ) %>%
  arrange ( var_label_norm, val_label_norm, filename )

trust_metadata %>%
  head(10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Scrolling through the normalized variable names (see `?label_normalize`) created from the GESIS SPSS variable labels, it looks like binary value trust in institutions questions have in their label `trust in institutions` or `trust political parties` or `- trust` at the end. Let's take a closer look at those questions and see what response options they have.

Several variables from `Eurobarometer 69.2` (ZA4744) ask more detailed questions about specific reasons for trusting or not trusting selected institutions. These are different types of questions with more categories than the more general trend questions with a binary value `tend to trust` - `tend not to trust`.

### Iteration 2

```{r}
trust_metadata <- trust_metadata %>% 
  filter (
    grepl( 
      "trust_in_institutions|trust_political_parties|_trust$",
      var_label_norm )
  ) %>%
  filter( ! (filename == "ZA4744_v5-0-0.sav" &
               var_name_orig %in% c("v336", "v347","v292", "v303",
                                    "v314", "v325", "v270", "v281")
             )
          )

trust_metadata %>%
  arrange(filename, var_name_orig, val_order_alpha) %>%
  head(10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Check how many unique response values do the selected variables have?

```{r}
trust_metadata %>%
  arrange(val_label_norm) %>%
  select(val_label_norm, everything()) %>% 
  count(n_categories) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

These variables have 2, 3 or 4, or as many as 13 unique response options. Let's take a look at the one with 13 unique responses:

```{r}
trust_metadata %>%
  filter(n_categories == 13) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

The question asks:
*Who do you trust the most to fight effectively in [country] against the European Union and its budget being defrauded?*
(see documentation in the [GESIS documentation](https://search.gesis.org/research_data/ZA3938)).

This refers to trust with regard to a specific activity, so does not fit into our concept definition (see [procedure](#procedure).) Let's see what we are left with after excluding it.

### Iteration 3

```{r}
trust_metadata <- trust_metadata %>%
  filter( ! ( filename == "ZA3938_v1-0-1.sav" &
                var_name_orig == "v511")
          )

```

```{r, results='asis'}
trust_metadata %>%
  count(val_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

* `dk` stands for "don't know",
* `inap*` stands for "inapplicable" for various reasons, which are maybe worth exploring,
* `mentioned` and `not mentioned` suggest questions where the respondent is asked to list all objects to which some condition applies, e.g. "here is a list of institutions. which of those would you say you trust?". These questions require special attention, because "no" is blended with "no answer / don't know". Plus, it's a very different type of question for the respondent,
* `na` and `lt_na_gt` are missing values,  
* `tend not to trust` is a negative answer to the trust question,
* `tend to trust` is a positive answer to the trust question,
* `you generally do not trust stories published on online social networks` is a mystery.

Let's start with the mystery.

```{r}
trust_metadata %>% 
  filter( 
    grepl("you_generally_do_not_trust", val_label_norm)
    ) %>%
  select(-var_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

That's likely not what we are interested in, but still worth checking with the [GESIS documentation](https://search.gesis.org/research_data/ZA6861).

The question reads `QD5 When you see or read a story published on online social networks, what makes you consider the story trustworthy?` and has several "mentioned/not mentioned" items. This follows the usual multiple choice question structure of Eurobarometer, where all possible choice options are coded as  separate variables, and one of them ended with "trust", which is why we filtered it out. We don't want it and will remove it. _This should generally be fixed with filtering out multiple choice questions because they follow a particular coding, see[Prefix conventions](http://eurobarometer.danielantal.eu/articles/variable_names.html)_

We refine the variable filter.

### Iteration 4

```{r}
trust_metadata <- trust_metadata %>% 
  filter( ! ( filename == "ZA6861_v1-2-0.sav" &
                var_name_orig == "qd5.6"
              )
          )
```

```{r, results='asis'}
trust_metadata %>%
  count(val_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Now let's at the questions that have the `mentioned` - `not mentioned` response options. It's not what we want.

```{r}
exclusions <- trust_metadata %>%
  filter(val_label_norm %in% c("mentioned", "not_mentioned")) %>%
  print() %>%
  count(filename, var_name_orig)
```

### Iteration 5

```{r}
trust_metadata <- trust_metadata %>%
  anti_join(
    # with a filtering join exclude all from iteration 4
    exclusions, 
    by = c("filename", "var_name_orig")
    )
```

```{r, results='asis'}
trust_metadata %>%
  count(val_label_norm) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

## Standardizing value labels across questions

Now we need to create a table with all value labels, as well as topics / variable types.

```{r}
val_labels_trust <- trust_metadata %>%
  count(val_label_norm)

trust_vocabulary <- tibble::tibble (
  # maybe we can use a generic controlled vocabulary, for example
  # Library of Congress
  topic_1 = 'trust institutions',
  # And if we find them, we can add GESIS or TNS/Kantar keywords here
  topic_2 = 'trust, binary',
  val_label_norm = val_labels_trust %>% pull(val_label_norm), 
  level = 3 # missingness should be harmonized in character form 
)
```

```{r, results='asis'}
trust_vocabulary %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

We add the harmonized values and labels, as well as a flag indicating whether the value is substantive of one of missing value codes, to create the `trust_values_table`.

```{r}
trust_values_table <- trust_vocabulary %>%
  mutate(character_value = case_when(
      # and create a surely harmonized character representation 
      grepl("dk|inap|na", val_label_norm) ~ NA_character_, 
      substr(val_label_norm, 1,7) == "tend_to" ~ "tend_to_trust",
      substr(val_label_norm, 1,11) ==
        "tend_not_to" ~ "tend_not_to_trust",
    TRUE ~ "ERROR"), 
    numeric_value = case_when ( 
      character_value == "tend_to_trust"     ~ 1,
      character_value == "tend_not_to_trust" ~ 0,
      TRUE ~ NA_real_),
    missing = case_when (
      # it is useful for faster filtering of missingness
      # and true value labels  
      is.na(numeric_value) ~ TRUE, 
      TRUE ~ FALSE)
  ) %>%
  arrange(numeric_value)
```

```{r, results='asis'}

trust_values_table %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

## Standardizing variable names

The `trust_variable_table` will include original and harmonized variable names, as well as keywords and topics to match with the `trust_values_table`.

```{r}
trust_variable_table <- trust_metadata %>%
  filter ( 
    val_label_norm %in% trust_values_table$val_label_norm 
    ) %>%
  filter ( ! grepl("_recoded", var_label_norm ) ) %>%
  mutate ( 
    # normalize the names of institutions
    institution = gsub(
      "trust_in_institutions_|in_|_trust|trust_|[0-9]_|^a_|^b_|^q16c_|the_", 
      "", var_label_norm ),
    institution = gsub(
      "charitable_org", "charities", institution ),
    institution = gsub(
      "europ_court_of_auditors|eur_court_of_auditors",
      "eu_court_of_auditors", institution ),
    institution = gsub(
      "europ_court_of_justice|european_court_of_justice",
      "eu_court_of_justice", institution ),
    institution = gsub(
      "justice_legal_system|justice_nat_legal_system",
      "justice", institution ),
    institution = gsub(
      "nat_parliament", "national_parliament", institution ),
    institution = gsub(
      "nat_government", "national_government", institution ),
    institution = gsub(
      "non_govmnt_org|non_govnmt_org", "ngo", institution ),
    institution = gsub(
      "polit_parties", "political_parties", institution ),
    institution = gsub(
      "reg_loc_public_authorities|reg_local_authorities|reg_local_public_authorities|rg_lc_authorities", 
      "reg_loc_authorities", institution ),
    institution = gsub(
      "written_press", "press", institution ),
    institution = gsub(
      "econ_and_social_committee|economic_and_soc_committee", 
      "econ_and_soc_committee", institution ),
  ) %>%
  mutate ( 
    ## handle exceptions from original questionnaire variants
    geo_qualifier = case_when(
      grepl("_tcc", institution) ~ "tcc", #Turkish Cypriot Community
      TRUE ~ NA_character_),
    institution = gsub("_tcc", "", institution)
  ) %>%
  mutate ( 
    ## follow exception handling for special questionnaires
    var_name_suggested = paste0(
      "trust_", institution, "_", geo_qualifier)
    ) %>%
  mutate (
    # remove _na where the no geo_qualifier is present as base case
    var_name_suggested = gsub("_NA", "", var_name_suggested),
  ) %>%
  mutate ( 
    ## add some topical keywords to our table by taste
    topic_1 = "trust institutions",
    topic_2 = "trust, binary"
    ) %>%
  select ( 
    ## remove aux variables 
    -contains("label"),
    -all_of(c("val_numeric_orig", 
              "val_order_alpha", 
              "n_categories"))
  ) %>%
  rename ( keyword_1 = institution ) %>%
  distinct_all() %>%
  mutate(var_name_suggested = ifelse(
    #one last exception to handle
    test = var_name_suggested == "trust_political_partiess_tcc",
    yes = "trust_political_parties_tcc", 
    no = var_name_suggested )
    ) %>%
  arrange(var_name_suggested)
```

Let's print the first 10 rows of the resulting table

```{r, results='asis'}
trust_variable_table %>% 
  head(10) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

How many different harmonized variables do we get? 54, each in between 1 and 42 Eurobarometer editions.

```{r}
trust_variable_table %>%
  count(var_name_suggested) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

### 2020-07-17 old_code/old_vignettes/variable_names.Rmd
new vignettes, old functions in not built directory

---
title: "Harmonizing Variable Names"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Harmonizing Variable Names}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
require(kableExtra)
```

```{r setup, message=FALSE, warnings=FALSE}
library(eurobarometer)
library(knitr)
library(kableExtra)
library(dplyr)
```

# Harmonizing Variable Names

## Normalization

We are using the SPSS `.sav` files as a data resource from GESIS.  The SPSS files are combining data with metadata labels.  Furthermore, GESIS often includes metadata, including constants, into the SPSS file. The [label_normalize()](http://eurobarometer.danielantal.eu/reference/label_normalize.html) function is used to create a first suggestion for a variable name using the GESIS SPSS file's variable labels.

The normalization rules can be reviewed in the source code of the `normalize_names()` function.

The variable names are following the `underscore_separated`naming convention which is often called the `snake_case_convention`.  We are using prefix and suffix conventions for easy separation and detachments of certain special variables and types.

* we are removing all special characters that may be understood as regular expressions, such as `%`, `]`, and all whitespace.

* We use Eurostat abbreviations similar to some Eurostat metadata vocabularies for `greater than`, `lesser than` and similar quasi-arithmetic expressions, and `pct` for percent, percentage or the `%` sign.

### Prefix conventions
* `id_`: easy selection and detachments of constant metadata, such as the document object identifier or the version number of the data file; except for the `uniqid` variable, which is an observation / row identifier in almost all files.

* `mc_` : multiple choice question items.  These are logically related variables that are given in the questionnaire in tabular format, but they are coded both in SPSS and R item by item as separate variables (columns).  For the same (structured) question, we always use the `mc_this_question_item_1`, `mc_this_question_item_2`, `mc_this_question_item_dk`  naming convention, where the `_dk` suffix contains the indication that the respondent rejected the entire complex question.

* `is_` : indicator questions, which may created from binary answer options, or from multiple-choice like questions without clear structure.

These questions can easily be selected for example with the `tidyselect::starts_with(“mc_”)` selector which is widely used in the `tidyverse`.

### Suffix conventions

Currently we are only using suffixes for questionnaire variations.  The most likely variation is related to the fact that some fieldwork areas are not part of the core Eurobarometer territory, and for some reason they separated in the original data file, because the respondent received a shorter, or slightly modified questionnaire. In these cases, some of the questions are the same as in the core questionnaire, and can be imputed to the core question.

* `_tcc`: Question variation for members of the Turkish Cypriot community

* `_mk`: Question variation used in North Macedonia.

## Preferred expressions
With often used variables, we try to use the nearest lower_snake_case version of the regularly used questionnaire name without alphanumeric IDs, because we believe that the researchers of Eurobarometer are familiar with these names.  We use this for example with `age_exact`, `type_of_community` and some other often used variables.

Often, repeating (trend) questions can be found under several, often dissimilar labels.  In these cases, we are choosing a preferred version of the variable name, and we are creating metadata tables to bring other variations to this format.

## Less frequent questions and variables
In less frequently used questions, questionnaire items we are creating programmatic variable names using roughly the convention of rOpenSci.  In these cases the creation of time series or data panels is less likely, or requires more work, because the questions or the questionnaire items are not the same. Our approach in this case is to create metadata tables for naming that are following topics.  We hope to find contributors who are familiar with certain topics, and can organize the similar but not exactly the same variables into topical groups with similar variable names, such as variables related to trust in institutions or climate change. 
In these cases we follow general rules:

* The variable name starts with a more general concept and goes towards a specificity

* When applicable, we prefer the rOpenSci `object_verb()` organization (see: [Function and argument naming](https://devguide.ropensci.org/building.html#package-api)), but overrule it for convenience or avoiding awkward variable names.

* Only fully identical questionnaire items have identical names

* Similar questionnaire items have similarly structured names for easier selection and comparison

* We try to avoid extremely long variable names

The preferred variable names are stored in a topical metadata table that contains standardized topical keywords, and serves as a data and questionnaire map to the researcher.

## Example Metadata Table

This metadata table is created by `not_included/vignette_vocabulary_examples.R`.  It is a good practice to create the metadata tables programmatically for reproducibility.

After the modification of the source code of this R file, `data-raw/create_categorical_var.R` can update the data file.  The documentation of the metadata table should be updated in `R/data-category_label_2.R`

Similar metadata tables should help the naming for similar variables (for example, in this case, often used two-level categories) or for topics (such as trust in institutions for not two-level variables, or air pollution and climate change related questions.)

```{r}
data("categorical_variables_2", package = "eurobarometer")

categorical_variables_2 %>%
  kable() %>%
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed"), 
                  fixed_thead = T, 
                  font_size = 10 ) %>%
    add_header_above(c("Filtering" = 3, 
                       "Preferred Term" = 1, 
                       "Keywords" = 2)
                     )
```

We must decide early on the column names of the metadata table itself, i.e. `filename`, `r_name`, `normalized_names`, `canonical_names`, `keyword_1`, `geo_qualifier`.

### 2020-07-17 R/retroharmonize_imports.R

@@ -1,10 +1,16 @@
-
+#' @inheritParams retroharmonize::read_rds
 #' @importFrom retroharmonize read_rds
 #' @export
 read_rds <- retroharmonize::read_rds
 
+#' @inheritParams retroharmonize::as_factor
+#' @export
+as_factor <- retroharmonize::as_factor
 
-library(usethis)
-use_build_ignore("inst/")
-path <-
-read_rds(path)
+#' @inheritParams retroharmonize::as_numeric
+#' @export
+as_numeric <- retroharmonize::as_numeric
+
+#' @inheritParams retroharmonize::as_character
+#' @export
+as_character <- retroharmonize::as_character

### 2020-07-17 R/utils.R

@@ -1,15 +1,3 @@
-#' @inheritParams retroharmonize::as_factor
-#' @export
-as_factor <- retroharmonize::as_factor
-
-#' @inheritParams retroharmonize::as_numeric
-#' @export
-as_numeric <- retroharmonize::as_numeric
-
-#' @inheritParams retroharmonize::as_character
-#' @export
-as_character <- retroharmonize::as_character
-
 #' Suggest Conversion
 #'
 #' @param metadata A metadata file created by metadata_create within

### 2020-07-17 old_code/old_functions/concatenate.R

@@ -1,0 +1,95 @@
+#' @title Concatenate haven_labelled_spss vectors
+#'
+#' @param x A haven_labelled_spss vector.
+#' @param y A haven_labelled_spss vector.
+#' @return A concatenated haven_labelled_spss vector. Returns an error
+#' if the attributes do not match. Gives a warning when only the variable
+#' label do not match.
+#' @examples
+#' s1 <-labelled_spss(
+#'   x = unclass(v1),         # remove labels from earlier defined
+#'   labels = val_labels(v1), # use the labels from earlier defined
+#'   na_values = NULL,
+#'   na_range = 8:9,
+#'   label = "Variable Example"
+#' )
+#'
+#' s2 <-labelled_spss(
+#'   x = unclass(v2),         # remove labels from earlier defined
+#'   labels = val_labels(v2), # use the labels from earlier defined
+#'   na_values = NULL,
+#'   na_range = 8:9,
+#'   label = "Variable Example"
+#' )
+#' concatenate (x,y)
+#' @importFrom haven is.labelled_spss labelled_spss
+#' @family joining functions
+#' @export
+
+concatenate <- function(x, y) {
+  if ( ! all(c(is.labelled_spss(x), is.labelled_spss(y))) ) {
+    stop ("Both arguments must be haven_labelled_spss")
+  }
+
+  if ( is.null(attr(x, "na_range")) ) {
+    if ( ! is.null(attr(y, "na_range")) ) {
+      stop("The first (x) na_range is NULL, the second (y) is not.")
+    }
+  } else if ( is.null(attr(y, "na_range")) ) {
+    stop("The second (y) na_range is NULL, the first (x) is not.")
+  } else {
+    if ( any( attr(x, "na_range" ) != attr(y, "na_range" )) ) {
+      stop ("The na_range attribute must be equal.")
+    }
+  }
+
+  if ( is.null(attr(x, "na_values")) ) {
+    if ( ! is.null(attr(y, "na_values")) ) {
+      stop("The first (x) na_values is NULL, the second (y) is not.")
+    }
+  } else if ( is.null(attr(y, "na_values")) ) {
+    stop("The second (y) na_values is NULL, the first (x) is not.")
+  } else {
+    if ( any( attr(x, "na_values" ) != attr(y, "na_values" )) ) {
+      stop ("The na_values attribute must be equal.")
+    }
+  }
+
+  if ( ! is.labelled (x) && is.labelled(y) ) {
+    stop ("Both x and y must be labelled.")
+  }
+
[…]

### 2020-07-17 old_code/old_functions/convert_class.R

@@ -1,0 +1,225 @@
+#' @title Convert To Basic R Classes
+#'
+#' @param dat A data frame like object with the raw statistical data.
+#' @param metadata A metadata data frame created by
+#' \code{\link{gesis_metadata_create}}.
+#' @param var_name Defaults to \code{"normalized_name"}.
+#' @param conversion Defaults to \code{"conversion_suggested"}.
+#' @return The \code{df} converted to numeric, character, factor classes.
+#' Dummies are converted to binary numeric variables and their name is
+#' prefixed with \code{'is_'}.
+#' @importFrom tibble tibble
+#' @importFrom dplyr mutate filter select bind_cols
+#' @importFrom tidyselect all_of
+#' @importFrom purrr set_names
+#' @importFrom labelled to_factor to_character
+#' @examples
+#' \dontrun{
+#' require(dplyr)
+#' small_sample_convert  <- ZA5913_sample %>%
+#'   mutate ( filename = "ZA5913_sample") %>%
+#'   select (all_of(
+#'      c("uniqid", "filename", "doi", "p3", "qa10_3", 'w1'))
+#'      ) %>%
+#'   sample_n(20)
+#'
+#' small_sample_metadata <- gesis_metadata_create(
+#'                                 small_sample_convert )
+#'
+#' convert_class  ( dat = small_sample_convert,
+#'                  metadata = small_sample_metadata,
+#'                  var_name = "var_name_orig")
+#' }
+#' @importFrom purrr set_names
+#' @importFrom dplyr filter select mutate_all bind_cols
+#' @importFrom tidyselect all_of
+#' @importFrom labelled to_factor
+#' @family harmonization functions
+#' @export
+
+convert_class <- function(dat, metadata,
+                          var_name = "var_name_suggested",
+                          conversion = "conversion_suggested" ) {
+
+  . <- var <- valid_range <- length_total_range <- NULL
+  length_cat_range <- NULL
+
+  original_name_order  <- names ( dat )
+  class_conversion <- metadata %>%
+    dplyr::select ( all_of(c(var_name, conversion,
+                             "length_cat_range", "length_total_range") )) %>%
+    purrr::set_names(c("var", "conversion", "length_cat_range", "length_total_range"))
+
[…]

### 2020-07-17 old_code/old_functions/data-vocabulary_nuts1.R

@@ -1,0 +1,10 @@
+#' Vocabulary of NUTS1 regions.
+#'
+#' @format A data frame with 49 observations and four variables:
+#' \describe{
+#'   \item{country_code}{ISO 3166 country codes}
+#'   \item{region_nuts_names}{Region names in Eurobarometer files.}
+#'   \item{code10}{NUTS2010 regional codes}
+#'   \item{code13}{NUTS2013 regional codes}
+#'   }
+"vocabulary_nuts1"

### 2020-07-17 old_code/old_functions/data-vocabulary_nuts2.R

@@ -1,0 +1,10 @@
+#' Vocabulary of NUTS2 regions.
+#'
+#' @format A data frame with 293 observations and four variables:
+#' \describe{
+#'   \item{country_code}{ISO 3166 country codes}
+#'   \item{region_nuts_names}{Region names in Eurobarometer files.}
+#'   \item{code10}{NUTS2010 regional codes}
+#'   \item{code13}{NUTS2013 regional codes}
+#'   }
+"vocabulary_nuts2"

### 2020-07-17 old_code/old_functions/eurobarometer_labelled.R

@@ -1,0 +1,95 @@
+#' Construct A Variable Of Type 'eurobarometer_labelled'
+#'
+#' @param x A vector of numeric values of class integer, numeric,
+#' haven_labelled or haven_labelled_spss.
+#' @param value_labels The \code{labels} attribute, defaults to \code{NULL}.
+#' @param na_values The \code{na_values} attribute, defaults to \code{NULL}.
+#' @param na_range The \code{na_range} attribute, defaults to \code{NULL}.
+#' @param labels_orig The \code{labels_orig} attribute which contains the original
+#' value_labels before harmonization or normalization, defaults to \code{NULL}.
+#' @param var_name_orig The \code{var_name_orig} attribute, which
+#' contains the original variable name,  defaults to \code{NULL}.
+#' @param qb The \code{qb} attribute, which contains the identifier of the
+#' question block, defaults to \code{NULL}.
+#' @param conversion The \code{conversion} attribute which contains the
+#' name of the function that created the variable, defaults to \code{NULL}.
+#' @return A vector with metadata attributes of class \code{eurobarometer_labelled}.
+#' @examples
+#' eurobarometer_labelled(x = c(6,7,8,9,6,7,8),
+#'    labels =c("HI" = 6, "HELLO" = 7,
+#'              "HOWDY" = 8, "BYE" = 9),
+#'    na_values = 9,
+#'    qb = "Example Block",
+#'    conversion = "eurobarometer_labelled"
+#' )
+#' @export
+
+eurobarometer_labelled <- function ( x = NULL,
+                                     value_labels = NULL,
+                                     na_values = NULL,
+                                     na_range = NULL,
+                                     labels_orig = NULL,
+                                     var_name_orig = NULL,
+                                     qb = NULL,
+                                     conversion = NULL) {
+
+  if ( is.null(x) ) {
+    stop ( "Parameter x must be given.")
+  }
+
+  if (class(x)[1] =='logical' ) {
+    warning ("class <logical> is converted to <integer>.")
+    x <- as.integer(x)
+  }
+  if ( any( class(x) %in% c("integer", "numeric")) ) {
+    ebl <- haven::labelled_spss( x = x,
+                          labels = value_labels,
+                          na_values = na_values,
+                          na_range = na_range )
+  } else if ( any(class(x) %in% "haven_labelled_spss") ) {
+    ebl <- x
[…]

### 2020-07-17 old_code/old_functions/gesis_metadata_create_old.R

@@ -1,0 +1,145 @@
+#' Create A Metadata Table
+#'
+#' Create a metadata file from your surveys.
+#'
+#' @details The structure of the returned tibble:
+#' \describe{
+#'   \item{filename}{The original file name}
+#'   \item{qb}{Identifier of a question block for further tasks.}
+#'   \item{var_name_orig}{The original variable name in SPSS.}
+#'   \item{class_orig}{The original variable class after importing with\code{\link[haven]{read_spss}}.}
+#'   \item{var_label_orig}{The original variable label in SPSS.}
+#'   \item{var_label_norm}{A normalized version of the variable labels.}
+#'   \item{var_name_suggested}{A partly canonized variable name.}
+#'   \item{factor_levels}{A list of factor levels, i.e. value labels in SPSS}
+#'   \item{na_levels}{Values marked as missing by GESIS.}
+#'   \item{valid_range}{Not missing factor(category) levels.}
+#'   \item{class_suggested}{A suggested class conversion.}
+#'   \item{length_cat_range}{Number of categories in the non-missing range.}
+#'   \item{length_na_range}{Number of categories marked as missing by GESIS.}
+#'   \item{length_total_range}{Number of categories or unique levels, which may be different from the sum of missing and category labels.}
+#'   \item{n_categories}{Number of categories of the variable, should be the sum of the former two.}
+#' }
+#' @param survey_list A list of data frames containing surveys, or a
+#' single survey in a single data frame. The filename should be added
+#' in the column \code{filename}.
+#' @importFrom labelled val_labels var_label
+#' @importFrom dplyr full_join mutate
+#' @importFrom tibble tibble
+#' @importFrom tidyselect all_of
+#' @return A data frame with the original variable attributes and
+#' suggested conversions and changes.
+#' @examples
+#' import_file_names <- c(
+#'   'ZA7576_sample','ZA5913_sample'
+#' )
+#'
+#' my_surveys <- read_surveys (
+#'     import_file_names,
+#'     .f = 'read_example_file' )
+#'
+#' metadata <- gesis_metadata_create(my_surveys)
+#' @export
+
+gesis_metadata_create <- function ( survey_list ) {
+
+  ## if input is a data.frame, place it in a list -----------
+  if ( ! "list" %in% class(survey_list) ) {
+    survey_list <- to_survey_list( x = survey_list )
+  }
+
+  ## start of internal function metadata_create -------------
+  metadata_create <- function (dat) {
+
[…]

### 2020-07-17 old_code/old_functions/harmonize_qb_vars.R

@@ -1,0 +1,154 @@
+#' @title Harmonize A Question Block in Surveys
+#'
+#' Harmonize the variable names and the class of a question block.
+#' @param survey_list A list of data frames containing surveys.
+#' @param metadata A metadata table that contains \code{var_name_orig} and
+#' \code{qb} for the questions blocks, and \code{var_name_suggested} or
+#' its user defined alternative, \code{class_suggested} or its user defined
+#' alternative.
+#' @param question_block A user defined question block, or pre-defined
+#' question blocks: \code{"id"}, \code{"metadata"}, \code{"weights"},
+#' \code{"socio-demography"}.
+#' @param id_vars A vector of ID variables to form a panel ID var,
+#' defaults to \code{c("uniqid", "doi")}.
+#' @param var_name Defaults to \code{var_name_suggested}. Must be
+#' a column in \code{metadata}.
+#' @param conversion Defaults to \code{conversion_suggested}. Must be
+#' a column in \code{metadata}.
+#' @importFrom dplyr select filter bind_cols distinct_at distinct
+#' @importFrom purrr set_names
+#' @importFrom magrittr %>%
+#' @importFrom tidyselect all_of
+#' @return A tibble containing all variables with harmonized names and
+#' classes, and a unique key column named \code{panel_id}.
+#' @family harmonization functions
+#' @examples
+#' \dontrun{
+#' harmonize_qb_vars (
+#'    survey_list,
+#'    metadata,
+#'    question_block = "socio-demography",
+#'    var_name = "var_name_suggested",
+#'    conversion = "conversion_suggested"
+#'    )
+#' }
+#' @export
+
+harmonize_qb_vars <- function ( survey_list,
+                                metadata,
+                                id_vars = c("uniqid", "doi"),
+                                question_block = "socio-demography",
+                                var_name = "var_name_suggested",
+                                conversion = "conversion_suggested" ) {
+
+  id_vars_input <- id_vars
+  id_vars <- qb <- survey_subset <- var_name_orig <- NULL
+  . <- panel_id <- filename <- NULL
+
+  if ( ! "list" %in% class(survey_list) ) {
+    survey_list <- to_survey_list(survey_list)
+  }
+
+  if ( ! var_name %in% names(metadata) ) {
+    stop( var_name , " is not in <metadata>." )
+  }
+
[…]

### 2020-07-17 old_code/old_functions/test-convert_class.R

@@ -1,0 +1,25 @@
+library(testthat)
+library(dplyr)
+data ("ZA5913_sample")
+
+small_sample_convert  <- ZA5913_sample %>%
+  mutate ( filename = "ZA5913_sample") %>%
+  select (all_of(c("uniqid", "filename", "doi", "p3", "qa10_3", 'w1'))) %>%
+  dplyr::sample_n(20)
+small_sample_metadata <- gesis_metadata_create( small_sample_convert )
+
+test_converted <- convert_class  ( dat = small_sample_convert,
+                   metadata = small_sample_metadata,
+                   var_name = "var_name_orig")
+
+test_that("Correct class conversion takes place", {
+  expect_equal(class(test_converted$filename), "character")
+  expect_equal(class(test_converted$w1), "numeric")
+  expect_equal(class(test_converted$p3), "character")
+  expect_equal(class(test_converted$qa10_3), c("haven_labelled", "vctrs_vctr", "double"))
+  expect_equal(sort(names(attr ( test_converted$qa10_3, "labels"))),
+  c("decline", "tend_not_to_trust", "tend_to_trust"))
+  expect_equal(sort(names(attr ( test_converted$qa10_3, "labels_orig"))),
+               c("DK", "Tend not to trust", "Tend to trust"))
+})
+

### 2020-07-17 old_code/old_functions/test-eurobarometer_labelled.R

@@ -1,0 +1,53 @@
+library(testthat)
+
+### Test wrong parameter classes ---------------------------------
+
+# "Hello, Hi, Hey":
+# https://open.spotify.com/track/35J5fWBYFx1LTut0JwXxgh?si=pspMB7AfT2G808_wP3Ysbw
+
+test_that("Appropriate warnings are given", {
+  expect_warning(eurobarometer_labelled (x = c(TRUE, TRUE, FALSE, NA))  )
+})
+
+test_that("Errors are found", {
+  expect_error(eurobarometer_labelled ( x = data.frame ( x = 1:5)))
+})
+
+### Test if the constructor works as expected -----------
+v <- labelled::labelled(c(3,4,4,3,8, 9),
+                        c(YES = 3, NO = 4, `WRONG LABEL` = 8, REFUSED = 9)
+)
+correct_labels <- c(3,4,8,9)
+names(correct_labels) <- c("YES", "NO", "WRONG LABEL", "REFUSED")
+correct_yes_no <- c(1,0)
+names(correct_yes_no) <- c("yes", "no")
+
+
+eurobarometer_labelled(x = c(6,7,8,9, 6,7,8),
+                       labels =c("HI" = 6, "HELLO" = 7,
+                                 "HOWDY" = 8, "BYE" = 9),
+                       na_values = 9,
+                       qb = "Example Block",
+                       conversion = "eurobarometer_labelled")
+
+x <-haven::labelled_spss(
+  x = c(1,2,2,1,99998,99999),
+  labels =  c(yes = 1, no = 2, wrong_label = 99998, refused = 99999),
+  na_values = NULL,
+  na_range = 99998:99999,
+  label = "Variable Example"
+)
+
+test_that("constructor works", {
+  expect_equal(labelled::val_labels ( eurobarometer_labelled(v)),
+               correct_labels)
+  expect_equal(labelled::val_labels (
+    eurobarometer_labelled(x = c(0,1),
+                           value_labels = c( 'yes' = 1, 'no' = 0)
+                           )),
+    correct_yes_no)
+  expect_equal(labelled::na_range (eurobarometer_labelled(x)),
+               c(99998,99999))
+  expect_equal(class ( eurobarometer_labelled(x)),
+               c("eurobarometer_labelled", class(x)))
+})

### 2020-07-17 old_code/old_functions/test-harmonize_to_numeric.R

@@ -1,0 +1,11 @@
+v <- labelled::labelled(c(3,4,4,3,8,9),
+             c(YES = 3, NO = 4, wrong_label = 8, refused = 9)
+             )
+
+input_labelled_var = harmonize_value_labels(v,2)
+
+test_that("correct values are retured", {
+  expect_equal(harmonize_to_numeric(
+    harmonized_labelled_var = input_labelled_var),
+               c(1,0,0,1,NA_real_, NA_real_))
+})

### 2020-07-17 old_code/old_vignettes/create_tabels.R

@@ -1,0 +1,299 @@
+## ---- include = FALSE---------------------------------------------------------
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+## ----setup, message=FALSE, warnings=FALSE-------------------------------------
+library(eurobarometer)
+library(dplyr)
+library(tibble)
+library(knitr)
+library(kableExtra)
+# The examples of this vignette can be found run with  
+# source(
+#  file.path("not_included", "vignette_vocabulary_examples.R")
+#  )
+
+## -----------------------------------------------------------------------------
+metadata_database <-  readRDS( 
+  file.path('..', 'data-raw',
+  'eb_metadata_database_large.rds')) 
+
+## -----------------------------------------------------------------------------
+select_metadata_vars <- c(
+  "filename", "var_name_orig", "var_label_orig",
+  "var_label_norm", "val_label_orig", "val_label_norm", 
+  "val_numeric_orig", "val_order_alpha", "n_categories")
+
+trust_metadata <- metadata_database  %>%
+  filter (
+    grepl( "trust", var_label_norm )
+  )  %>%
+  select ( all_of(select_metadata_vars) ) %>%
+  arrange ( var_label_norm, val_label_norm, filename )
+
+trust_metadata %>%
+  head(10) %>%
+  kable %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 )
+
+## -----------------------------------------------------------------------------
+trust_metadata <- trust_metadata %>% 
+  filter (
+    grepl( 
+      "trust_in_institutions|trust_political_parties|_trust$",
+      var_label_norm )
+  ) %>%
+  filter( ! (filename == "ZA4744_v5-0-0.sav" &
+               var_name_orig %in% c("v336", "v347","v292", "v303",
+                                    "v314", "v325", "v270", "v281")
+             )
+          )
+
+trust_metadata %>%
+  arrange(filename, var_name_orig, val_order_alpha) %>%
+  head(10) %>%
+  kable %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 )
+
+## -----------------------------------------------------------------------------
+trust_metadata %>%
+  arrange(val_label_norm) %>%
+  select(val_label_norm, everything()) %>% 
+  count(n_categories) %>%
+  kable %>%
+  kable_styling(bootstrap_options =
[…]

### 2020-07-17 old_code/old_vignettes/variable_names.R

@@ -1,0 +1,26 @@
+## ---- include = FALSE---------------------------------------------------------
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>"
+)
+
+## ----setup, message=FALSE, warnings=FALSE-------------------------------------
+library(eurobarometer)
+library(knitr)
+library(kableExtra)
+library(dplyr)
+
+## -----------------------------------------------------------------------------
+data("categorical_variables_2", package = "eurobarometer")
+
+categorical_variables_2 %>%
+  kable() %>%
+  kable_styling(bootstrap_options = 
+                  c("striped", "hover", "condensed"), 
+                  fixed_thead = T, 
+                  font_size = 10 ) %>%
+    add_header_above(c("Filtering" = 3, 
+                       "Preferred Term" = 1, 
+                       "Keywords" = 2)
+                     )
+

### 2020-07-17 old_code/old_vignettes/vocabulary.R

@@ -1,0 +1,196 @@
+## ---- include = FALSE---------------------------------------------------------
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE, 
+  eval=FALSE,
+  collapse = TRUE,
+  comment = "#>"
+)
+
+## ----setup--------------------------------------------------------------------
+#  library(eurobarometer)
+#  library(dplyr)
+#  library(tibble)
+#  library(knitr)
+#  library(kableExtra)
+#  # The examples of this vignette can be found run with
+#  # source(
+#  #  file.path("not_included", "vignette_vocabulary_examples.R")
+#  #  )
+
+## -----------------------------------------------------------------------------
+#  metadata_database <- readRDS(
+#    file.path("..", "data-raw", "eb_metadata_database.rds")
+#    )
+
+## -----------------------------------------------------------------------------
+#  select_metadata_vars <- c("filename", "var_name_orig",
+#                            "var_label_norm",
+#                            "var_name_suggested",
+#                            "val_label_norm" , "val_label_orig",
+#                            "val_order_alpha", "val_order_length")
+#  
+#  trust_metadata <- metadata_database  %>%
+#    filter (
+#      grepl( "tend_to_trust|tend_not_to_trust", val_label_norm )
+#    )  %>%
+#    select ( all_of(select_metadata_vars)) %>%
+#    arrange ( var_label_norm, val_label_norm, filename )
+
+## -----------------------------------------------------------------------------
+#  require(kableExtra)
+#  trust_metadata %>% filter (
+#     grepl( "tend_to_trust", val_label_norm )
+#    ) %>%
+#    sample_n(15) %>% # Print only a sample of 15 rows
+#  kable() %>%
+#    kable_styling(bootstrap_options =
+#                  c("striped", "hover", "condensed"),
+#                  fixed_thead = T,
+#                  font_size = 7)
+
+## -----------------------------------------------------------------------------
+#  trust_metadata %>%
+#    select (
+#      # previously we filtered tend to (not) agree, remove it
+#      -all_of(c("val_label_norm", "val_label_orig"))
+#      ) %>%
+#    left_join (
+#      # add back all value labels, i.e. various declines
+#      metadata_database %>%
+#                  select ( all_of(select_metadata_vars))
+#      )  %>%
+#    sample_n(15) %>% # Print only a sample of 15 rows
+#  kable() %>%
[…]

### 2020-07-17 old_code/old_vignettes/workflow.R

@@ -1,0 +1,180 @@
+## ---- include = FALSE---------------------------------------------------------
+knitr::opts_chunk$set(
+  eval = FALSE,
+  collapse = TRUE,
+  comment = "#>"
+)
+example_panel_id <- data.frame(
+  panel_id = rep(NA_character_, 8),
+  id_uniqid = c(101,102,103,104, 108,199,104,188), 
+  id_survey = c( rep("eb_23.4", 4), 
+                 rep("eb_28.9", 4)),
+  filename = c( rep("GESIS_A.sav", 4), 
+                rep("GESIS_B.sav", 4)), 
+  w1 =  round(runif(8, 0, 1.5),2),
+  wex = round(runif(8, 0, 1.5)*runif(8, 1000,9000),2)
+  )
+example_panel_id$panel_id <- paste0(
+  example_panel_id$id_survey, "_", 
+  example_panel_id$id_uniqid, "_",
+  example_panel_id$filename
+)
+
+my_panel <- example_panel_id
+
+## ----setup, eval=TRUE, echo=TRUE----------------------------------------------
+## This vignette is hypothetical at this stage, and the chunks, 
+## unless stated otherwise, are NOT evaluated.
+library(eurobarometer)
+library(knitr)
+
+## -----------------------------------------------------------------------------
+#  ## read in all GESIS SPSS files and validated the
+#  ## reading process, save only those that are valid GESIS files.
+#  ## must canonize some id vars, such as uniqid, and filename
+#  ##or anything that is present in all files.
+#  
+#  gesis_spss_to_rds ( import_dir, working_dir )
+
+## -----------------------------------------------------------------------------
+#  ## filter out the most basic and omnipresent id variables, and the
+#  ## most basic weights, w1 and its projected version wex
+#  
+#  my_panel <- panel_create ( working_dir,
+#                 # do not start with id_ to make filtering other IDs
+#                 # easier later
+#                            panel_id = "panel_id",
+#                 ## must be at least two, and one must be the uniqid
+#                 ## of the file or row_id
+#                            id_components = c("uniqid", "filename")
+#                 )
+
+## ---- eval=TRUE---------------------------------------------------------------
+kable(my_panel)
+
+## -----------------------------------------------------------------------------
+#  ## read in all GESIS SPSS files and validated the reading process,
+#  ## save only those that are valid GESIS files.
+#  working_metadata <- gesis_metadata_create ( working_dir )
+#  
[…]

### 2020-07-17 R/data_eurobarometer_corpus.R
delete broken links fromm docs

@@ -1,7 +1,6 @@
 #' GESIS Eurobarometer Variable Labels
 #'
-#' Variable labels used for the harmonization of questionnaire items,
-#' created by \code{\link{gesis_vocabulary_create}}.
+#' Variable labels used for the harmonization of questionnaire items.
 #'
 #' @format A data_frame:
 #' \describe{
@@ -13,7 +12,6 @@
 #'   \item{label_normalized}{Normalized version of the questionnaire item}
 #'   \item{filename}{The name of the original GESIS file.}
 #' }
-#' @seealso gesis_vocabulary_create
 "eurobarometer_corpus"
 
 

### 2020-07-17 R/eurobarometer.R

@@ -7,7 +7,8 @@
 #' @section labelling functions:
 #' The naming functions make the GESIS SPSS files usable in a programmatic
 #' context.
-#' \code{\link{label_normalize}} removes regex symbols, whitespace, and
+#' \code{\link{var_label_normalize}} and \code{\link{val_label_normalize}}
+#' removes regex symbols, whitespace, and
 #' basic inconsistencies in abbreviations.
 #' \code{\link{label_suggest}} harmonizes the names across various GESIS
 #' files, so that they can be joined into panels across time.
@@ -16,24 +17,6 @@
 #' The metadata functions map the SPSS file and suggest variable names,
 #' re-labelling and class conversions.
 #' \code{\link{gesis_metadata_create}}
-#' \code{\link{gesis_vocabulary_create}}
-#'
-#' @section harmonization functions:
-#' \code{\link{id_create}} creates truly unique IDs, and its wrapper
-#' \code{\link{panel_create}} creates a skeleton panel with unique
-#' \code{panel_id} and other identification variables.
-#' \code{\link{convert_class}} to make class conversions based on a
-#' metadata analysis.
-#' \code{\link{harmonize_qb_vars}} is harmonizing the names and classes of
-#' questions in a question block.
-#' \code{\link{harmonize_missing_values}} is harmonizing the missing values
-#' in a labelled vector.
-#' \code{\link{harmonize_value_lables}} is harmonizing the the value
-#' labels in 2 or 3 category labelled variables.
-#' \code{\link{eurobarometer_special_values}} returns the harmonized
-#' numeric codes and harmonized labels of special values in the
-#' Eurobarometer surveys, for example \code{"inap"}, \code{"declined"} or
-#' or \code{"still_studying"}.
 #'
 #' @section sample data files:
 #' \code{\link{ZA5913_sample}} A small subset of the ZA5913:

### 2020-07-17 R/gesis_vocabulary_create.R

@@ -1,179 +1,0 @@
-#' Get The Vocabulary Of A Survey
-#'
-#' Get every single unique questionnaire item from a survey file.
-#'
-#' @details The variables in the returned data frame
-#' \describe{
-#'   \item{var_label_orig}{The name of the variable in the data file.}
-#'   \item{val_numeric_orig}{The original SPSS numeric values, if applicable}
-#'   \item{val_label_orig}{The original SPSS character value, if applicable}
-#'   \item{val_order_alpha}{The rank number of the alphabetically sorted answer option.}
-#'   \item{val_order_length}{The number of answer options in the questionnaire item.}
-#'   \item{val_label_normalized}{Normalized version of the questionnaire item}
-#' }
-#' @param survey_list A list of data frames containing surveys, or a
-#' single survey in a single data frame. The filename should be added
-#' in the column \code{filename}.
-#' @importFrom dplyr bind_rows mutate arrange
-#' @importFrom labelled var_label to_character
-#' @importFrom tibble tibble
-#' @seealso label_suggest
-#' @return The vocabulary of the survey in a data frame. The return
-#' data frame has 6 columns.
-#' @examples
-#' import_file_names <- c(
-#'   'ZA7576_sample','ZA5913_sample'
-#' )
-#'
-#' my_surveys <- read_surveys (
-#'     import_file_names,
-#'     .f = 'read_example_file' )
-#'
-#' my_vocabulary <- gesis_vocabulary_create(my_surveys)
-#' @export
-
-gesis_vocabulary_create <- function ( survey_list ) {
-
-  ## if input is a data.frame, place it in a list -----------
-  if ( ! "list" %in% class(survey_list) ) {
-    survey_list <- to_survey_list( x = survey_list )
-  }
-
-  ## start internal function vocabulary_create --------------
-  vocabulary_create <- function(dat) {
-    class_orig   <- vapply (
-      dat, function(x) class(x)[1], character(1)
-    )
-
-    get_items <- function ( this_var_name_orig  = 'nuts' ) {
-
-      . <- val_label_orig <- val_numeric_orig <- NULL
-
-      itemize <- dat[, this_var_name_orig ]
-      names(itemize) <- "item"
-      items <- unique(itemize$item)
-
-      numeric_value <- rep(NA_real_, length(items))
-      character_value <- rep(NA_character_, length(items))
-
[…]

### 2020-07-17 R/read_rds.R
vignette changes

@@ -41,7 +41,7 @@ read_rds <- function(file, id = NULL, doi = NULL) {
     doi <- tmp$doi[1]
     }
 
-  amend_survey(tmp)
+  amend_survey(tmp,filename = filename ,doi=doi)
 
 }
 
@@ -85,14 +85,14 @@ read_spss <- function(file,
     doi = doi )
 
 
-  amend_survey(tmp, filename = filename)
+  amend_survey(tmp, filename = filename, doi = doi)
 
 
 }
 
 #' @importFrom retroharmonize read_spss survey
 #' @keywords internal
-amend_survey <- function (survey, filename) {
+amend_survey <- function (survey, filename, doi = doi) {
   tmp <- retroharmonize::survey(
     df = survey, id = id, filename = filename, doi = doi
   )

### 2020-07-18 data-raw/variable-groups.Rmd
variable groups from ZACAT for 157 waves, and description of their merging with metadata from 97 files

---
title: "Variable groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vocabulary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
options(width = 160)
```

```{r setup}
library(eurobarometer)
library(rio)
library(tidyverse)
library(tibble)
library(knitr)
library(kableExtra)
```

`eb_var_groups_zacat.rds` contains info on variable groups for 157 EB waves from 44.2bis (ZA2828) to 92.1 (ZA7579), copied from [GESIS ZACAT](https://zacat.gesis.org/webview/).

The columns are mostly self-explanatory:

`wave` = wave number (e.g., 72.1)    
`archive_id` = GESIS archive id (e.g., ZA2828)    
`group` = variable group as identified by GESIS ZACAT (e.g., "Protocol variables") , slightly cleaned to exclude variable names/ranges in parentheses following group name   
`var_label_zacat` = variable label as presented in GESIS ZACAT (e.g., "EU COMMON CURRENCY - KNOW NAME"), slightly cleaned to exclude question number if it occurs at the beginning of the value label - more on that below     
`wave_desc` = wave number and description / title (e.g., "Eurobarometer 85.1OVR (April 2016): European Youth in 2016")  
`timeframe` = timing of the wave (e.g., "April 2016"), not always available  
`description` = wave description / title (e.g., "European Youth in 2016")

```{r}
eb_var_groups_zacat <- readRDS( file.path('../data-raw', 'eb_var_groups_zacat.rds') )

eb_var_groups_zacat %>%
  select(wave, archive_id, group, var_label_zacat) %>%
  head() %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Some examples of how the groups look like:

```{r}
eb_var_groups_zacat %>%
  distinct(wave, group) %>%
  head(., 25) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

Most common groups. Note that some groups changed names, e.g. "Archive and survey ID variables" and "Archive and ID variables". They could be standardized across all waves.

```{r}
eb_var_groups_zacat %>%
  count(wave, group) %>%
  count(group) %>%
  arrange(desc(n)) %>%
  head(., 25) %>%
  kable %>%
  kable_styling(bootstrap_options =
                  c("striped", "hover", "condensed"),
                  fixed_thead = T,
                  font_size = 10 )
```

`eb_metadata_var_groups.rds` is created by joining `eb_var_groups_zacat.rds` with `eb_metadata_database_20200628.rds` after some cleaning of `var_label_orig`.

The reason is that, in GESIS ZACAT, the variable labels (`var_label_zacat`) sometimes don't include the question number at the beginning (in the recent waves) and sometimes, especially in the early waves, they do (e.g., `Q25 EU COMMON CURRENCY - KNOW NAME`). So, for matching with metadata from the data files, these question numbers have been eliminated, but they also need to be eliminated from `var_label_orig` in the metadata files.

Note: Variable labels in GESIS ZACAT for waves 85.3 (ZA6695) and 86.1 (ZA6697) are for some reason different than in the data and can't be joint easily. This is why, for these two waves some variables don't have groups assigned. Matching variable lables from these two waves would require some better regex magic or manual work.

The resulting file is for 96 waves, the intersection of `eb_metadata_database_20200628.rds` (which has 97 files, but one is not from the Standard Eurobarometer) and `eb_var_groups_zacat.rds` (which has 157 waves).

```{r}
# waves included in eb_var_groups_zacat
archive_ids <- eb_var_groups_zacat %>% distinct(archive_id) %>% pull(archive_id)

metadata_database <- readRDS( file.path('../data-raw', 'eb_metadata_database_20200628.rds')) %>%
  mutate(archive_id = substr(filename, 1, 6))

eb_metadata_var_groups <- metadata_database %>%
  # filter only waves for which variable groups are available
  filter(archive_id %in% archive_ids,
         # exclude this strange variable
         var_label_orig != "filename") %>%
  # keep one row per variable
  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm, archive_id) %>%
  # original variable labels in the data and in GESIS ZACAT are not exactly the same
  # --> need some cleaning
  mutate(var_label_orig2 = gsub("^([A-Z])*[0-9]{1,3}([A-Z]{1,2})* ", "",
                                var_label_orig),
         var_label_orig2 = gsub("^[A-Z][0-9] [0-9]{1,2} ", "",
                                var_label_orig2),
         var_label_orig2 = gsub("^W3A/W4A ", "", var_label_orig2),
         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2}([A-Z])*_[A-Z]{1,2} ", "",
                                var_label_orig2),
         var_label_orig2 = gsub("^\\s+|\\s+$", "",
                                var_label_orig2),
         var_label_orig2 = gsub('“)', "",
                                var_label_orig2),
         var_label_orig2 = gsub("  ", " ",
                                var_label_orig2),
         # these 2 waves have problems and labels in data and in ZACAT are different
         var_label_orig2 = ifelse(archive_id %in% c("ZA6695", "ZA6697"),
                                  gsub("\\s*\\([^\\)]+\\)","", var_label_orig2),
                                  var_label_orig2)) %>%
  # join with table with variable groups
  left_join(eb_var_groups_zacat, by = c("var_label_orig2" = "var_label_zacat", "archive_id"))

# saveRDS(eb_metadata_var_groups, file.path('../data-raw', 'eb_metadata_var_groups.rds') )
```

### 2020-07-20 README.Rmd

@@
 # install.packages("devtools")
+devtools::install_github("antaldaniel/retroharmonize")
 devtools::install_github("antaldaniel/eurobarometer")
 ```
+
+You do not need to load the [retroharmonize](http://retroharmonize.satellitereport.com/) package with `library(retroharmonize)` but eurobarometer relies on some if its classes, methods and functions.
 
@@
 1.
-We import the GESIS SPSS files with the help of the [haven](https://haven.tidyverse.org/) package.
-In order to preserve the metadata in the files (particularly variable- and value labels, and user-defined missing values) these are converted to labelled classes.
+We import the GESIS SPSS files with the help of the adopted version of `read_spss()` from the [haven](https://haven.tidyverse.org/) package.
+The `haven::write_sav()` function does not work perfectly, it loses some metadata, so the best is to save you data as an `.rds` file.
+You can read surveys then with `read_rds()`.
+Both importing function will try to add important metadata to the `retroharmonize::survey()` tibble/data.frame class, including the original file name, the survey's ID and DOI.
 
@@
 We analyse the files with `gesis_metadata_create()` to understand what are the possible problems in the SPSS file and how to resolve them.
-(See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html) or `vignette("metadata")`)
+(See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html) or `vignette("metadata")`) This is a modified version of the more generic [metadata_create](http://retroharmonize.satellitereport.com/reference/metadata_create.html) function in `retroharmonize`.
+See the [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html) vignette for some more details.
 
@@
 4.
-We convert the variables to a modified version of the `haven_labelled_spss` class that contains the harmonized values, harmonized labels, harmonized missing values codes of the variable with their history (original values, labels, and function applied to change them)  for reproducibility.
-The [eurobarometer_labelled](http://eurobarometer.danielantal.eu/articles/eurobarometer_class.html) class inherits most of the functionality of the packages [haven](https://haven.tidyverse.org/),  [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) and are expected to work well with [sjlabelled](https://strengejacke.github.io/sjlabelled/articles/labelleddata.html).
+We convert the variables to a modified version of the `retroharmonize_labelled_spss_survey` class that contains the harmonized values, harmonized labels, harmonized missing values codes of the variable with their history (original values, labels, and function applied to change them)  for reproducibility.
+The [labelled_spss_survey](http://retroharmonize.satellitereport.com/articles/labelled_spss_survey.html) class inherits most of the functionality of the packages [haven](https://haven.tidyverse.org/),  [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) and are expected to work well with [sjlabelled](https://strengejacke.github.io/sjlabelled/articles/labelleddata.html).
+
+5.
+The harmonization of the variable names is currently possible with retroharmonize only (See: [Harmonize Value Labels](http://retroharmonize.satellitereport.com/articles/harmonize_labels.html)).
+Further, Eurobarometer specific helpers will be added soon.
+See the article [Survey Data Harmonization](http://eurobarometer.danielantal.eu/articles/harmonization.html) as a case study.
 

### 2020-07-22 data-raw/article.Rmd

@@
 author: "Daniel Antal, Marta Kolczynska"
-date: "7/10/2020"
+date: "7/22/2020"
 bibliography: [library.bib, packages.bib]
@@
 
-<Here more information about the EUrobarometer, the contents, frequency of waves, amount of data, fieldwork and number of respondents.
-Maybe even the paragraph from the introduction would fit better here than there>
+The Eurobarometer is a biannual survey conducted by the European Commission with the goal of monitoring the public opinion of populations of EU member states and -- occasionally -- also in candidate countries.
+The surveys are administered via face-to-face interviews to samples (selected via multi-stage probability sampling) of adult populations of the respective countries.
+Sample sizes hover around 1000 respondents in most countries, and around 500 respondents in smaller nations.
+Target populations in the Eurobarometer are defined as the “[P]opulation of the respective nationalities of the European Union Member States and other EU nationals, resident in each of the 28 Member States and aged 15 years and over“.
+In non-EU countries, the target population includes “the national population of citizens and the population of citizens of all the European Union Member States that are residents in these countries and have a sufficient command of the national languages to answer the questionnaire“ [@EuropeanCommission2017].
+
+Each EB wave is devoted to a particular topic, and most waves ask some "trend questions", i.e.
+questions that are repeated frequently in the same form.
+Especially these trend questions create unique opportunities for longitudinal research.
+Each wave is published as a separate file that contains responses from all countries that participated in that wave, i.e.
+all EU countries and sometimes also candidate countries.
+As a consequence, the data are ex-ante harmonized within waves, which makes cross-country comparisons readily possible, and also reduces the burden of ex-post harmonization by  a factor of about 20 (the number of countries in an average EB wave).
+
+TO given an idea of the size of the EB data sets, each file consists of between around 150 and 4000 variables, with an average of some 750 among the test files we used, and between 15,000 and around 65.000 cases corresponding to individual respondents.
 
@@
 Thus, in this case the harmonization of the data across multiple Eurobarometer surveys boils down to data data transformation and documentation.
-For the purposes of presenting the process of harmonization with `retroharmonize` and eurobarometer packages, we use the case of the Eurobarometer without the loss of generality of the described procedures, since it includes all data processing steps involved in harmonization.
+For the purposes of presenting the process of harmonization with `retroharmonize` and `eurobarometer` packages, we use the case of the Eurobarometer without the loss of generality of the described procedures, since it includes all data processing steps involved in harmonization.
 
@@
 
-We should keep an eye on [snakecase](#https://github.com/Tazinho/snakecase) and maybe janitor, too.
+While many EB waves include the same trend questions, the naming conventions of these variables and their variable labels have been changing and are not consistent across all waves.
+For example, questions about trust in parliament are labelled as "TRUST IN INSTITUTIONS: NAT PARLIAMENT" or "TRUST IN INSTITUTIONS: NATIONAL PARLIAMENT", where the label also typically contains at the beginning the number of the question in the questionnaire, which changes from wave to wave.
+Similarly, the question about Respondent's occupation is sometimes labelled as "RESPONDENT OCCUPATION SCALE" and other times "OCCUPATION OF RESPONDENT SCALE".
+
+Identifying the same variables behind these different labels can be automated in cases when the labels contain the same words but in a different order or separated by different punctuation or stop words.
+
+Another part of the standardization of variable names can be performed with a dictionary that maps abbreviations (often non-standard ones due to limits in the length of variable labels) to standardized terms.
+For example, 'gov', 'govnmt', and 'govmnt' are all variants of the term 'government'.
+
+Only applying these two steps enable the standardization of a number of variables referring to socio-demographic characteristics (age, gender, education, occupation, household size, marital status, community type), well-being (life satisfaction), political attitudes and engagement (left-right placement, trust in institutions, discussing politics), and living standard (ownership of durables, internet use).
+
+**Note:** We should keep an eye on [snakecase](#https://github.com/Tazinho/snakecase) and maybe janitor, too.
 

### 2020-07-22 data-raw/article.Rmd
first few paragraphs - technology overview and grammar

@@
 knitr::write_bib(c(
-  .packages(), 'haven', 'labelled', 'vctrs'
+  .packages(), 'haven', 'labelled', 'vctrs', 'retroharmonize'
 ), 'packages.bib')
@@
 The increasing availability of microdata creates new opportunities for joint exploitation of these resources for comparative research.
-Before joint analysis is possible, the data need to be harmonized, i.e.
-transformed and combined in a single data set in a way that matches the measurement of the same characteristics and standardizes the coding of responses.
-We introduce the `retroharmonize` and `eurobarometer` packages that help the reproducible ex-post harmonization of survey data stored in various, individual files with rich metadata.
+Before joint analysis is possible, the data needs to be harmonized.
+This means a transformation and combination in a way that the same measurements are coded in the same way.
+
+We introduce the `retroharmonize` and `eurobarometer` packages that help the reproducible, metadata-rich ex-post harmonization of survey data stored in various individual files.
 We illustrate the proposed workflow using the case of the Eurobarometer surveys, which provide rich data on European publics in form of single-wave SPSS files from a period of over 25 years.
-Our tools are flexible and can be applied to other survey projects.
+Our tools are flexible and can be applied to other survey projects: we also introduce [a small case study](http://retroharmonize.satellitereport.com/articles/afrobarometer.html) with Afrobarometer.
 
@@
 
-* Define a common [taxonomy](#taxonomy), probably based on a thesaurus for consistentcy.
+* Define a common [taxonomy](#taxonomy) based on a thesaurus for consistency.
 
@@
 
---	Creating a [joint data table](#datajoin) where each variable that represents the same class and metadata attributes; documenting all the changes and 	Providing methods that re-export the data files to SPSS, Stata, SAS.
+--	Creating a [joint data table](#datajoin) where each variable that represents the same class and metadata attributes; documenting all the changes and providing methods that re-export the data files to SPSS, Stata, SAS.
 
@@
 
-Our workflow generally fits into the tidyverse.
+The workflow that we created fits into the [tidyverse](https://www.tidyverse.org/), and relies on [dplyr](https://dplyr.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/) for data processing, [haven](https://haven.tidyverse.org/) for importing,  [tibble](https://tibble.tidyverse.org/) for more modern data framing.
+
+For easier further development, we rely on the r-lib family of [rlang](https://rlang.r-lib.org/) for easier exception handling, [vctrs](https://vctrs.r-lib.org/) for type casting and coercion, and implementation of class methods, and [fs](https://fs.r-lib.org/) for platform-independent file management.
+Of course, the tidyverse relies on these modern re-implementations of base R and they improve the tidy workflow and benefit from the efficiency gains with of a basic C++ infrastructure.
 

### 2020-07-22 data-raw/article.Rmd
citations

@@
 knitr::write_bib(c(
-  .packages(), 'haven', 'labelled', 'vctrs', 'retroharmonize'
+  .packages(), 'haven', 'labelled', 'vctrs', 'retroharmonize', 'eurobarometer'
 ), 'packages.bib')
@@
 
-We introduce the `retroharmonize` and `eurobarometer` packages that help the reproducible, metadata-rich ex-post harmonization of survey data stored in various individual files.
+We introduce the `retroharmonize` [@R-retroharmonize] and `eurobarometer` [@R-eurobarometer] packages that help the reproducible, metadata-rich ex-post harmonization of survey data stored in various individual files.
 We illustrate the proposed workflow using the case of the Eurobarometer surveys, which provide rich data on European publics in form of single-wave SPSS files from a period of over 25 years.

### 2020-07-26 data-raw/variable-keywords-table.Rmd
just one space
one comment
updates variable name harmonization with label types

@@
 ---
-title: "Tagging variables with keywords"
+title: "Harmonizing variables names"
 output: rmarkdown::html_vignette
@@
 
-```{r setup}
+```{r setup, warning = F, error = F}
 library(eurobarometer)
+library(rio)
 library(tidyverse)
@@
 
-An important aspect of survey data harmonization - in addition to standardizing response coding and value labels - is the identification of variables that measure the same thing, and standardizing the names of these variables.
+An important aspect of survey data harmonization -- in addition to standardizing response coding and value labels -- is the identification of variables that measure the same thing in the same way, and standardizing the names of these variables.
 The process is faciliated by the fact that variables typically have labels that succintly describe the content of the variable, and sometimes also contain information about the coding of responses (e.g., length of scale), the unit (e.g., age in years or categories) and whether the variable contains "original" respondent responses or some transformation of responses (e.g.
 recoding or index).
+Variable labels, on the other hand, provide information about the response options, including their number (corresponding to the number of substantive response options) and content.
 
-Fortunately, in many multi-wave surveys, such as the Eurobarometer, there is a degree of consistency in how variables are labelled in different editions of the survey, and most survey projects try to follow some labelling conventions.
+Thus, there is quite a lot of information in the variable metadata that can be used to identify:
 
-Unfortunately, consistency is limited and conventions are not followed strictly, which makes standardizing labels a semi-manual task.
-Some of these issues stem from the limits in the number of characters of variable labels in common statistical software packages (e.g., SPSS), which leads to non-standard abbreviations.
+1. variables that have the same or very similar **variable labels**,
 
-This vignette outlines a procedure for arriving at standardized variable names via tokenization of normalized variable labels (`var_label_norm`), which are created by [`label_normalize`](http://eurobarometer.danielantal.eu/reference/label_normalize.html) as part of [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html).
+2.
+variables of the same type in terms of item design in terms of **value labels** - especially items of the "mentioned / not mentioned" type that are used to represent questions with multiple items where the Respondent chooses more than one option, but also -- occasionally -- questions, where the Respondent chooses only one option.
+
+Thanks to variable and value labels, it becomes possible to identify variables that measure the same concept in the same way in different EB waves without studying the documentation.
+
+This vignette outlines these procedures, including the tokenization of normalized variable labels (`var_label_norm`), which are created by [`label_normalize`](http://eurobarometer.danielantal.eu/reference/label_normalize.html) as part of [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html).
 The logic is similar to text analysis, in that a document-term matrix is created with each variable label treated as a (very short) document, and terms treated as keywords.
-variables that contain the same sets of keywords can be - with some margin of error - assumed to correspond to the same or equivalent survey questions.
 
@@
 
-0. Read survey and variable metadata, match variables to variable groups as defined by GESIS-ZACAT, filter out technical and protocol variables,
+0. Read survey and variable metadata
 
-1. Standardize the spelling of certain often abbreviated words (e.g., 'gov', 'govnmt', and 'govmnt' for 'government'),
+    a. Match variables to variable groups as defined by GESIS-ZACAT, filter out technical and protocol variables
 
-2. Tokenize,
+1. Standardize variable labels
 
-3. Eliminate stop words (but be careful to not eliminate too many, e.g. 'right' or 'working'),
+    a. Standardize the spelling of certain often abbreviated words (e.g., 'gov', 'govnmt', and 'govmnt' for 'government')
 
-4. Create a document-term matrix (actually, variable label - keyword matrix),
+    b. Tokenize, eliminate stop words (but be careful to not eliminate too many, e.g. 'right' or 'working')
 
-5. Sort the columns according to the declining frequency of the keywords,
+    c. Create a document-term matrix (actually, variable label - keyword matrix)
 
-6. Create standardized variable labels by concatenating keywords.
+    d. Sort the columns according to the declining frequency of the keywords
+
+    e. Create standardized variable labels by concatenating keywords
+
+2. Identify identical or equivalent sets of value labels
 
@@
 
-The table contains metadata from 97 Eurobarometer files (each file corresponding to one EB round).
-The metadata are extracted from SPSS files with [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html).
+The table `eb_metadata_var_groups` contains metadata from 96 Eurobarometer files (each file corresponding to one EB round) combined with information on variable groups from [GESIS ZACAT](https://zacat.gesis.org/webview/).
 
 ```{r}
-metadata_database <- readRDS( file.path('../data-raw', 'eb_metadata_database_20200628.rds')) %>%
-  mutate(archive_id = substr(filename, 1, 6))
+# uses eb_metadata_database_20200628.rds
+
+eb_metadata_var_groups <- readRDS( file.path('../data-raw', 'eb_metadata_var_groups.rds')) %>%
+  mutate(archive_id = substr(filename, 1, 6)) %>%
+  distinct()
 ```
 
-#### Waves - Archive IDs
+`eb_metadata_var_groups` is created by joining `eb_var_groups_zacat.rds` with 'eb_metadata_database_20200628.rds' after some cleaning of `var_label_orig`.
 
-Mapping of GESIS Archive IDs (as used in file names) to wave numbers, from EB/44.2bis (ZA2828) to EB/92.1 (ZA7579) - 156 EB waves altogether.
+`eb_var_groups_zacat.rds` contains info on variable groups for 157 EB waves from 44.2bis (ZA2828) to 92.1 (ZA7579).
 
-```{r, warning = F}
-eb_names <- rio::import("../data-raw/eb_waves_filenames.xlsx") %>%
-  separate(x, c("archive_id", "x2", "title"), sep = ":") %>%
-  separate(x2, c("wave", "dates"), sep = " \\(|\\)") %>%
-  mutate(wave = gsub("Eurobarometer | ", "", wave))
+Code for this (not run):
 
-eb_names %>%
-  select(archive_id, wave, dates) %>%
-  head() %>%
-  kable %>%
-  kable_styling(bootstrap_options =
-                  c("striped", "hover", "condensed"),
-                  fixed_thead = T,
-                  font_size = 10 )
+```{r, eval = FALSE}
+
+metadata_database <- readRDS( file.path('../data-raw', 
+                                        'eb_metadata_database_20200628.rds')) %>%
+  mutate(archive_id = substr(filename, 1, 6))
+
+eb_var_groups_zacat <- readRDS( file.path('../data-raw', 
+                                          'eb_var_groups_zacat.rds') ) %>% distinct()
+
+archive_ids <- eb_var_groups_zacat %>% distinct(archive_id) %>% pull(archive_id)
+
+eb_metadata_var_groups <- metadata_database %>%
+  # filter only waves for which variable groups are available
+  filter(archive_id %in% archive_ids,
+         # exclude this strange variable
+         var_label_orig != "filename") %>%
+  # valid_range is a list, which messes things up, so let's make it a lower case string
+  mutate(valid_range_str = tolower(as.character(valid_range))) %>%
+  # keep one row per variable
+  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, 
+           var_label_norm, archive_id, valid_range_str, length_cat_range) %>%
+  # original variable labels in the data and in GESIS ZACAT are not exactly the same
+  # --> need some cleaning
+  mutate(var_label_orig2 = gsub("^([A-Z])*[0-9]{1,3}([A-Z]{1,2})* ", "",
+                                var_label_orig),
+         var_label_orig2 = gsub("^[A-Z][0-9] [0-9]{1,2} ", "",
+                                var_label_orig2),
+         var_label_orig2 = gsub("^W3A/W4A ", "", var_label_orig2),
+         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2}([A-Z])*_[A-Z]{1,2} ", "",
+                                var_label_orig2),
+         var_label_orig2 = gsub("^\\s+|\\s+$", "",
+                                var_label_orig2),
+         var_label_orig2 = gsub('“)', "",
+                                var_label_orig2),
+         var_label_orig2 = gsub("  ", " ",
+                                var_label_orig2),
+         # these 2 waves have problems and labels in data and in ZACAT are different
+         var_label_orig2 = ifelse(archive_id %in% c("ZA6695", "ZA6697"),
+                                  gsub("\\s*\\([^\\)]+\\)","", var_label_orig2),
+                                  var_label_orig2)) %>%
+  # join with table with variable groups
+  left_join(eb_var_groups_zacat, by = c("var_label_orig2" = "var_label_zacat", "archive_id"))
 ```
 
-#### Variable groups
+### 1. Standardize variable labels
 
-Mapping of variable labels to variable groups, as defined in [GESIS ZACAT](https://zacat.gesis.org/webview/), for 51 waves from EB 79.1 (ZA5687) through EB 92.1 (ZA7579).
+In many multi-wave surveys, such as the Eurobarometer, there is a degree of consistency in how variables are labelled in different editions of the survey, and other survey projects also try to follow some labelling conventions.
+
+This consistency is limited and conventions are not followed strictly, which makes standardizing labels a semi-manual task.
+Some of the issues stem from the limits in the number of characters of variable labels in common statistical software packages (e.g., SPSS), which leads to non-standard abbreviations.
+
+#### 1.1. Translation tables ...
+
+The dictionary of variable labels is a project-specific (i.e., specific for the Eurobarometer) table that standardizes the spelling of words and abbreviations used in variable names.
 
 ```{r}
-varlabs <- rio::import("../data-raw/eb_variable_groups_zacat.xlsx", sheet = 4) %>%
-  mutate(wave_desc = ifelse(substr(var_label_zacat, 1, 13) == "Eurobarometer", var_label_zacat, NA),
-         var_label_zacat = ifelse(substr(var_label_zacat, 1, 13)=="Eurobarometer", NA, var_label_zacat),
-         # group names are writte in camel case and != toupper(group names)
-         group = ifelse(var_label_zacat != toupper(var_label_zacat), var_label_zacat, NA),
-         # exceptions from the rule identifying group names
-         group = ifelse(substr(group, 1, 9) == "OWNERSHIP", NA, group), 
-         group = ifelse(substr(group, 1, 19) == "POSITIVE EU RESULTS", NA, group),
-         group = ifelse(substr(group, 1, 24) == "OPINION LEADERSHIP INDEX", NA, group),
-         var_label_zacat = ifelse(!is.na(group), NA, var_label_zacat)) %>%
-  fill(wave_desc, group) %>%
-  drop_na(group, var_label_zacat) %>%
-  mutate(wave = substr(wave_desc, 1, 18),
-         wave = gsub("^Eurobarometer ", "", wave),
-         var_label_zacat = gsub("^W[0-9]{1,2}[A-Z] |^W3A/W4A ", "", var_label_zacat)) %>%
-  left_join(eb_names) %>%
-  select(wave, archive_id, group, var_label_zacat, everything())
-
-varlabs %>%
-  select(archive_id, wave, group, var_label_zacat) %>%
-  head() %>%
-  kable %>%
-  kable_styling(bootstrap_options =
-                  c("striped", "hover", "condensed"),
-                  fixed_thead = T,
-                  font_size = 10 )
-```
-
-```{r}
-# vector of archive IDs
-archive_ids <- varlabs %>% distinct(archive_id) %>% pull(archive_id)
-
-length(archive_ids)
-```
-
-### 1. Standardize the spelling
-
-All procedures below applied to 51 waves, for which all metadata are available.
-
-#### 1.1. Translation table ...
-
-```{r}
-dict <- rio::import("../data-raw/var_lab_dict.xlsx", sheet = "dict") %>%
+var_dict <- import( file.path('../data-raw', 'var_lab_dict.xlsx'), 
+                sheet = "dict") %>% 
   gather(id, from, 2:6) %>%
@@
 
-head(dict) %>%
+head(var_dict) %>%
   kable %>%
@@
 
-dat2 <- metadata_database %>%
-  # filter only waves for which variable groups are available
-  filter(archive_id %in% archive_ids,
-         # exclude this strange variable
-         var_label_orig != "filename") %>%
-  # keep one row per variable
-  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm, archive_id) %>%
-  # original variable labels in the data and in GESIS ZACAT are not exactly the same 
-  # --> need some cleaning
-  mutate(var_label_orig2 = gsub("^W[0-9]{1,2}", "", var_label_orig),
-         var_label_orig2 = gsub("^W[0-9]{1,2}[A-Z]", "", var_label_orig),
-         var_label_orig2 = gsub("^[0-9]{1-3} ", "", var_label_orig2),
-         var_label_orig2 = gsub("^[A-Z][0-9]{1,2}[A-Z] ", "", var_label_orig2),
-         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2} ", "", var_label_orig2),
-         var_label_orig2 = gsub("^[A-Z]{1,2}[0-9]{1,2}[A-Z] ", "", var_label_orig2),
-         var_label_orig2 = gsub("^[0-9]{1,2} ", "", var_label_orig2),
-         var_label_orig2 = gsub("^[A-Z][0-9] [0-9]{1,2} ", "", var_label_orig2),
-         var_label_orig2 = gsub("^W3A/W4A ", "", var_label_orig2)) %>%
-  # join with table with variable groups
-  left_join(varlabs, by = c("var_label_orig2" = "var_label_zacat", "archive_id")) %>%
+dat2 <- eb_metadata_var_groups %>%
   # remove variables in the following groups
-  filter(!group %in% c("Archive and survey ID variables",
-                       "Country group variables",
-                       "Nationality (Q1)",
-                       "Protocol variables (P)",
-                       "Standard nation ID variables",
-                       "Weight variables (W)",
-                       "Nation ID variables",
-                       "Protocol variables",
-                       "Weight variables",
-                       "Archive and ID variables",
-                       "Household Weights"),
+  filter(!grepl("protocol|weight|nationality|nation id|original id|id variable|country group", 
+                tolower(var_group)),
          # remove recoded variables
@@
          # leftover question numbers at the beginning of value labels
-         var_label_norm1 = gsub("^w[0-9]{1,2}", "", var_label_norm),
-         var_label_norm1 = gsub("^[0-9]{1-3}_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[0-9]{1,2}_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z]_", "", var_label_norm1),
-         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", var_label_norm1)) %>%
-  # apply the translation table
+         var_label_norm1 = gsub("^w[0-9]{1,2}", "", 
+                                var_label_norm),
+         var_label_norm1 = gsub("^[0-9]{1-3}_", "", 
+                                var_label_norm1),
+         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", 
+                                var_label_norm1),
+         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}_", "", 
+                                var_label_norm1),
+         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", 
+                                var_label_norm1),
+         var_label_norm1 = gsub("^[0-9]{1,2}_", "", 
+                                var_label_norm1),
+         var_label_norm1 = gsub("^[a-z]_", "", 
+                                var_label_norm1),
+         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", 
+                                var_label_norm1)) %>%
+    # apply the variable label dictionary
   mutate(var_label_norm1 = stri_replace_all_regex(var_label_norm1,
-                                                  pattern = dict$from,
-                                                  replacement = dict$to,
+                                                  pattern = var_dict$from,
+                                                  replacement = var_dict$to,
                                                   vectorize_all = FALSE)) %>%
-  # replace underscore with space
-  mutate(text = gsub("_", " ", var_label_norm1)) %>%
   ungroup() %>%
   # select the needed variables
-  select(filename, qb, var_label_orig, var_label_norm, var_label_norm1, text, var_id, group)
-
+  select(filename, qb, var_label_orig, var_label_orig2, var_name_orig,
+         var_label_norm, var_label_norm1, var_id, var_group, 
+         valid_range_str, length_cat_range)
 ```
@@
 ```{r}
-# list of country codes to be converted to country names
-eu_countries <- c(
-  "at", "be", "bg", "cy", "cz", "dk", "de", "ee", "es", "fr",
-  "fi", "gb", "gr", "hr", "hu", "ie", "it", "lt", "lv", "lu",
-  "mt", "nl", "pl", "pt", "ro", "se", "si", "sk", "tr")
-
-# !!! the problem here is that dk sometimes means "don't know" and sometimes Demnark !!!
-# this needs to be treated in gesis_metadata_create, where the
-# user defined missings are separately analyzed, they dk should 
-# be converted there to do_not_know.
-
-# it's not clear hot to do this, because both types of DK appear at the end of labels e.g. in
-# mentioned/not mentioned questions, and they look very much the same
-
 tidy <- dat2 %>%
   # tokenize
-  unnest_tokens(word, text, 
-                token = "regex", pattern = " |  ") %>%
+  unnest_tokens(word, var_label_norm1, 
+                token = "regex", 
+                pattern = "_",
+                drop = FALSE) %>%
   # eliminate stop words, with some exceptions
   anti_join(stop_words %>%
-              filter(!word %in% c("working", "right", "other")
+              filter( !word %in% c("working", "right", "other")
                      )) %>%
   # eliminate numbers
-  filter(!grepl("\\b\\d+\\b", word)) %>%
-  # convert country codes to country names
-  mutate(
-    country = ifelse(word %in% eu_countries, word, NA),
-         country = countrycode::countrycode(
-           toupper(country), "iso2c", "country.name"),
-    country = tolower(country),
-    word = ifelse(!is.na(country), country, word),
-    word = ifelse(word == "united kingdom", "united-kingdom", word),
-    word = ifelse(word == "tcc", "turkish-cypriot-community", word)
-    ) %>%
-  select(-country)
+  filter( !grepl("\\b\\d+\\b", word)) 
 ```
@@
 
-Note: this Denmark is probably a "don't know"
+### 4. Create a document-term matrix
+### 5. Sort the columns according to the declining frequency of the keywords
+### 6. Create standardized variable labels by concatenating keywords
 
-### 4. Create a document-term matrix
-
-Filter out labels that have words that appear in < 5 different waves. If we want to create a trend file, we're only interested in variables that are repeated often.
+First, we filter out labels that have words that appear in < 5 different waves. If we want to create a trend file, we're only interested in variables that are repeated often.
 
 ```{r}
-tidy_small_2 <- tidy %>%
+tidy_small <- tidy %>%
   group_by(word) %>%
@@
   # flag words that appear in < 5 different files
-  mutate(to_drop = as.numeric(n_wave < 10)) %>%
+  mutate(to_drop = as.numeric(n_wave < 5)) %>%
   # group by variable ID (file x variable name)
@@
   ungroup()
+```
 
-dtm <- tidy_small_2 %>%
+The next steps are performed separately within variable groups, and programmed as a loop.
+
+Note that a change to the earlier process is that **selected words** (so far only "cat" which stands for "categories", I presume) **are placed at the end of the document-term matrix**, even though they tend to occur frequently, so that they are always at the end.
+
+```{r}
+# split into list by `group`
+tidy_small_list <- split( tidy_small , f = tidy_small$var_group )
+
+# list of words that go last
+words_last <- c("cat")
+
+# create empty list for standardized variable names
+dtm2_list <- list()
+
+for (i in names(tidy_small_list)) {
+
+  # create a document-term matrix within each variable group
+  dtm <- tidy_small_list[[i]] %>%
     distinct(var_id, var_label_norm1, word) %>%
@@
 
-dim(dtm)
-```
-
-### 5. Sort columns according to the declining frequency of the keywords
-
-```{r}
-# get the order of variables according to declining proportion of missing value
-# (declining popularity of keywords)
+  # find the order of declining frequency of the keywords
   order <- dtm %>%
-  summarise_all(funs(sum(is.na(.))/length(.))) %>%
+    summarise_all(
+      list(~ sum(is.na(.))/length(.))
+    ) %>%
     t() %>% as.vector()
@@
   names(order) <- names(dtm)
-```
 
-### 6. Create standardized variable labels by concatenating keywords
-
-Standardized variable labels `var_label_std` are concatenated keywords, where the keywords occur in the order of their frequency in the entire corpus of variable labels.
-As a consequence, more common - likely broader - terms will come first, and less common - more specific - terms will come later.
-
-```{r}
-dtm2 <- dtm %>%
-  select( names(sort(order)) ) %>%
-  mutate_at(vars(3:ncol(dtm)), 
-            funs(ifelse(is.na(.), "", .))
+  # concatenate keywords to create standardized variable names
+  dtm2_list[[i]] <- dtm %>%
+    select( names(sort(order)), 
+            # !!! put selected words last
+            -any_of("cat"), any_of(words_last) ) %>%
+    mutate_at( vars(3:ncol(dtm) ),
+               list( ~ifelse(is.na(.), "", .))
     ) %>%
@@
 
-dim(dtm2)
+}
+
+dtm2 <- bind_rows(dtm2_list)
 ```
 
-These standardized variable names still have problems that need to be fixed.
+Standardized variable labels `var_label_std` are concatenated keywords, where the keywords occur in the order of their frequency in the entire corpus of variable labels.
+As a consequence, more common - likely broader - terms will come first, and less common - more specific - terms will come later.
 
 ```{r}
-dtm2 %>% 
+metadata_standard_var_names <- dtm2 %>%
+  left_join(dat2)
+```
+
+```{r}
+metadata_standard_var_names %>% 
+  select(var_id, var_label_std, var_group) %>%
   head(., 25) %>%
@@
 ```{r}
-dtm2 %>%
+metadata_standard_var_names %>%
   count(var_label_std) %>%
@@
 # filter on var_label_std
-dtm2 %>%
-  filter(grepl("trust|satisf|left-right|interest|class", var_label_std)) %>% 
+metadata_standard_var_names %>%
+  filter( grepl("trust|satisf|left-right|interest|class",
+                var_label_std)) %>% 
   count(var_label_std) %>%
@@
 ```
+
+### 2. Standardize value labels
+
+The dictionary of value labels identifies identical or equivalent sets of value labels.
+
+```{r}
+val_dict <- import( file.path('../data-raw', 'val_lab_dict.xlsx'),
+                sheet = "dict") %>%
+  mutate(from = tolower(from)) %>%
+  select(from, to)
+
+head(val_dict) %>%
+  kable %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 )
+
+```
+
+So far, seven types of response scales are identified:
+
+1.
+`mentioned / not mentioned` - coding of responses to questions where the Respondent is asked to choose one or more elements from a list of possibilities; these variables account for the majority of variables in the Eurobarometer, but -- interestingly -- are not generally found in other cross-national survey projects,
+2. `trust / no trust` - binary response scales to questions about trust in institutions,  
+3. `national government / European Union` - binary responses to questions about whether national governments or the European Union have authority over a given issue domain,  
+4. `yes / no` - binary response scales (exclude types 1-3),  
+5. `positive / negative / neutral` - three-point scales, where the order of options is: positive, negative, neutral,  
+6. `positive / neutral / negative` - three-point scales, where the order of options is: positive, neutral, negative,  
+7.
+`ordinal, 4 points` - four-point scales that contain ordinal evaluations of a phenomenon, starting from the most positive, typically on a bipolar scale, e.g.: very positive, fairly positive, fairly negative, very negative.
+
+```{r}
+metadata_standard_var_names_scales <- metadata_standard_var_names %>%
+  # use values dictionary to tag variables to values_type
+  mutate(values_type = stri_replace_all_fixed(as.character(valid_range_str),
+                                                  pattern = val_dict$from,
+                                                  replacement = val_dict$to,
+                                                  vectorize_all = FALSE)) %>%
+  # identify "mentioned/not mentioned" as a separate type
+  mutate(values_type = ifelse(
+           grepl(".*not mentioned.*", valid_range_str), 
+           "mentioned", 
+           values_type )) %>%
+  # if type == valid_range, nothing has been learnt, so replace with NA
+  mutate(values_type = ifelse(
+           values_type == as.character(valid_range_str), 
+           NA, 
+           values_type)
+         )
+
+metadata_standard_var_names_scales %>% 
+  count(values_type) %>%
+  kable %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 )
+
+```
+
+## Application, ctd.: political attitudes
+
+Filter on substrings of standardized variable labels `var_label_std`, and count the repetitions across datasets.
+
+```{r}
+# filter on var_label_std
+metadata_standard_var_names_scales %>%
+  filter( grepl("trust|satisf|left-right|interest|class",
+                var_label_std)) %>% 
+  count(var_label_std, values_type) %>%
+  arrange(desc(n)) %>%
+  head(., 25) %>%
+  kable %>%
+  kable_styling(bootstrap_options =
+                  c("striped", "hover", "condensed"),
+                  fixed_thead = T,
+                  font_size = 10 )
+```
